### 2021-05-27 docs/manuscript.Rmd
Add writing

---
title: "Linkage of trial registrations and results publications: A cross-sectional, automated evaluation of German university medical centers"
author: "Maia Salholz-Hillel, Daniel Strech, Benjamin Gregory Carlisle"
output: 
  bookdown::word_document2:
    reference_docx: quest-word-template.docx
    latex_engine: xelatex
bibliography: references.bib
editor_options:
    markdown:
        canonical: true
    chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
ggplot2::theme_set(ggplot2::theme_minimal())

library(dplyr)
library(tidyr)
library(stringr)
library(glue)
library(purrr)
library(here)
library(readr)

library(ggplot2)
library(DiagrammeR)
library(gtsummary)
# library(papaja)

# Get data
trials_all <- read_rds(here("data", "processed", "trials.rds"))
# cross_registrations_all <- read_rds(here("data", "processed", "cross-registrations.rds"))
pubmed_all <-
  read_rds(here("data", "processed", "pubmed-main.rds"))

# Source analyses
source(here("R", "functions", "screening.R"))
# # source(here("R", "functions", "report-results.R"))
source(here("R", "analysis", "01_screen-trials.R"))
# # source(here("R", "analysis", "02_flowchart-trials.R"))
source(here("R", "analysis", "03_tbl-trials.R"))
source(here("R", "analysis", "04_link-reg-pub.R"))
# # source(here("R", "analysis", "05_upset-link-reg-pub.R"))
source(here("R", "analysis", "06_plot-change-over-time.R"))
source(here("R", "analysis", "trn-regex-analysis.R"))
source(here("R", "analysis", "cross-registrations.R"))
# # fs::dir_ls(here("R", "analysis")) %>% 
# #   purrr::walk(source)
```

# Abstract

**Objective** To evaluate, using a novel automated approach, the linkage of registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). The International Committee of Medical Journal Editors (ICMJE) and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linkage to promote discoverability of both the clinical trial publication and registration and thus increase the value of clinical research.

**Design** Cross-sectional bibliographic study

**Setting** German UMC clinical trials

**Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered clinical trials conducted across German UMCs and completed between 2009 and 2017. We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and TRNs in publication metadata, abstract, and full-text. <!--# We verified our automated approach on a manually coded sample. -->

**Main outcome measures** The proportion of trial registrations incorporating a structured reference (DOI or PMID) to a known results publication. The proportion of clinical trial results publications that reported the known TRN in the metadata, abstract, and full-text.

**Secondary analyses** We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and publication linkage in the registration. We also explored the relationship between cross-linkage<!--# practices --> and registry. We also explored the relationship between cross-linkage<!--# practices --> and prospective registration <!--# not sure about this. i think if we do this then glm for the various practices -->. We looked at change in registration-publication linkage over time. We also looked at cross-registrations in the registries as well as additional TRNs reported in trial publications.<!--#  We also looked at reporting PMC-version full-text. We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status. We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting. All analyses are exploratory. -->

**Results** Of `r n_trials_ft_pdf` trials with PubMed-indexed clinical trial results publications available as full-text, we found `r report_n_pct_trials(n_trials_link_full)` trials with comprehensive registration-publication linkage, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links. We found that only `r report_n_pct_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, whereas `r report_n_pct_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials did so, mostly via automatic publication indexing. TRNs were most commonly reported in the full-text, appearing in `r report_n_pct_trials(n_trials_link_trn_ft)`. Linkage rates appear overall lower for DRKS trials than for ClinicalTrials.gov. The most common link is TRN reporting in the full-text which reached `r stringr::str_extract(report_link_year("TRN (Full-text)", 2017), "\\d{2}%")` in 2017 in both registries. <!--# add more secondary? -->

**Conclusion** Despite encouragement from CONSORT and ICMJE, German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications. Publication linkage was noticeably greater in ClinicalTrials.gov which uses a more structured data entry process as well as automatically indexes publications from PubMed. German and international university medical centers (UMCs) should incentivize, promote, and educate their researchers on this responsible research practice. Automated strategies, such as those used in this study, could support bibliometric databases and registries to enrich their registration-publication linkage. For example, to improve TRN inclusion in its metadata, PubMed could automatically extract TRNs from publication abstracts or full-text. DRKS could adopt strategies like those used by ClinicalTrials.gov to automatically index publications reporting the TRN in PubMed. As compared with other responsible research practices (such as trial registration itself), reporting TRN and publications are copy-and-paste tasks that may be integrated into biomedical researchers' existing workflows.

**Registration** This study was not preregistered. <!--# The study protocol is available at [insert link]. -->

**Open Data and Code** All code and the final analysis data for this study are available at [insert link]. Raw data (with the exception of the full-text of publications) is available at [insert link].

# Introduction

Clinical trial registration and results reporting are important tools for clinical research transparency [@worldmedicalassociation2013; @worldhealthorganization2017].<!--# I removed background on registration and results reporting for succinctness and because there's a robust literature, which I can add additional citations to. Let me know if you think it's worth adding back a few sentence of background. --> Linkage of the trial registrations and results publications promotes discoverability of both components <!--# adds further transparency beyond registration and results reporting independently -->and increases the value of clinical research. Such threaded evidence <!--# Linked registrations-publications-->empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd]. Robust linkage includes both the incorporation of a structured reference to publications in the registration, and reporting of trial registration numbers (TRN) in the full-text, abstract, and metadata of publications.

The Consolidated Standards of Reporting Trials (CONSORT) as well as the ICMJE's *Recommendations for the Conduct, Reporting, Editing, and Publication of Scholarly Work in Medical Journals* call for reporting "trial registration number and name of the trial register" in both the full-text and abstract of trial results publication [@internationalcommitteeofmedicaljournaleditorsicmje2019; @hopewell2008pm]. Reporting TRN solely in the full-text does not suffice since readers may not have access to the full-text or may base their judgment of the trial or topic on the abstract alone [@hopewell2008tl]. Abstracts may also be published independently of a full-text publication, such as for conference abstracts, highlighting the need for TRN reporting in abstracts as well [@hopewell2008pm].

Furthermore, the inclusion of TRNs in publication and bibliographic metadata enhances discoverability for both human readers and computerized analysis. For example, PubMed MEDLINE metadata includes TRNs under the "secondary source id" tag [@nationallibraryofmedicine2018], which can then be searched on the web interface (e.g., "NCT00002568[secondary identifier]" or "NCT00002568[si]") or accessed programmatically via the underlying XML [@nationallibraryofmedicine2017c]. Currently, these tags can be generated from publisher-provided data or assigned manually by NLM Indexing Data Reviewers who copy-and-paste secondary source ids (including trial ids) found in the abstract and full-text (personal communication via NLM Helpdesk Ticket CAS-552810-T3H7V5). <!--# @daniel: this had been in the intro but you said to shorten so i moved to discussion but then you had several quetion about how SI is made, so moved it back to intro... -->

Machine-readable TRN metadata moreover facilitates automatic incorporation of the publication in the registry for "easy linking of trial related publications with clinical trial registry site records" [@nationallibraryofmedicine2017c]. For example, ClinicalTrials.gov uses PubMed secondary identifiers to automatically index publications in the registry entry under "Additional publications automatically indexed to this study by ClinicalTrials.gov Identifier (NCT Number)" [@nationallibraryofmedicine2017c]. This therefore increases the discoverability of the publication from the registration, with no additional effort from the trial investigator for manually adding the publication to the registry.

Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings. <!--# add policy background on why pub in reg --> The insertion and structuring of references depends on the registry; whereas ClinicalTrials.gov includes both automatically indexed publications and manual entries in quasi-structured datafields, DRKS currently accepts only manual entries in minimally structured datafields. References, like those automatically indexed, may include one or more publication unique identifier such as a DOI or PMID, or may be limited to a free-text citation with journal, article, and author names, but no unique identifier. While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval and linkage [@huser2013po].

Trial evidence appears to be insufficiently linked in both registries and publications. @huser2013po found that only 28% of 8907 completed trials contained a structured trial-article link with a publication unique identifier in the registry. In a systematic review of 16 studies examining cohorts of trial publications for links to registrations<!--# of results publications with associated registrations/investigating linked registrations-publications in publication cohorts found via both automated and manual approaches...couldn't find the overall TRN reporting rate from all studies looking at TRN, only the 16 using both automated and manual approaches, but this level of detail feels like too much to explain -->, @bashir2017sr found an overall rate of 49% of results publications <!--# (range: 8-97%) --> reported a TRN in either the metadata, abstract, or full-text. More recently, @carlisle2020m found 38,001 ClinicalTrials.gov ids across 488,365 abstracts from records indexed by PubMed as clinical trial publications, meaning a maximum of `r round(38001/488365 * 100, digits = 1)`% of publications report a ClinicalTrials.gov id, assuming one TRN per publication. @aldurra2020b found that only 28.8% (7473/10,500) records classified as RCTs based on a bespoke PubMed query reported trial ids in any of the metadata, abstract, or full-text. <!--# @murph sanity check of numbers from other papers -->

These previous projects, however, faced limitations due to the unclear publication cohort and unknown true registration rates. PubMed queries currently do not suffice to identify the true populations of clinical trials or even RCTs, both excluding true positives (i.e., sensitivity) and including false positives (i.e., specificity) [@glanville2020jmla]. Therefore not all PubMed-indexed clinical trials or RCTs should be registered <!--# as they may not reflect true clinical trials or RCTs -->, and not all that should be are. Furthermore, not all reported TRNs correspond to bona fide registrations of trial whose results are reported in the publication, i.e., TRNs may be reported for related trials in the background section. As such, conclusions drawn by these previous studies relying on PubMed-query based samples, about registration reporting rates must be qualified as the rate of reported trial registration numbers in PubMed-indexed (randomized) clinical trials with unknown registration rates. <!--# remove previous sentence? --> In addition, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link. These studies of publication linkage in registrations have furthermore focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of other studies investigating publication linkage in DRKS.

## Objectives

We evaluated the linkage of registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017. <!--# @daniel comment: add reference to intovalue here? it's in methods -->We also looked at change in these linkage practices over time. By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication linkage, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata. We use an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs. We did not pre-specify the analyses.

<!--# We also explored the reporting of additional TRNs, beyond the known associated TRN, in publications, including reported cross-registrations, meaning more than one TRN in more than one registry associated with a single publication. We also looked at reporting PMC-version full-text.  -->

# Methods

## Data sources and collection

We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset. In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.

We <!--# programmatically -->downloaded data from both registries on 15 May 2021. We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data. We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti]. As DRKS does not provide an API, we built a webscraper to capture the necessary fields [@federalinstitutefurdrugsandmedicaldevicesbfarm]. ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information. For example, ClinicalTrials.gov has a separate summary results reporting section, whereas summary results are included with other references in DRKS. We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse"). <!--# remove if not including summary results -->

We then queried the PubMed Entrez Programming Utilities (E-utilities) API via an R client for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter]. From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata, as well as MeSH terms, keywords, author details, and summary metadata. <!--# remove unused pubmed components? -->

We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies? --> [@ourresearch; @jahn]. We then used the GROBID machine learning library for technical and scientific publications to parse the PDFs into machine-readable XMLs [@2008]. In order to isolate the main body from the publication abstract for our analysis, we extracted the `<body>` sections of the papers.

Figure \@ref(fig:flow-methods) show the methods flow using automated techniques with the exception of some full-text PDFs.

```{r flow-methods, fig.cap = "Data collection sources and methods", out.width="50%"}
knitr::include_graphics(here("docs", "figures", "flow-methods.png"))
```

### Exclusion/inclusion criteria

After updating registry data, we reapplied the IntoValue exclusion criteria to exclude trials that were completed before 2009 or after 2017, were non-interventional, or did not meet the recruitment or study status inclusion criteria (see @wieschowski2019joce and @riedel2021 for further details on these criteria). We also excluded trials without a German UMC lead, as only one of the two IntoValue cohorts included trials not led by German UMCs. For duplicate trials (i.e., appearing in both IntoValue cohorts), we excluded the earlier record. We also excluded trials without results, and trials with results for which we could not find a PMID and full-text publication.

## Detection and classification of publication identifiers in registrations

As reference data is available in difference formats depending on the registry, we developed parallel approaches to extract publication identifiers for each. For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT. We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field. For DRKS, we scraped the reference type, citation, and URL link, when available. We used regexes to extract any DOIs and PMIDs from either the citation or link. In the rare cases when conflicting DOIs or PMIDs were found in the citation vs. the link, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation. In addition, we determined whether a reference was manually or automatically indexed in the registry. DRKS allows only manually added references; for ClinicalTrials.gov, references of type "derived" were marked as automatically indexed. We then used DOIs and PMIDs to match the referenced publications to those in our trial dataset. We did not attempt to match publications without identifiers (i.e., publication title only).

## Detection and classification of trial registration numbers in publications

We developed <!--# and evaluated [or more methods details] --> regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(trn-regular-expressions) Table \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text. To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata for true positives and true negatives.<!--# In order to isolate the full-text main body from the publication abstract for our analysis, we included only `<body>` sections of the PMC and PDF samples for the full-text source. --> For each source, we first extracted all unique TRN patterns within a source. Our regex allowed for minor formatting errors in the TRN patterns (such as additional punctuation). We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source introduced through the cleaning process. We also coalesced across sources to deduce a list of unique TRNs by publication along with the source(s) in which each was reported.

The output of this TRN extraction, cleaning, and deduplication included TRNs beyond those of the known registrations. These additional TRNs could be cross-registrations of the same trial or provided as background or discussion. For this study, we were interested in whether known registrations were reported in the publication, so we filtered for reported TRNs matching the known registrations. <!--# change this if considering cross-reg -->

<!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration. currently placeholder "TRUE" such that all TRNs resolve in the registries -->

<!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns? then would have to change my backend work. or can keep this section but maybe not report all the details in text but just appendix? -->

## Software, code, and data

Data collection, preparation, and analysis was performed in R version 4.0.2 (R Foundation for Statistical Computing) and RStudio version 1.4 (RStudio Team). PDF conversion to TEI XML was performed using GROBID Server 0.6.1 [@2008] and GROBID Python Client [@lopez2021]. All code and the final analysis data for this study are available at [insert link]. Raw data (with the exception of the full-text of publications) is available at [insert link].

# Results

## Trial and publication characteristics

The IntoValue trials dataset included the full sample of trials led by a German UMC that were registered on ClinicalTrials.gov or the DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria <!--# excluding non-German UMC-led trials, deduplicating trials, excluding trials without results, enhancing the publication metadata, getting full-text, and excluding trials with results without a PMID and full-text -->, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications. <!--# QUESTION: what summary info about trials and publications? I also include a long version of demographic data in case we want to include any of these, ideally with collapsed categories, in Table @ref(tab:tbl-trials-full) -->

```{r flow-trials, fig.cap = "Trials and publications screening", out.width="50%", out.height="50%"}
knitr::include_graphics(here("docs", "figures", "flow-trials.png"))
```

```{r tbl-trials}
tbl_trials %>% 
  gtsummary::as_flex_table() %>% 
  flextable::set_caption(caption = "Characteristics of included trials with results publications")
```

## Registration-publication linkage

While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linkage, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication. Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the breakdown of registration-publication linkage across the various components for ClinicalTrial.gov and DRKS trials.

```{r plot-upset-links-reg-pub-registries, fig.cap = "Registristion-publication linkages in ClinicalTrials.gov and DRKS", out.width="80%"}
knitr::include_graphics(here("docs", "figures", "plot-upset-links-reg-pub-registry.png"))
```

### Publications linked in registrations

We found that `r report_pct_num_denom_trials(n_trials_link_reg)` registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results. Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed.

<!--# @daniel: you had several comments about automatically linked pubs so I don't think the language I had was clear. I tried several wordings below. Let's discuss and figure out what is clearest. -->

Of ClinicalTrials.gov trials with the TRN in the PubMed metadata (n = `r n_trials_ctgov_link_reg_trn_si`), `r report_n_pct_trials(n_trials_ctgov_link_reg_trn_si_manual + n_trials_ctgov_no_link_reg_trn_si, n_trials_ctgov_link_reg_trn_si)` were not automatically indexed.

Of ClinicalTrials.gov trials with the TRN in the PubMed metadata (n = `r n_trials_ctgov_link_reg_trn_si`), `r report_n_pct_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_ctgov_link_reg_trn_si)` have the publication automatically linked in the registry, `r report_n_pct_trials(n_trials_ctgov_link_reg_trn_si_manual, n_trials_ctgov_link_reg_trn_si)` have the publication manually linked in the registry, and only `r report_n_pct_trials(n_trials_ctgov_no_link_reg_trn_si, n_trials_ctgov_link_reg_trn_si)` trial is not linked in the registry.

All `r n_trials_ctgov_link_reg_trn_si_derived` trials with automatically indexed publications in ClinicalTrials.gov include the TRN in the PubMed metadata. An additional `r n_trials_ctgov_link_reg_trn_si_manual` trials include the TRN in the PubMed metadata and a link publication that was not automatically indexed, as it may have been entered by the trialist manually prior to being indexed. `r if (n_trials_ctgov_no_link_reg_trn_si == 1) "Only one"` ClinicalTrials.gov trial (NCT01614301) has a TRN in the PubMed metadata (PMID: 26417987) and no linked publication in the registration. While the TRN is indeed indexed in the secondary identifier field, PubMed mistakenly associated it with an incorrect databank (i.e., GENBANK) which presumably prevented ClinicalTrials.gov from making the link.

When limiting to only manually linked publications in the registration, performance is similar between ClinicalTrials.gov (24%, 186/783) and DRKS (22%, 96/447). <!--# change to code -->

### Registrations linked in publications

We found that `r report_pct_num_denom_trials(n_trials_link_trn_none)` trials did not report the known TRN in any of the metadata, abstract, or full-text of the corresponding publication. Only `r report_pct_num_denom_trials(n_trials_link_trn_all)` reported the TRN across all three locations. TRNs were most commonly reported in the full-text, appearing in `r report_n_pct_trials(n_trials_link_trn_ft)`. TRNs were included in `r report_n_pct_trials(n_trials_link_trn_abs)` of abstracts, as compared with `r report_n_pct_trials(n_trials_link_trn_si)` of PubMed metadata.

### Linkage by registry

We found that trials with a ClinicalTrials.gov TRN were significantly more likely than trials with a DRKS TRN to include a publication reference in the registry (`r apa::apa(chisq_link_by_reg_reg, print_n = TRUE)`) and to include the TRN in the PubMed metadata (`r apa::apa(chisq_link_by_reg_si, print_n = TRUE)`) and abstract (`r apa::apa(chisq_link_by_reg_abs, print_n = TRUE)`). We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`). Table \@ref(tab:tbl-link-by-reg) shows summary registration-publication linkage by registry.

<!--# point estimtes in table should have CI, right? -->

```{r tbl-link-by-reg}
tbl_link_by_reg %>% 
  gtsummary::as_flex_table() %>% 
  flextable::set_caption(caption = "Registration-publication linkage by registry")
```

### TRN reporting in PubMed metadata

<!--# @daniel asked me to test relationship between trn in metadata and abs/ft. many options. total in abs 714 and ft 1131. chisquare trn reporting abs given diff combos of abs/ft reporting OR just abs or ft? ideally, compare registry, trn in abs/ft (not sure what levels), and trn in si, but 3 variable so not chisquare-->

For the 1,372 trials that provided a TRN in either the abstract or full-text, a TRN was also available in the metadata in 55% (n = 767). ClinicalTrials.gov TRNs (65%, 705/1077) were more likely than DRKS TRNs (21%, 62/295) to be included in PubMed metadata. Furthermore, 97% (563/580) of trials with a ClinicalTrials.gov TRN in the abstract also had the TRN in the PubMed metadata, compared with 23% (31/134) of trials with a DRKS TRN. Similarly, 58% (507/878) of trials with a ClinicalTrials.gov TRN in the full-text also had the TRN in the PubMed metadata, compared with 18% (48/253) of trials with a DRKS TRN.

A chi-square test of independence was performed to examine the relation between registry and TRN in PubMed metadata, for trials with TRN reported in both the abstract and the full-text. The relation between these variables was significant, X2 (1, N = 473) = 201.0, p = 1.07e-45. Trials with a ClinicalTrials.gov TRN (Proportion with TRN in SI = 95.5%) were more likely than trials with a DRKS TRN (Proportion with TRN in SI = 33.7%) to have a TRN in PubMed metadata if reported in both the abstract and the full-text (Difference in reporting = 61.8%, CI = 52.5, 71.4).

### TRN reporting given prospective registration

<!--# Are TRNs reported more or less often in abstract OR full text for prospective or retrospective registrations? -->

A chi-square test of independence was performed to examine the relation between prospective registration and TRN reporting in either the abstract or the full-text (abstract-only, full-text-only, both, neither). Registration was categorized as prospective if the trial was registered in the same or preceding months to trial start, since trial start on ClinicalTrials.gov is sometimes available with month and year only. Two trials (NCT02127736, NCT02213406) without start dates and were excluded from this analysis as we could not determine whether registration was prospective. The relation between these variables was significant, X2 (1, N = 1894) = 22.8, p = .00000182. Prospectively registered trials (Proportion reporting = X) were more likely than retrospectively registered trials (Proportion reporting = X) to report the TRN in either the abstract and/or the full-text (Difference in reporting = X, CI = X, X).

### Change in registration-publication linkage over time, by registry

Linkage rates appear overall lower for DRKS trials than for ClinicalTrials.gov. Overall, we observed positive trends for TRN reporting in full-text and abstracts. <!--# add test? --> Figure \@ref(fig:plot-links-year-reg) shows the registration-publication linkages over time for ClinicalTrials.gov and DRKS.

```{r plot-links-year-reg, fig.cap = "Change over time in registration-publication linkages in ClinicalTrials.gov and DRKS. Completion year from registry."}
knitr::include_graphics(here("docs", "figures", "plot-links-year-reg.png"))
```

## Cross-registrations

We found `r report_pct_num_denom_trials(n_trials_crossreg_reg)` of trials indicated a total of `r n_crossreg_reg` cross-registrations as secondary identifiers in their registrations. Of these `r n_crossreg_reg` cross-registrations, X were reported in pubs (X in metadata, X in abs, X in ft). Beyond the IntoValue primary TRNs and these registry-indicated cross-registrations, we detected X additional TRNs in X publications. Table \@ref(tab:tbl-crossreg) shows details on cross-registrations and additional TRNs by registry.

```{r tbl-crossreg}
tbl_crossreg %>% 
  gtsummary::as_flex_table() %>% 
  flextable::set_caption(caption = "Cross-registrations and additional TRNs by registry")
```

## Evaluation of TRN regular expressions

<!--# TODO: add to methods if this is staying? -->

To gauge the sensitivity and specificity of the TRN regexes, we visually inspected all PubMed secondary identifiers extracted for our sample (n = `r nrow(si)`). PubMed's secondary identifier field can contain identifiers not only from clinical trial registries, but also from other databanks, such as molecular sequence and gene expression identifiers or open data repository designators (i.e., figshare, Dryad) [@nationallibraryofmedicine2018]. Most ids were clear true positive TRNs (n = `r si_tp` ) and a handful were clear true negative other ids (n = `r si_figshare`) (i.e., figshare). We also found `r si_drks_misformatted` ids which, based on additional PubMed data, we could determine were DRKS ids, but which were so severely misformatted that on visual inspection alone, we did not classify as TRNs (i.e., missing the preceding letters "DRKS" and just a string of numbers such as "00000711"). As such, we consider a true negative of n = `r si_tn`. Our regexes correctly classified all TRNs and non-TRNs, resulting in a sensitivity and specificity of 100%. If we had instead categorized the `r si_drks_misformatted` severely misformatted DRKS ids as true positives, our regexes would have sensitivity of `r style_percent(si_sens_alt, digits = 1)`% and a specificity of `r style_percent(si_spec_alt)`%.

# Discussion

The cross-linkage of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research. This responsible practice comes at minimal costs to researchers, on the order of seconds for pasting TRNs in papers and minutes for adding a publication link to the registration.

Our study shows that German UMCs can improve in both TRN reporting in publications and references to publications in the registration. In our sample (n = `r n_trials_ft_pdf`), `r report_n_pct_trials(n_trials_link_none)` trials had no links between registration and publication and only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linkage

The issue of TRN reporting in publications is also reflected in CONSORT and ICMJE guidelines. Only `r report_n_pct_trials(n_trials_link_trn_ft_abs)` trials in our sample fully met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract. An additional `r report_n_pct_trials(n_trials_link_trn_ft)` report in full-text only and `r report_n_pct_trials(n_trials_link_trn_abs)` did so in the abstract only. This situation indicates that there is much room for improvement, with involvement across stakeholders, from individual trialists and UMCs, to journals, bibliographic databases, and trial registries.

Reporting appears to have increased over our study time in the full-text and abstracts<!--# while an increase in reporting in abstracts seems to lag behind -->. The upward trend suggests this practice is gaining traction, however more slowly than advisable per CONSORT and ICMJE guidelines. Low TRN reporting poses a barrier to identifying trial publications via automated approaches and instead require researchers to perform intensive manual keyword-based searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].

Trials registered in ClinicalTrials.gov showed better linkage than trials registered in DRKS<!--# or be explicit that not significant for full-text -->. Notably, TRNs in PubMed metadata and publication references in registrations were almost three-fold more frequent in ClinicalTrials.gov than in DRKS.

<!--# n addition to incorporating TRNs provided in publisher data, NLM Indexing Data Reviewers manually assign metadata tags -->

TRNs in PubMed metadata may be either provided by publishers or manually assigned by National Library of Medicine (NLM) staff who copy-and-paste TRNs found in the abstract and full-text (personal communication via NLM Helpdesk Ticket CAS-552810-T3H7V5). As such, we would expect similar rates of TRNs in PubMed metadata across registries for trials with a TRN in abstract or full-text. However, even when limiting to trials with a TRN in abstract or full-text, ClinicalTrial.gov TRNs are more likely than DRKS TRNs to be included in the PubMed metadata. This suggests that NLM staff are more attentive to ClinicalTrial.gov TRNs and that DRKS TRNs more often go unnoticed in their manual labor. This is supported by our finding that TRNs were more likely to be included in PubMed metadata for ClinicalTrials.gov than DRKS trials with their TRN in the abstract or full-text. Futhermore, it appears that NLM staff may be more attentive to abstracts than to full-text, since almost all trials with ClinicalTrials.gov TRNs in the abstract, but only about half of those with ClinicalTrials.gov TRNs in the full-text, had the TRN in the metadata.

The almost three-fold discrepancy in the rates of publication references in registrations is driven by automatic indexing of trial publications based on TRNs in the PubMed metadata. ClinicalTrials.gov uses PubMed metadata to automatically link publications, whereas DRKS currently allows only for manual submission of references by trialists. When limiting to manually submitted publication references in registrations, performance across registries is comparable.

Accurate TRN formatting in publication data is critical for machine-readability, which in turn enables automated indexing of publications in the registration. While our regular expressions allow for and correct minor formatting errors (such as erroneous punctuation), egregious misformatting may make the TRN undetectable. For example, in our visual inspection of the PubMed metadata TRNs, we found `r si_drks_misformatted` severely misformatted DRKS TRNS (i.e., numbers only with no preceding letters), which we could only identify as DRKS TRNs using additional metadata and which prevented the regex classifying them as TRNs based on the pattern alone. Classifying these cases as true negative, our regexes correctly classified all TRNs and non-TRNs in our sample's the PubMed metadata secondary identifier field (sensitivity and specificity of 100%).<!--# remove since not about trn formatting alone? Futhermore, ClinicalTrials.gov presumably relies not only on the reported TRN but also on the registry, and as such failed to index a publication (PMID: 26417987) despite the TRN (NCT01614301) appearing in the PubMed metadata, as the associated databank was incorrect (i.e., GENBANK). -->

We detected X TRNs in PubMed metadata, abstracts, and full-text which did not correspond to either the associated trial's IntoValue TRN or a cross-registration listed on the registry. While the TRNs in the PubMed metadata (n = X) and abstracts (n = X) are likely supplemental cross-registrations, the additional TRNs in the full-text may also have been provided as background or discussion. Such discrimination falls beyond the scope of these regexes.

## Strengths and Limitations

This approach has numerous strengths. In contrast with previous studies relying on PubMed queries to identify potential (randomized) clinical trial results publications, we relied on a sample of bona fide results publications from registered trials, allowing us to evaluate the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.

We used an automated approach of regular expressions with high sensitivity and specificity to identify and classify publication identifiers and TRNs which allowed for a larger sample size that manual data extraction would have permitted. This study demonstrates the application of an code-based strategy which provides a scalable method for evaluating publication links in registrations and TRN reporting in publications and can be applied to other trial sets.

The approach also faces limitations. Our input IntoValue dataset was limited to German UMC-led trials found on ClinicalTrials.gov and DRKS. It is possible that this sample and its cross-linkage rate do not reflect the broader practices across other registries and/or other countries. Future projects should look at similar multi-national samples of bona fide results publication.

Furthermore, we relied on IntoValue for trial deduplication and bona fide results publications. The dataset may have had a small number of unaccounted for cross-registrations (e.g., DRKS00004156 and NCT00215683) and publications that are not trial results (e.g., DRKS00009565 associated publication is 10.1097/aap.0000000000000573, a systematic review, with many reported TRNs).

Finally, the approach relies quality input text data, and since journal articles are still most commonly distributed as PDFs, parsing errors may reduce text quality. We minimized this risk by relying on GROBID to parse full-text PDFs into XML with the body separated out from the abstract, header, and back. While GROBID algorithms are trained on academic papers and have been used in large-scale bibliometric projects [e.g., @wang2020ac], it does not perfectly classify the full-text main body. As such, TRNs appearing in the main body main be mistakenly excluded and TRNs appearing in the header or back material may be mistakenly included.

## Implications for policy and practice

<!--# The cross-linkage of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research. This responsible practice comes at negligible costs to researchers, on the order of seconds for pasting TRNs in papers and minutes for adding a publication link to the registration. Our study shows that German UMCs can improve in both TRN reporting in publications and references to publications in the registration. While TRN reporting has improved gradually over the past decade, the rate of change is insufficient and demands a systemic change in practices across the scientific ecosystem to improve this good scientific research practice. -->

This study reveals shortcomings in the cross-linkage of German UMC-led trials registrations and results publications and highlights several promising avenues forward. In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication cross-linkage is relatively straightforward and can be improved with action across stakeholders. Researchers should comply with CONSORT and ICMJE and include the TRN in their publication full-text and abstract. Research institutions like UMCs can facilitate this behavior by educating, supporting, and incentivizing their scientists to adopt this practice. For example, ethics review offices or UMC-core facilities for clinical research could provide researchers a reminder sheet with trial registration and reporting guidelines when reviewing trials. Additionally, *Leistungsorientierte Mittel (LOM)* or bonuses could be disseminated for trial registration, reporting, and reporting of registration. Publishers could check for correctly formatted TRNs for all submitted trial results publications, either via manual review or through simple software such as the regular expressions used for this study. As some already do, journals and preprint servers could also include a separate TRN metadata field during manuscript submission and could provide this data to PubMed and other bibliographic databases. The National Library of Medicine relies on publisher-provided data and manual indexers to create the TRN PubMed metadata if the TRN appears in the abstract or full-text. PubMed and other bibliographic databases could extend and semi-automate these methods in order to detect more TRNs by integrating regex-based software at least to flag potential TRNs which can then be verified by manual indexers if desired. Registries can support cross-linkage by highlighting the correct TRN formatting on their websites and by reminding researchers to include their TRN in publications and link their publication in the registration. Furthermore, they could adopt an automated linking approach using PubMed metadata, similar to that used by ClinicalTrials.gov. Figure \@ref(fig:flow-discussion) shows a process diagram of potential actions to improve registration-publication linkage from stakeholders across clinical research.

With improved TRN inclusion in bibliographic metadata and increased automatic indexing of publications based on this metadata, full cross-linkage could be achieved with negligible work by researchers: simply pasting the TRN into the publication abstract and full-text. In our sample alone, an additional `r report_pct_num_denom_trials(n_trials_link_trn_ft_or_abs_no_si, n_trials_link_trn_ft_or_abs)` of trials with TRN in abstract or full-text could have the TRN included in PubMed's metadata, and `r report_pct_num_denom_trials(n_trials_drks_trn_si_not_linked_pub, n_trials_drks)` of DRKS trials with TRN in the PubMed metadata could be automatically indexed in the registration.

```{r flow-discussion, fig.cap = "Cross-linkage of registrations and publications across the clinical research ecosystem"}
knitr::include_graphics(here("docs", "figures", "flow-discussion.png"))
```

# Acknowledgments

[TO ADD]

# Funding

This work was funded under a grant from the Federal Ministry of Education and Research of Germany (Bundesministerium für Bildung und Forschung - BMBF) [insert grant number]. The funder had no involvement in the study design, data collection, analysis, or interpretation, writing of the manuscript, or the decision to submit for publication.

# Author disclosures

MSH, BGC, and DS declare they have no direct conflict of interest related to this work. MSH is employed as a researcher under the project funding and additional grants from the German Bundesministerium für Bildung und Forschung (BMBF). DS is a member of the Sanofi Advisory Bioethics Committee and receives an honorarium for his contribution to meetings. <!--# BGC? -->

# References {.unnumbered}

::: {#refs}
:::

# (APPENDIX) Appendix {.unnumbered}

# TRN Regular Expressions

```{r tbl-trn-regexes}
ctregistries::registries %>% 
  select(
    Registry = registry,
    "Registry Title" = databank_full_name,
    "TRN Regex" = trn_regex
  ) %>% 
  flextable::flextable(cwidth = 2) %>% 
  flextable::set_caption(caption = "Trial registration number regular expressions")
```

### 2021-07-08 docs/manuscript.Rmd
Update abstract, intro, methods

@@
 ---
-title: "Linkage of trial registrations and results publications: A cross-sectional, automated evaluation of German university medical centers"
+title: "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers"
 author: "Maia Salholz-Hillel, Daniel Strech, Benjamin Gregory Carlisle"
@@
     latex_engine: xelatex
+    fig_caption: true
+    number_sections: false
 bibliography: references.bib
@@
 
-**Objective** To evaluate, using a novel automated approach, the linkage of registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
-The International Committee of Medical Journal Editors (ICMJE) and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linkage to promote discoverability of both the clinical trial publication and registration and thus increase the value of clinical research.
+**Objective** To evaluate, using a novel automated approach, the links between registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to promote discoverability of both clinical trial publications and registrations and thus increase the value of clinical research.
 
@@
 
-**Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered clinical trials conducted across German UMCs and completed between 2009 and 2017.
-We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and TRNs in publication metadata, abstract, and full-text.
+**Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017.
+We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
 <!--# We verified our automated approach on a manually coded sample.
@@
 
-**Secondary analyses** We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and publication linkage in the registration.
-We also explored the relationship between cross-linkage<!--# practices --> and registry.
-We also explored the relationship between cross-linkage<!--# practices --> and prospective registration <!--# not sure about this.
+**Secondary analyses** We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and a linked publication in the registration.
+We also explored the relationship between registration-publication links and (1) registry, and (2) prospective registration <!--# not sure about this.
 i think if we do this then glm for the various practices -->.
-We looked at change in registration-publication linkage over time.
-We also looked at cross-registrations in the registries as well as additional TRNs reported in trial publications.<!--#  We also looked at reporting PMC-version full-text.
-We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status.
+We looked at change in registration-publication linking practices over time.
+We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.<!--# We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status.
 We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting.
@@
 
-**Results** Of `r n_trials_ft_pdf` trials with PubMed-indexed clinical trial results publications available as full-text, we found `r report_n_pct_trials(n_trials_link_full)` trials with comprehensive registration-publication linkage, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata.
-In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links.
-We found that only `r report_n_pct_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, whereas `r report_n_pct_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials did so, mostly via automatic publication indexing.
-TRNs were most commonly reported in the full-text, appearing in `r report_n_pct_trials(n_trials_link_trn_ft)`.
-Linkage rates appear overall lower for DRKS trials than for ClinicalTrials.gov.
-The most common link is TRN reporting in the full-text which reached `r stringr::str_extract(report_link_year("TRN (Full-text)", 2017), "\\d{2}%")` in 2017 in both registries.
-<!--# add more secondary?
+**Results** Only `r report_pct_num_denom_trials(n_trials_link_full)` of trials comprehensively link their registrations and publications, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata.
+In contrast, `r report_n_pct_trials(n_trials_link_none)` trials had no links.
+The most common link was TRN reported in the full-text, which occurred for `r report_n_pct_trials(n_trials_link_trn_ft)` trials <!--# and reached `r stringr::str_extract(report_link_year("TRN (Full-text)", 2017), "\\d{2}%")` in 2017 in both registries.
+--><!--# add more secondary?
 -->
 
-**Conclusion** Despite encouragement from CONSORT and ICMJE, German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
-Publication linkage was noticeably greater in ClinicalTrials.gov which uses a more structured data entry process as well as automatically indexes publications from PubMed.
+DRKS trials were overall more poorly linked than ClinicalTrials.gov trials.
+For example, links to results publications appeared in `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov registrations, but only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS registrations.
+Most [`r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)`] of the links in ClinicalTrials.gov were generated via automatic publication indexing from PubMed metadata, which DRKS does not do.
+
+**Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
+ClinicalTrials.gov trials were more extensively linked and rely on a more structured data entry process and automatic publication indexing from PubMed; DRKS could adopt similar strategies to improve linking.
+
 German and international university medical centers (UMCs) should incentivize, promote, and educate their researchers on this responsible research practice.
-Automated strategies, such as those used in this study, could support bibliometric databases and registries to enrich their registration-publication linkage.
+Automated strategies, such as those used in this study, could support bibliometric databases and registries to enrich their registration-publication links.
 For example, to improve TRN inclusion in its metadata, PubMed could automatically extract TRNs from publication abstracts or full-text.
-DRKS could adopt strategies like those used by ClinicalTrials.gov to automatically index publications reporting the TRN in PubMed.
 As compared with other responsible research practices (such as trial registration itself), reporting TRN and publications are copy-and-paste tasks that may be integrated into biomedical researchers' existing workflows.
@@
 
-Clinical trial registration and results reporting are important tools for clinical research transparency [@worldmedicalassociation2013; @worldhealthorganization2017].<!--# I removed background on registration and results reporting for succinctness and because there's a robust literature, which I can add additional citations to.
-Let me know if you think it's worth adding back a few sentence of background.
---> Linkage of the trial registrations and results publications promotes discoverability of both components <!--# adds further transparency beyond registration and results reporting independently -->and increases the value of clinical research.
-Such threaded evidence <!--# Linked registrations-publications-->empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].
-Robust linkage includes both the incorporation of a structured reference to publications in the registration, and reporting of trial registration numbers (TRN) in the full-text, abstract, and metadata of publications.
+Clinical trial results risk being left out of evidence synthesis and health policy decision making if publications cannot be found.
+Incomplete evidence threatens healthcare and the lives and welfare of human volunteers.
+Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp].
+Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing the clinical research value[@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->.
+Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].
 
+Trial results should be bidirectionally linked, meaning the inclusion of both a structured reference to publications in the registration, and trial registration numbers (TRN) in the full-text, abstract, and metadata of publications.
 The Consolidated Standards of Reporting Trials (CONSORT) as well as the ICMJE's *Recommendations for the Conduct, Reporting, Editing, and Publication of Scholarly Work in Medical Journals* call for reporting "trial registration number and name of the trial register" in both the full-text and abstract of trial results publication [@internationalcommitteeofmedicaljournaleditorsicmje2019; @hopewell2008pm].
@@
 
-Furthermore, the inclusion of TRNs in publication and bibliographic metadata enhances discoverability for both human readers and computerized analysis.
-For example, PubMed MEDLINE metadata includes TRNs under the "secondary source id" tag [@nationallibraryofmedicine2018], which can then be searched on the web interface (e.g., "NCT00002568[secondary identifier]" or "NCT00002568[si]") or accessed programmatically via the underlying XML [@nationallibraryofmedicine2017c].
-Currently, these tags can be generated from publisher-provided data or assigned manually by NLM Indexing Data Reviewers who copy-and-paste secondary source ids (including trial ids) found in the abstract and full-text (personal communication via NLM Helpdesk Ticket CAS-552810-T3H7V5).
-<!--# @daniel: this had been in the intro but you said to shorten so i moved to discussion but then you had several quetion about how SI is made, so moved it back to intro...
--->
-
-Machine-readable TRN metadata moreover facilitates automatic incorporation of the publication in the registry for "easy linking of trial related publications with clinical trial registry site records" [@nationallibraryofmedicine2017c].
-For example, ClinicalTrials.gov uses PubMed secondary identifiers to automatically index publications in the registry entry under "Additional publications automatically indexed to this study by ClinicalTrials.gov Identifier (NCT Number)" [@nationallibraryofmedicine2017c].
-This therefore increases the discoverability of the publication from the registration, with no additional effort from the trial investigator for manually adding the publication to the registry.
-
+The inclusion of TRNs in publication and bibliographic metadata especially enhances machine readability and discoverability.
+Trial results can then be found more efficiently using TRNs to search publication databases [@huser2013po].
+Publications can then also be automatically linked within trial registrations, as is currently done with ClinicalTrials.gov using TRN metadata from PubMed [@nationallibraryofmedicine2017c].
 Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings.
-<!--# add policy background on why pub in reg --> The insertion and structuring of references depends on the registry; whereas ClinicalTrials.gov includes both automatically indexed publications and manual entries in quasi-structured datafields, DRKS currently accepts only manual entries in minimally structured datafields.
-References, like those automatically indexed, may include one or more publication unique identifier such as a DOI or PMID, or may be limited to a free-text citation with journal, article, and author names, but no unique identifier.
-While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval and linkage [@huser2013po].
+<!--# add policy background on why pub in reg --> Results references may be entered manually in many registries, with varying degrees of structure.
+ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas DRKS offers an unstructured free-text field.
+<!--# add some background on ctgov vs.
+drks.
+and how they include trns and publications.
+and ctgov and pubmed by same nlm so may be better at seing those ids -->While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval [@huser2013po].
 
 Trial evidence appears to be insufficiently linked in both registries and publications.
+Previous studies on trial registrations found links to results publications in as few as 13% to as many as 42% of trials [@bashir2017sr; @huser2013po].
+TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b].
+For example, some studies included trial publications limited to specific journals or fields whereas others relied on PubMed "clinical trial" publication type queries.
+
+<!--# REMOVE or @murph sanity check of numbers from other papers -->
+
 @huser2013po found that only 28% of 8907 completed trials contained a structured trial-article link with a publication unique identifier in the registry.
-In a systematic review of 16 studies examining cohorts of trial publications for links to registrations<!--# of results publications with associated registrations/investigating linked registrations-publications in publication cohorts found via both automated and manual approaches...couldn't find the overall TRN reporting rate from all studies looking at TRN, only the 16 using both automated and manual approaches, but this level of detail feels like too much to explain -->, @bashir2017sr found an overall rate of 49% of results publications <!--# (range: 8-97%) --> reported a TRN in either the metadata, abstract, or full-text.
+In a systematic review of 16 studies examining cohorts of trial publications for links to registrations<!--# of results publications with associated registrations/investigating linked registrations-publications in publication cohorts found via both automated and manual approaches...couldn't find the overall TRN reporting rate from all studies looking at TRN, only the 16 using both automated and manual approaches, but this level of detail feels like too much to explain -->, @bashir2017sr found that a median of 49% of results publications (range: 8-97%) reported a TRN in either the metadata, abstract, or full-text.
 More recently, @carlisle2020m found 38,001 ClinicalTrials.gov ids across 488,365 abstracts from records indexed by PubMed as clinical trial publications, meaning a maximum of `r round(38001/488365 * 100, digits = 1)`% of publications report a ClinicalTrials.gov id, assuming one TRN per publication.
-@aldurra2020b found that only 28.8% (7473/10,500) records classified as RCTs based on a bespoke PubMed query reported trial ids in any of the metadata, abstract, or full-text.
-<!--# @murph sanity check of numbers from other papers -->
+@aldurra2020b found that only 28.8% (7473/10,500) of publications classified as RCTs based on a bespoke PubMed query reported trial ids in any of the metadata, abstract, or full-text.
 
-These previous projects, however, faced limitations due to the unclear publication cohort and unknown true registration rates.
+In order to draw robust conclusions about registration-publication linking, the analysis population must be limited to registered clinical trials with published results.
+Previous projects faced limitations due to the unclear publication cohort and unknown true registration rates.
 PubMed queries currently do not suffice to identify the true populations of clinical trials or even RCTs, both excluding true positives (i.e., sensitivity) and including false positives (i.e., specificity) [@glanville2020jmla].
 Therefore not all PubMed-indexed clinical trials or RCTs should be registered <!--# as they may not reflect true clinical trials or RCTs -->, and not all that should be are.
-Furthermore, not all reported TRNs correspond to bona fide registrations of trial whose results are reported in the publication, i.e., TRNs may be reported for related trials in the background section.
+Furthermore, not all reported TRNs correspond to bona fide registrations of the trial whose results are reported in the publication, i.e., TRNs may be reported for related trials in the background section.
 As such, conclusions drawn by these previous studies relying on PubMed-query based samples, about registration reporting rates must be qualified as the rate of reported trial registration numbers in PubMed-indexed (randomized) clinical trials with unknown registration rates.
 <!--# remove previous sentence?
---> In addition, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link.
-These studies of publication linkage in registrations have furthermore focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of other studies investigating publication linkage in DRKS.
+--> Similarly, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link.
+Furthermore, these studies of publication links in registrations have focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of other studies investigating publication links in DRKS.
 
@@
 
-We evaluated the linkage of registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017.
-<!--# @daniel comment: add reference to intovalue here?
-it's in methods -->We also looked at change in these linkage practices over time.
-By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication linkage, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.
+We evaluated linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017.
+We also looked at change in these practices over time.
+By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.
 We use an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs.
@@
 
-## Data sources and collection
+## Automated pipeline for data sources and collection
 
@@
 We <!--# programmatically -->downloaded data from both registries on 15 May 2021.
-We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data.
-We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti].
+<!--# We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data.
+-->We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti].
 As DRKS does not provide an API, we built a webscraper to capture the necessary fields [@federalinstitutefurdrugsandmedicaldevicesbfarm].
-ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information.
+<!--# ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information.
 For example, ClinicalTrials.gov has a separate summary results reporting section, whereas summary results are included with other references in DRKS.
-We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse").
-<!--# remove if not including summary results -->
+We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse").-->
 
-We then queried the PubMed Entrez Programming Utilities (E-utilities) API via an R client for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter].
+We queried the PubMed Entrez Programming Utilities (E-utilities) API for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter].
 From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata, as well as MeSH terms, keywords, author details, and summary metadata.
@@
 We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies?
---> [@ourresearch; @jahn].
-We then used the GROBID machine learning library for technical and scientific publications to parse the PDFs into machine-readable XMLs [@2008].
+-->[@ourresearch; @jahn].
+We then used the GROBID machine learning library for technical and scientific publications (v0.6.1) via Python to parse the PDFs into machine-readable XMLs [@2008; @lopez2021].
 In order to isolate the main body from the publication abstract for our analysis, we extracted the `<body>` sections of the papers.
 
-Figure \@ref(fig:flow-methods) show the methods flow using automated techniques with the exception of some full-text PDFs.
+Figure \@ref(fig:flow-methods) show the methods flow using automated techniques<!--# with the exception of some full-text PDFs -->.
 
@@
 We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field.
-For DRKS, we scraped the reference type, citation, and URL link, when available.
-We used regexes to extract any DOIs and PMIDs from either the citation or link.
+For DRKS, we scraped the reference type, citation, and URL, when available.
+We used regexes to extract any DOIs and PMIDs from the citation and URL.
 In the rare cases when conflicting DOIs or PMIDs were found in the citation vs.
-the link, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation.
+the URL, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation.
 In addition, we determined whether a reference was manually or automatically indexed in the registry.
@@
 
-We developed <!--# and evaluated [or more methods details] --> regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(trn-regular-expressions) Table \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
-To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata for true positives and true negatives.<!--# In order to isolate the full-text main body from the publication abstract for our analysis, we included only `<body>` sections of the PMC and PDF samples for the full-text source.
---> For each source, we first extracted all unique TRN patterns within a source.
+We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(trn-regular-expressions) Table \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
+To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata for true positives and true negatives.
+For each source, we first extracted all unique TRN patterns within a source.
 Our regex allowed for minor formatting errors in the TRN patterns (such as additional punctuation).
@@
 <!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration.
-currently placeholder "TRUE" such that all TRNs resolve in the registries -->
-
-<!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns?
+currently placeholder "TRUE" such that all TRNs resolve in the registries --><!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns?
 then would have to change my backend work.
@@
 Data collection, preparation, and analysis was performed in R version 4.0.2 (R Foundation for Statistical Computing) and RStudio version 1.4 (RStudio Team).
-PDF conversion to TEI XML was performed using GROBID Server 0.6.1 [@2008] and GROBID Python Client [@lopez2021].
+Additional packages and libraries are specified in the methods.
 All code and the final analysis data for this study are available at [insert link].
@@
 
-The IntoValue trials dataset included the full sample of trials led by a German UMC that were registered on ClinicalTrials.gov or the DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
-After applying our exclusion criteria <!--# excluding non-German UMC-led trials, deduplicating trials, excluding trials without results, enhancing the publication metadata, getting full-text, and excluding trials with results without a PMID and full-text -->, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
+The IntoValue trials dataset included the full sample of trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
+After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
 Figure \@ref(fig:flow-trials) details the trial and publication screening.
@@
 
-## Registration-publication linkage
+## Registration-publication links
 
-While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linkage, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and metadata.
+While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and metadata.
 In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication.
-Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the breakdown of registration-publication linkage across the various components for ClinicalTrial.gov and DRKS trials.
+Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the breakdown of registration-publication links across the various components for ClinicalTrial.gov and DRKS trials.
 
-```{r plot-upset-links-reg-pub-registries, fig.cap = "Registristion-publication linkages in ClinicalTrials.gov and DRKS", out.width="80%"}
+```{r plot-upset-links-reg-pub-registries, fig.cap = "Registration-publication links in ClinicalTrials.gov and DRKS", out.width="80%"}
 knitr::include_graphics(here("docs", "figures", "plot-upset-links-reg-pub-registry.png"))
@@
 
-We found that `r report_pct_num_denom_trials(n_trials_link_reg)` registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results.
+We found that `r report_pct_num_denom_trials(n_trials_link_reg)` of registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results.
 Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed.
@@
 
-### Linkage by registry
+### Linking by registry
 
@@
 We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`).
-Table \@ref(tab:tbl-link-by-reg) shows summary registration-publication linkage by registry.
+Table \@ref(tab:tbl-link-by-reg) shows summary registration-publication links by registry.
 
@@
   gtsummary::as_flex_table() %>% 
-  flextable::set_caption(caption = "Registration-publication linkage by registry")
+  flextable::set_caption(caption = "Registration-publication links by registry")
 ```
@@
 
-Linkage rates appear overall lower for DRKS trials than for ClinicalTrials.gov.
+Overall, DRKS trials appear to be less robustly linked than ClinicalTrials.gov trials.
 Overall, we observed positive trends for TRN reporting in full-text and abstracts.
 <!--# add test?
---> Figure \@ref(fig:plot-links-year-reg) shows the registration-publication linkages over time for ClinicalTrials.gov and DRKS.
+--> Figure \@ref(fig:plot-links-year-reg) shows the registration-publication links over time for ClinicalTrials.gov and DRKS.
 
-```{r plot-links-year-reg, fig.cap = "Change over time in registration-publication linkages in ClinicalTrials.gov and DRKS. Completion year from registry."}
+```{r plot-links-year-reg, fig.cap = "Change over time in registration-publication links in ClinicalTrials.gov and DRKS. Completion year from registry."}
 knitr::include_graphics(here("docs", "figures", "plot-links-year-reg.png"))
@@
 
-We found `r report_pct_num_denom_trials(n_trials_crossreg_reg)` of trials indicated a total of `r n_crossreg_reg` cross-registrations as secondary identifiers in their registrations.
+We found `r report_pct_num_denom_trials(n_trials_crossreg_reg)` of trials listed additional TRNs in their registrations, indicating likely cross-registrations.
+Of these, 18 trials (6%) listed more than 1 cross-registration, for a total of `r n_crossreg_reg` cross-registrations.
 Of these `r n_crossreg_reg` cross-registrations, X were reported in pubs (X in metadata, X in abs, X in ft).
@@
 
-The cross-linkage of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research.
+Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research.
 This responsible practice comes at minimal costs to researchers, on the order of seconds for pasting TRNs in papers and minutes for adding a publication link to the registration.
@@
 Our study shows that German UMCs can improve in both TRN reporting in publications and references to publications in the registration.
-In our sample (n = `r n_trials_ft_pdf`), `r report_n_pct_trials(n_trials_link_none)` trials had no links between registration and publication and only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linkage
+In our sample (n = `r n_trials_ft_pdf`), `r report_n_pct_trials(n_trials_link_none)` trials had no links between registration and publication and only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication links.
 
@@
 
-Trials registered in ClinicalTrials.gov showed better linkage than trials registered in DRKS<!--# or be explicit that not significant for full-text -->.
+Trials registered in ClinicalTrials.gov showed better linking than trials registered in DRKS<!--# or be explicit that not significant for full-text -->.
 Notably, TRNs in PubMed metadata and publication references in registrations were almost three-fold more frequent in ClinicalTrials.gov than in DRKS.
@@
 This suggests that NLM staff are more attentive to ClinicalTrial.gov TRNs and that DRKS TRNs more often go unnoticed in their manual labor.
-This is supported by our finding that TRNs were more likely to be included in PubMed metadata for ClinicalTrials.gov than DRKS trials with their TRN in the abstract or full-text.
-Futhermore, it appears that NLM staff may be more attentive to abstracts than to full-text, since almost all trials with ClinicalTrials.gov TRNs in the abstract, but only about half of those with ClinicalTrials.gov TRNs in the full-text, had the TRN in the metadata.
+<!--# This is supported by our finding that TRNs were more likely to be included in PubMed metadata for ClinicalTrials.gov than DRKS trials with their TRN in the abstract or full-text.
+--> Furthermore, it appears that NLM staff may be more attentive to abstracts than to full-text, since almost all trials with ClinicalTrials.gov TRNs in the abstract, but only about half of those with ClinicalTrials.gov TRNs in the full-text, had the TRN in the metadata.
 
@@
 
-We detected X TRNs in PubMed metadata, abstracts, and full-text which did not correspond to either the associated trial's IntoValue TRN or a cross-registration listed on the registry.
+We detected X TRNs in PubMed metadata, abstracts, and full-text <!--# Some trial publications also reported TRNs --> which did not correspond to either the associated trial's IntoValue TRN or a cross-registration listed on the registry.
 While the TRNs in the PubMed metadata (n = X) and abstracts (n = X) are likely supplemental cross-registrations, the additional TRNs in the full-text may also have been provided as background or discussion.
@@
 We used an automated approach of regular expressions with high sensitivity and specificity to identify and classify publication identifiers and TRNs which allowed for a larger sample size that manual data extraction would have permitted.
-This study demonstrates the application of an code-based strategy which provides a scalable method for evaluating publication links in registrations and TRN reporting in publications and can be applied to other trial sets.
+This study demonstrates the application of an automated strategy which provides a scalable method for evaluating publication links in registrations and TRN reporting in publications and can be applied to other trial sets.
 
@@
 Our input IntoValue dataset was limited to German UMC-led trials found on ClinicalTrials.gov and DRKS.
-It is possible that this sample and its cross-linkage rate do not reflect the broader practices across other registries and/or other countries.
+It is possible that this sample and its linking rate do not reflect the broader practices across other registries and/or other countries.
 Future projects should look at similar multi-national samples of bona fide results publication.
@@
 
-<!--# The cross-linkage of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research.
+<!--# Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research.
 This responsible practice comes at negligible costs to researchers, on the order of seconds for pasting TRNs in papers and minutes for adding a publication link to the registration.
@@
 
-This study reveals shortcomings in the cross-linkage of German UMC-led trials registrations and results publications and highlights several promising avenues forward.
-In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication cross-linkage is relatively straightforward and can be improved with action across stakeholders.
+This study reveals shortcomings in the linking of German UMC-led trials registrations and results publications and highlights several promising avenues forward.
+In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is relatively straightforward and can be improved with action across stakeholders.
 Researchers should comply with CONSORT and ICMJE and include the TRN in their publication full-text and abstract.
@@
 Publishers could check for correctly formatted TRNs for all submitted trial results publications, either via manual review or through simple software such as the regular expressions used for this study.
-As some already do, journals and preprint servers could also include a separate TRN metadata field during manuscript submission and could provide this data to PubMed and other bibliographic databases.
+As some already do (see for example Taylor & Francis [@taylorfrancisgroup2020]), publishers, journals, and preprint servers could also extract TRNs from the paper and/orinclude a separate TRN metadata field during manuscript submission, display this on the publication webpage, and provide this data to PubMed, CrossRef, and other bibliographic databases [@crossref2020c].
 The National Library of Medicine relies on publisher-provided data and manual indexers to create the TRN PubMed metadata if the TRN appears in the abstract or full-text.
 PubMed and other bibliographic databases could extend and semi-automate these methods in order to detect more TRNs by integrating regex-based software at least to flag potential TRNs which can then be verified by manual indexers if desired.
-Registries can support cross-linkage by highlighting the correct TRN formatting on their websites and by reminding researchers to include their TRN in publications and link their publication in the registration.
+Registries can support linking by highlighting the correct TRN formatting on their websites and by reminding researchers to include their TRN in publications and link their publication in the registration.
 Furthermore, they could adopt an automated linking approach using PubMed metadata, similar to that used by ClinicalTrials.gov.
-Figure \@ref(fig:flow-discussion) shows a process diagram of potential actions to improve registration-publication linkage from stakeholders across clinical research.
+Figure \@ref(fig:flow-discussion) shows a process diagram of potential actions to improve registration-publication links from stakeholders across clinical research.
 
-With improved TRN inclusion in bibliographic metadata and increased automatic indexing of publications based on this metadata, full cross-linkage could be achieved with negligible work by researchers: simply pasting the TRN into the publication abstract and full-text.
+With improved TRN inclusion in bibliographic metadata and increased automatic indexing of publications based on this metadata, full linking could be achieved with negligible work by researchers: simply pasting the TRN into the publication abstract and full-text.
 In our sample alone, an additional `r report_pct_num_denom_trials(n_trials_link_trn_ft_or_abs_no_si, n_trials_link_trn_ft_or_abs)` of trials with TRN in abstract or full-text could have the TRN included in PubMed's metadata, and `r report_pct_num_denom_trials(n_trials_drks_trn_si_not_linked_pub, n_trials_drks)` of DRKS trials with TRN in the PubMed metadata could be automatically indexed in the registration.

### 2021-07-22 docs/manuscript.Rmd
Add logistic regression to manuscript

@@
 source(here("R", "analysis", "cross-registrations.R"))
+source(here("R", "analysis", "logistic-regression.R"))
 # # fs::dir_ls(here("R", "analysis")) %>% 
@@
 
-**Secondary analyses** We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and a linked publication in the registration.
+**Secondary analyses** We constructed logistic regression models to explore the relationship between trial completion date, registry, and registration-publication linking.
+We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.
+
+We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and a linked publication in the registration.
 We also explored the relationship between registration-publication links and (1) registry, and (2) prospective registration <!--# not sure about this.
@@
 We looked at change in registration-publication linking practices over time.
-We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.<!--# We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status.
+<!--# We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status.
 We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting.
@@
 
-## Automated pipeline for data sources and collection
+## Automated pipeline for data collection
 
@@
 
-### Exclusion/inclusion criteria
+### Inclusion and exclusion criteria
 
@@
 <!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration.
-currently placeholder "TRUE" such that all TRNs resolve in the registries --><!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns?
+currently placeholder "TRUE" such that all TRNs resolve in the registries -->
+
+<!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns?
 then would have to change my backend work.
@@
 -->
+
+## Analysis
+
+We generated descriptive statistics on trial and publication characteristics, overall and by registry.
+We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.
+<!--#with each type of link and different combination of links.-->
+
+To explore change over time, the relationships between types of links, and differences between registries, we ran exploratory logistic regressions for each link type, with all other link types as well as registry and year as explanatory variables.
+In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata.
+
+Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs.
+manually linked publications in ClinicalTrials.gov.
+To explore registry differences for only manually linked references, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration.
+<!--#add automatic links to regression for pub in registry?
+in which case would need to recode NA as FALSE.
+To explore whether automated links explains registry differences in publication links in the registry...-->
 
@@
 
-The IntoValue trials dataset included the full sample of trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
+The IntoValue trials dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
 After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
 Figure \@ref(fig:flow-trials) details the trial and publication screening.
-Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications.
+Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.
 <!--# QUESTION: what summary info about trials and publications?
@@
 
-While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and metadata.
+While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata.
+Disregarding metadata, which is beyond a trialist's control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_n_pct_trials(n_trials_link_ft_abs_reg)`.
 In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication.
-Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the breakdown of registration-publication links across the various components for ClinicalTrial.gov and DRKS trials.
+The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_n_pct_trials(n_trials_link_trn_ft_only)` trials.
+The inclusion of the various link types ranged from `r report_pct_num_denom_trials(n_trials_link_trn_abs)` in abstracts to `r report_pct_num_denom_trials(n_trials_link_trn_ft)` in full-text.
+Table \@ref(tab:tbl-link-by-reg) shows registration-publication links overall and by registry.
+<!--# point estimates in table should have CI, right?
+-->Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the intersection of registration-publication links across the various types for ClinicalTrial.gov and DRKS trials.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.-->
+
+```{r tbl-link-by-reg}
+tbl_link_by_reg %>% 
+  gtsummary::as_flex_table() %>% 
+  flextable::set_caption(caption = "Registration-publication links by registry")
+```
 
@@
 
+### Logistic regression
+
+```{r tbl-trials-regression}
+tbl_trials_regression %>% 
+  gtsummary::as_flex_table() %>% 
+  flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
+```
+
+Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables.
+Completion year did not have a strong relationship with linking practices, although trials completing more recently were significantly more likely report the TRN in the abstract (`r inline_text(cor_abs, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(cor_ft, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
+Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.
+
+```{r plot-links-year-reg, fig.cap = "Change over time in registration-publication links in ClinicalTrials.gov and DRKS. Completion year from registry."}
+knitr::include_graphics(here("docs", "figures", "plot-links-year-reg.png"))
+```
+
+With regards to the relationship between link types,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link.
+<!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were significantly more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not significantly more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`).
+Futhermore, DRKS TRNs were significantly less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
+Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (62% vs.
+18%).
+Table X shows the number and proportion of trials with a TRN in the metadata given TRN in the abstract, full-text, or either, overall and by registry.
+
+**Table X.
+Number and Proportion of trials with TRN in metadata given TRN in abstract, full-text, or either.
+Denominator indicates number of trials with TRN in abstract or full-text, from the respective registry.**
+
+|                                         | Overall        | ClinicalTrials.gov | DRKS         |
+|-----------------------------------------|----------------|--------------------|--------------|
+| **TRN in either abstract or full-text** | 53% (723/1372) | 62% (669/1077)     | 18% (54/295) |
+| **TRN in abstract**                     | 85% (605/714)  | 97% (562/580)      | 32% (43/134) |
+| **TRN in full-text**                    | 45% (513/1131) | 54% (471/878)      | 17% (42/253) |
+
+DRKS trials were also significantly less likely than ClinicalTrials.gov trials to reference the publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
+<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)`) ClinicalTrials.gov trials referencing the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
+Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (24%, 186/783 vs.
+21%, 96/447).
+
 ### Publications linked in registrations
@@
 We found that `r report_pct_num_denom_trials(n_trials_link_reg)` of registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results.
-Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed.
+Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed from PubMed metadata.
 
@@
 We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`).
-Table \@ref(tab:tbl-link-by-reg) shows summary registration-publication links by registry.
-
-<!--# point estimtes in table should have CI, right? -->
-
-```{r tbl-link-by-reg}
-tbl_link_by_reg %>% 
-  gtsummary::as_flex_table() %>% 
-  flextable::set_caption(caption = "Registration-publication links by registry")
-```
 
@@
 Trials with a ClinicalTrials.gov TRN (Proportion with TRN in SI = 95.5%) were more likely than trials with a DRKS TRN (Proportion with TRN in SI = 33.7%) to have a TRN in PubMed metadata if reported in both the abstract and the full-text (Difference in reporting = 61.8%, CI = 52.5, 71.4).
-
-### TRN reporting given prospective registration
-
-<!--# Are TRNs reported more or less often in abstract OR full text for prospective or retrospective registrations? -->
-
-A chi-square test of independence was performed to examine the relation between prospective registration and TRN reporting in either the abstract or the full-text (abstract-only, full-text-only, both, neither).
-Registration was categorized as prospective if the trial was registered in the same or preceding months to trial start, since trial start on ClinicalTrials.gov is sometimes available with month and year only.
-Two trials (NCT02127736, NCT02213406) without start dates and were excluded from this analysis as we could not determine whether registration was prospective.
-The relation between these variables was significant, X2 (1, N = 1894) = 22.8, p = .00000182.
-Prospectively registered trials (Proportion reporting = X) were more likely than retrospectively registered trials (Proportion reporting = X) to report the TRN in either the abstract and/or the full-text (Difference in reporting = X, CI = X, X).
-
-### Change in registration-publication linkage over time, by registry
-
-Overall, DRKS trials appear to be less robustly linked than ClinicalTrials.gov trials.
-Overall, we observed positive trends for TRN reporting in full-text and abstracts.
-<!--# add test?
---> Figure \@ref(fig:plot-links-year-reg) shows the registration-publication links over time for ClinicalTrials.gov and DRKS.
-
-```{r plot-links-year-reg, fig.cap = "Change over time in registration-publication links in ClinicalTrials.gov and DRKS. Completion year from registry."}
-knitr::include_graphics(here("docs", "figures", "plot-links-year-reg.png"))
-```
 
@@
 ```
+
+# Sensitivity Analyses

### 2021-07-22 R/analysis/04_link-reg-pub.R

@@ -36,6 +36,15 @@ n_trials_link_none <-
   ) %>%
   nrow()
 
+n_trials_link_ft_abs_reg <-
+  trials %>%
+  filter(has_reg_pub_link &
+           # has_iv_trn_secondary_id &
+           has_iv_trn_abstract &
+           has_iv_trn_ft_pdf
+  ) %>%
+  nrow()
+
 
 # Publication reference in registration -----------------------------------
 
@@ -101,6 +110,11 @@ n_trials_link_trn_ft <-
 n_trials_link_trn_ft_abs <-
   trials %>%
   filter(has_iv_trn_ft_pdf & has_iv_trn_abstract) %>%
+  nrow()
+
+n_trials_link_trn_ft_only <-
+  trials %>%
+  filter(has_iv_trn_ft_pdf & !has_iv_trn_abstract & !has_iv_trn_secondary_id & !has_reg_pub_link) %>%
   nrow()
 
 # How many trials with trn in ft or abs do not have trn in si?
@@ -269,9 +283,13 @@ tbl_link_by_reg <-
       has_iv_trn_ft_pdf ~ "TRN (Full-text)"
     )
   ) %>%
+
+  #TODO: figure out what test/stats to include
+  add_difference() %>%
+  # add_p() %>%
+  add_overall() %>% #show_header_names()
   # Move stats legend to each line
   # add_stat_label(location = NULL) %>%
-  add_p() %>%
   modify_header(label = "**Registration-Publication Linkage**") %>%
   # modify_caption("**German UMC-led trials with published result** (N = {N})") %>%
   bold_labels()

### 2021-07-26 docs/manuscript.Rmd
Update manuscript with additional analyses and papaja

@@
 Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp].
-Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing the clinical research value[@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->.
+Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing the clinical research value[@kubben2019]<!--# looked for pub on FAIR/findable clinical trials and not finding much beyond just data-->.
 Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].
@@
 
-## Trial and publication characteristics
-
 The IntoValue trials dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
@@
 ```
-
-## Registration-publication links
 
@@
 
-### Logistic regression
+Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables.
 
@@
 
-Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables.
 Completion year did not have a strong relationship with linking practices, although trials completing more recently were significantly more likely report the TRN in the abstract (`r inline_text(cor_abs, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(cor_ft, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
@@
 
-With regards to the relationship between link types,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link.
+Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link.
 <!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were significantly more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not significantly more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`).
@@
 18%).
-Table X shows the number and proportion of trials with a TRN in the metadata given TRN in the abstract, full-text, or either, overall and by registry.
+Table X shows the number and proportion of trials with a TRN in the metadata given TRN in the abstract, full-text, or either, both overall and by registry.
 
@@
 21%, 96/447).
-
-### Publications linked in registrations
-
-We found that `r report_pct_num_denom_trials(n_trials_link_reg)` of registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results.
-Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed from PubMed metadata.
-
-<!--# @daniel: you had several comments about automatically linked pubs so I don't think the language I had was clear.
-I tried several wordings below.
-Let's discuss and figure out what is clearest.
--->
-
-Of ClinicalTrials.gov trials with the TRN in the PubMed metadata (n = `r n_trials_ctgov_link_reg_trn_si`), `r report_n_pct_trials(n_trials_ctgov_link_reg_trn_si_manual + n_trials_ctgov_no_link_reg_trn_si, n_trials_ctgov_link_reg_trn_si)` were not automatically indexed.
-
-Of ClinicalTrials.gov trials with the TRN in the PubMed metadata (n = `r n_trials_ctgov_link_reg_trn_si`), `r report_n_pct_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_ctgov_link_reg_trn_si)` have the publication automatically linked in the registry, `r report_n_pct_trials(n_trials_ctgov_link_reg_trn_si_manual, n_trials_ctgov_link_reg_trn_si)` have the publication manually linked in the registry, and only `r report_n_pct_trials(n_trials_ctgov_no_link_reg_trn_si, n_trials_ctgov_link_reg_trn_si)` trial is not linked in the registry.
-
-All `r n_trials_ctgov_link_reg_trn_si_derived` trials with automatically indexed publications in ClinicalTrials.gov include the TRN in the PubMed metadata.
-An additional `r n_trials_ctgov_link_reg_trn_si_manual` trials include the TRN in the PubMed metadata and a link publication that was not automatically indexed, as it may have been entered by the trialist manually prior to being indexed.
-`r if (n_trials_ctgov_no_link_reg_trn_si == 1) "Only one"` ClinicalTrials.gov trial (NCT01614301) has a TRN in the PubMed metadata (PMID: 26417987) and no linked publication in the registration.
-While the TRN is indeed indexed in the secondary identifier field, PubMed mistakenly associated it with an incorrect databank (i.e., GENBANK) which presumably prevented ClinicalTrials.gov from making the link.
-
-When limiting to only manually linked publications in the registration, performance is similar between ClinicalTrials.gov (24%, 186/783) and DRKS (22%, 96/447). <!--# change to code -->
-
-### Registrations linked in publications
-
-We found that `r report_pct_num_denom_trials(n_trials_link_trn_none)` trials did not report the known TRN in any of the metadata, abstract, or full-text of the corresponding publication.
-Only `r report_pct_num_denom_trials(n_trials_link_trn_all)` reported the TRN across all three locations.
-TRNs were most commonly reported in the full-text, appearing in `r report_n_pct_trials(n_trials_link_trn_ft)`.
-TRNs were included in `r report_n_pct_trials(n_trials_link_trn_abs)` of abstracts, as compared with `r report_n_pct_trials(n_trials_link_trn_si)` of PubMed metadata.
-
-### Linking by registry
-
-We found that trials with a ClinicalTrials.gov TRN were significantly more likely than trials with a DRKS TRN to include a publication reference in the registry (`r apa::apa(chisq_link_by_reg_reg, print_n = TRUE)`) and to include the TRN in the PubMed metadata (`r apa::apa(chisq_link_by_reg_si, print_n = TRUE)`) and abstract (`r apa::apa(chisq_link_by_reg_abs, print_n = TRUE)`).
-We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`).
-
-### TRN reporting in PubMed metadata
-
-<!--# @daniel asked me to test relationship between trn in metadata and abs/ft.
-many options.
-total in abs 714 and ft 1131.
-chisquare trn reporting abs given diff combos of abs/ft reporting OR just abs or ft?
-ideally, compare registry, trn in abs/ft (not sure what levels), and trn in si, but 3 variable so not chisquare-->
-
-For the 1,372 trials that provided a TRN in either the abstract or full-text, a TRN was also available in the metadata in 55% (n = 767).
-ClinicalTrials.gov TRNs (65%, 705/1077) were more likely than DRKS TRNs (21%, 62/295) to be included in PubMed metadata.
-Furthermore, 97% (563/580) of trials with a ClinicalTrials.gov TRN in the abstract also had the TRN in the PubMed metadata, compared with 23% (31/134) of trials with a DRKS TRN.
-Similarly, 58% (507/878) of trials with a ClinicalTrials.gov TRN in the full-text also had the TRN in the PubMed metadata, compared with 18% (48/253) of trials with a DRKS TRN.
-
-A chi-square test of independence was performed to examine the relation between registry and TRN in PubMed metadata, for trials with TRN reported in both the abstract and the full-text.
-The relation between these variables was significant, X2 (1, N = 473) = 201.0, p = 1.07e-45.
-Trials with a ClinicalTrials.gov TRN (Proportion with TRN in SI = 95.5%) were more likely than trials with a DRKS TRN (Proportion with TRN in SI = 33.7%) to have a TRN in PubMed metadata if reported in both the abstract and the full-text (Difference in reporting = 61.8%, CI = 52.5, 71.4).
-
-## Cross-registrations
-
-We found `r report_pct_num_denom_trials(n_trials_crossreg_reg)` of trials listed additional TRNs in their registrations, indicating likely cross-registrations.
-Of these, 18 trials (6%) listed more than 1 cross-registration, for a total of `r n_crossreg_reg` cross-registrations.
-Of these `r n_crossreg_reg` cross-registrations, X were reported in pubs (X in metadata, X in abs, X in ft).
-Beyond the IntoValue primary TRNs and these registry-indicated cross-registrations, we detected X additional TRNs in X publications.
-Table \@ref(tab:tbl-crossreg) shows details on cross-registrations and additional TRNs by registry.
-
-```{r tbl-crossreg}
-tbl_crossreg %>% 
-  gtsummary::as_flex_table() %>% 
-  flextable::set_caption(caption = "Cross-registrations and additional TRNs by registry")
-```
 
@@
 
-We detected X TRNs in PubMed metadata, abstracts, and full-text <!--# Some trial publications also reported TRNs --> which did not correspond to either the associated trial's IntoValue TRN or a cross-registration listed on the registry.
-While the TRNs in the PubMed metadata (n = X) and abstracts (n = X) are likely supplemental cross-registrations, the additional TRNs in the full-text may also have been provided as background or discussion.
-Such discrimination falls beyond the scope of these regexes.
-
 ## Strengths and Limitations
@@
 
-[TO ADD]
+We would like to thank Kerstin Rubarth from the Charité Institut für Biometrie und Klinische Epidemiologie for her consultation on the statistical analysis.
 

### 2021-07-26 docs/manuscript_papaja.Rmd
Update manuscript with additional analyses and papaja

--- 2021-07-26 docs/manuscript.Rmd
@@
 ---
-title: "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers"
-author: "Maia Salholz-Hillel, Daniel Strech, Benjamin Gregory Carlisle"
-output: 
-  bookdown::word_document2:
-    reference_docx: quest-word-template.docx
-    latex_engine: xelatex
-    fig_caption: true
-    number_sections: false
-bibliography: references.bib
-editor_options:
-    markdown:
-        canonical: true
-    chunk_output_type: console
+title             : "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers"
+shorttitle        : "Registration-publication links in German clinical trials"
+
+author: 
+  - name          : "Maia Salholz-Hillel\\orcidlink{0000-0003-1934-9504}"
+    affiliation   : "1"
+    corresponding : yes
+    address       : "Anna-Louisa-Karsch-Str. 2, 10178, Berlin, Germany"
+    email         : "maia.salholz-hillel@charite.de"
+    role:         # CRediT roles (https://casrai.org/credit/)
+      - Conceptualization
+      - Methodology
+      - Software
+      - Investigation
+      - Data Curation
+      - Formal Analysis
+      - Visualization
+      - Validation
+      - Project Administration
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
+  - name          : "Daniel Strech\\orcidlink{0000-0002-9153-079X}"
+    affiliation   : "1"
+    role:
+      - Funding Acquisition
+      - Resources
+      - Conceptualization
+      - Writing - Review & Editing
+      - Supervision
+  - name          : "Benjamin Gregory Carlisle\\orcidlink{0000-0001-8975-0649}"
+    affiliation   : "1"
+    role:
+      - Writing - Review & Editing
+      - Methodology
+      - Software
+      - Validation
+      - Supervision
+
+affiliation:
+  - id            : "1"
+    institution   : "QUEST Center for Responsible Research, Berlin Institute of Health (BIH) at Charité Universitätsmedizin Berlin"
+
+authornote: >
+
+abstract: |
+  **Objective** To evaluate, using a novel automated approach, the links between registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make both clinical trial publications and registrations findable and thus increase the value of clinical research.
+
+  **Design** Cross-sectional bibliographic study
+
+  **Setting** German UMC clinical trials
+
+  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017.
+We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
+<!--# We verified our automated approach on a manually coded sample.
+-->
+
+  **Main outcome measures** The proportion of trial registrations incorporating a structured reference (DOI or PMID) to a known results publication.
+The proportion of clinical trial results publications that reported the known TRN in the metadata, abstract, and full-text.
+
+  **Secondary analyses** We constructed exploratory logistic regression models to investigate the relationship between trial completion date, registry, and registration-publication linking.
+<!--# We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.
+-->
+
+  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials comprehensively link their registrations and publications, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata.
+In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links.
+The most common link was TRN reported in the full-text, which occurred for `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_trn_ft")` trials.
+ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists.
+Of trials reporting a TRN in the abstract, ClinicalTrials.gov trials were more likely than DRKS trials to have the TRN captured in the PubMed metadata.
+Additionally, most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_ctgov_link_reg_trn_si_derived")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
+
+  **Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
+In addition to improved linking by trialists, changes in the PubMed TRN capturing process (such as the incorporation of automated strategies like those used in this study) and automated indexing of publications in DRKS would make trial registrations and results more findable.
+
+  **Open Data and Code** All code and the final analysis data for this study are available at https://github.com/maia-sh/reg-pub-link.
+Raw data (with the exception of the full-text of publications) is available at https://github.com/maia-sh/intovalue-data.<!--#TODO:move data processing into intovalue-data repo--><!--#**Registration** This study was not preregistered.
+The study protocol is available at [insert link].-->
+
+keywords          : "clinical trials, registration, reporting, meta-research"
+wordcount         : "3714"
+
+bibliography      : ["references.bib", "r-references.bib"]
+# appendix          : "appendix.Rmd"
+
+numbersections    : yes
+floatsintext      : no
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+linenumbers       : yes
+mask              : no
+draft             : no
+
+lang              : "en-US"
+documentclass     : "apa6"
+classoption       : "man"
+output            : papaja::apa6_word
+
+header-includes   :
+  - \usepackage{orcidlink}
 ---
@@
 ```{r setup, include = FALSE}
+
 knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
@@
 library(gtsummary)
-# library(papaja)
+library(gt)
+library(papaja)
 
@@
 trials_all <- read_rds(here("data", "processed", "trials.rds"))
-# cross_registrations_all <- read_rds(here("data", "processed", "cross-registrations.rds"))
-pubmed_all <-
-  read_rds(here("data", "processed", "pubmed-main.rds"))
+pubmed_all <- read_rds(here("data", "processed", "pubmed-main.rds"))
+
+# Prepare references
+r_refs("r-references.bib")
+r_citations <- 
+  cite_r(file = "r-references.bib", footnote = TRUE) # pkgs, withhold #r_citations$r and r_citations$pkgs
 
@@
 source(here("R", "analysis", "logistic-regression.R"))
+source(here("R", "analysis", "si-if-abs-ft.R"))
+source(here("R", "analysis", "reg-auto-link.R"))
+source(here("R", "analysis", "tbl-stakeholders.R"))
 # # fs::dir_ls(here("R", "analysis")) %>% 
@@
 
-# Abstract
+```{r abstract-vars, cache = TRUE}
+abs_n_trials_link_full <- report_pct_num_denom_trials(n_trials_link_full)
+abs_n_trials_link_none <- report_n_pct_trials(n_trials_link_none)
+abs_n_trials_link_trn_ft <- report_n_pct_trials(n_trials_link_trn_ft)
 
-**Objective** To evaluate, using a novel automated approach, the links between registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
-The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to promote discoverability of both clinical trial publications and registrations and thus increase the value of clinical research.
-
-**Design** Cross-sectional bibliographic study
-
-**Setting** German UMC clinical trials
-
-**Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017.
-We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
-<!--# We verified our automated approach on a manually coded sample.
--->
-
-**Main outcome measures** The proportion of trial registrations incorporating a structured reference (DOI or PMID) to a known results publication.
-The proportion of clinical trial results publications that reported the known TRN in the metadata, abstract, and full-text.
-
-**Secondary analyses** We constructed logistic regression models to explore the relationship between trial completion date, registry, and registration-publication linking.
-We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.
-
-We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and a linked publication in the registration.
-We also explored the relationship between registration-publication links and (1) registry, and (2) prospective registration <!--# not sure about this.
-i think if we do this then glm for the various practices -->.
-We looked at change in registration-publication linking practices over time.
-<!--# We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status.
-We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting.
-All analyses are exploratory.
--->
-
-**Results** Only `r report_pct_num_denom_trials(n_trials_link_full)` of trials comprehensively link their registrations and publications, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata.
-In contrast, `r report_n_pct_trials(n_trials_link_none)` trials had no links.
-The most common link was TRN reported in the full-text, which occurred for `r report_n_pct_trials(n_trials_link_trn_ft)` trials <!--# and reached `r stringr::str_extract(report_link_year("TRN (Full-text)", 2017), "\\d{2}%")` in 2017 in both registries.
---><!--# add more secondary?
--->
-
-DRKS trials were overall more poorly linked than ClinicalTrials.gov trials.
-For example, links to results publications appeared in `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov registrations, but only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS registrations.
-Most [`r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)`] of the links in ClinicalTrials.gov were generated via automatic publication indexing from PubMed metadata, which DRKS does not do.
-
-**Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
-ClinicalTrials.gov trials were more extensively linked and rely on a more structured data entry process and automatic publication indexing from PubMed; DRKS could adopt similar strategies to improve linking.
-
-German and international university medical centers (UMCs) should incentivize, promote, and educate their researchers on this responsible research practice.
-Automated strategies, such as those used in this study, could support bibliometric databases and registries to enrich their registration-publication links.
-For example, to improve TRN inclusion in its metadata, PubMed could automatically extract TRNs from publication abstracts or full-text.
-As compared with other responsible research practices (such as trial registration itself), reporting TRN and publications are copy-and-paste tasks that may be integrated into biomedical researchers' existing workflows.
-
-**Registration** This study was not preregistered. <!--# The study protocol is available at [insert link]. -->
-
-**Open Data and Code** All code and the final analysis data for this study are available at [insert link]. Raw data (with the exception of the full-text of publications) is available at [insert link].
+abs_n_trials_ctgov_link_reg_trn_si_derived <-
+  inline_text(tbl_reg_pub_link_ctgov_auto_v_manual, reference_derived, column = "stat_1", pattern = "{p}%")
+```
 
@@
 Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp].
-Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing the clinical research value[@kubben2019]<!--# looked for pub on FAIR/findable clinical trials and not finding much beyond just data-->.
+Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing clinical research value [@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->.
 Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].
@@
 Reporting TRN solely in the full-text does not suffice since readers may not have access to the full-text or may base their judgment of the trial or topic on the abstract alone [@hopewell2008tl].
-Abstracts may also be published independently of a full-text publication, such as for conference abstracts, highlighting the need for TRN reporting in abstracts as well [@hopewell2008pm].
+Abstracts may also be published independently of a full-text publication, such as for conference abstracts [@hopewell2008pm].
 
@@
 Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings.
-<!--# add policy background on why pub in reg --> Results references may be entered manually in many registries, with varying degrees of structure.
+<!--# add policy background on why pub in reg --> Results references may also be entered manually in many registries, with varying degrees of structure.
 ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas DRKS offers an unstructured free-text field.
@@
 TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b].
-For example, some studies included trial publications limited to specific journals or fields whereas others relied on PubMed "clinical trial" publication type queries.
-
-<!--# REMOVE or @murph sanity check of numbers from other papers -->
-
-@huser2013po found that only 28% of 8907 completed trials contained a structured trial-article link with a publication unique identifier in the registry.
-In a systematic review of 16 studies examining cohorts of trial publications for links to registrations<!--# of results publications with associated registrations/investigating linked registrations-publications in publication cohorts found via both automated and manual approaches...couldn't find the overall TRN reporting rate from all studies looking at TRN, only the 16 using both automated and manual approaches, but this level of detail feels like too much to explain -->, @bashir2017sr found that a median of 49% of results publications (range: 8-97%) reported a TRN in either the metadata, abstract, or full-text.
-More recently, @carlisle2020m found 38,001 ClinicalTrials.gov ids across 488,365 abstracts from records indexed by PubMed as clinical trial publications, meaning a maximum of `r round(38001/488365 * 100, digits = 1)`% of publications report a ClinicalTrials.gov id, assuming one TRN per publication.
-@aldurra2020b found that only 28.8% (7473/10,500) of publications classified as RCTs based on a bespoke PubMed query reported trial ids in any of the metadata, abstract, or full-text.
+For example, some studies included trial publications limited to specific journals or fields, whereas others relied on PubMed "clinical trial" publication type queries.
 
@@
 --> Similarly, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link.
-Furthermore, these studies of publication links in registrations have focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of other studies investigating publication links in DRKS.
+Furthermore, these studies of publication links in registrations have focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of previous studies investigating publication links in DRKS.
 
@@
 
-We evaluated linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017.
-We also looked at change in these practices over time.
-By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.
-We use an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs.
-We did not pre-specify the analyses.
-
-<!--# We also explored the reporting of additional TRNs, beyond the known associated TRN, in publications, including reported cross-registrations, meaning more than one TRN in more than one registry associated with a single publication.
+The exploratory study evaluates linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017.
+We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs, which may be applied to other trial cohorts.
+<!--# We did not pre-specify the analyses.
+We also explored the reporting of additional TRNs, beyond the known associated TRN, in publications, including reported cross-registrations, meaning more than one TRN in more than one registry associated with a single publication.
 We also looked at reporting PMC-version full-text.
@@
 We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset.
-In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.
+In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches [ADD DATASET REFERENCE].
 
@@
 
-We queried the PubMed Entrez Programming Utilities (E-utilities) API for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter].
-From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata, as well as MeSH terms, keywords, author details, and summary metadata.
-<!--# remove unused pubmed components?
--->
+We queried the PubMed Entrez Programming Utilities (E-utilities) API<!--#on DAY MONTH YEEAR--> for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter].
+From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->
 
@@
 
-We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(trn-regular-expressions) Table \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
-To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata for true positives and true negatives.
+We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
+To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata<!--# for true positives and true negatives-->.
 For each source, we first extracted all unique TRN patterns within a source.
 Our regex allowed for minor formatting errors in the TRN patterns (such as additional punctuation).
-We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source introduced through the cleaning process.
+We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source uncovered through the cleaning process.
 We also coalesced across sources to deduce a list of unique TRNs by publication along with the source(s) in which each was reported.
@@
 These additional TRNs could be cross-registrations of the same trial or provided as background or discussion.
-For this study, we were interested in whether known registrations were reported in the publication, so we filtered for reported TRNs matching the known registrations.
+For this study, we were interested in whether known registrations were reported in the publication and so filtered for reported TRNs matching the known registrations.
 <!--# change this if considering cross-reg -->
-
-<!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration.
-currently placeholder "TRUE" such that all TRNs resolve in the registries -->
-
-<!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns?
-then would have to change my backend work.
-or can keep this section but maybe not report all the details in text but just appendix?
--->
 
@@
 We generated descriptive statistics on trial and publication characteristics, overall and by registry.
-We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.
-<!--#with each type of link and different combination of links.-->
-
-To explore change over time, the relationships between types of links, and differences between registries, we ran exploratory logistic regressions for each link type, with all other link types as well as registry and year as explanatory variables.
+We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran exploratory logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables.
 In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata.
-
 Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs.
@@
 
-Data collection, preparation, and analysis was performed in R version 4.0.2 (R Foundation for Statistical Computing) and RStudio version 1.4 (RStudio Team).
-Additional packages and libraries are specified in the methods.
-All code and the final analysis data for this study are available at [insert link].
-Raw data (with the exception of the full-text of publications) is available at [insert link].
+Data collection, preparation, and analysis was performed in `r r_citations$r`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> All code and the final analysis data for this study are available at <https://github.com/maia-sh/reg-pub-link>.
+Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>.
+`r r_citations$pkgs`
 
@@
 
-The IntoValue trials dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
+The IntoValue dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
 After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
@@
 tbl_trials %>% 
-  gtsummary::as_flex_table() %>% 
-  flextable::set_caption(caption = "Characteristics of included trials with results publications")
+  modify_caption("Characteristics of included trials with results publications")
+  # gtsummary::as_flex_table() %>% 
+  # flextable::set_caption(caption = "Characteristics of included trials with results publications")
 ```
@@
 While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata.
-Disregarding metadata, which is beyond a trialist's control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_n_pct_trials(n_trials_link_ft_abs_reg)`.
+Disregarding metadata, which is largely beyond trialists' control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_n_pct_trials(n_trials_link_ft_abs_reg)`.
 In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication.
@@
 tbl_link_by_reg %>% 
-  gtsummary::as_flex_table() %>% 
-  flextable::set_caption(caption = "Registration-publication links by registry")
+  modify_caption("Registration-publication links by registry")
+  # gtsummary::as_flex_table() %>% 
+  # flextable::set_caption(caption = "Registration-publication links by registry")
 ```
@@
 
-Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables.
-
 ```{r tbl-trials-regression}
 tbl_trials_regression %>% 
-  gtsummary::as_flex_table() %>% 
-  flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
+  modify_caption("Crude and adjuusted odds ratios for factors associated with publication-registration links")
+  # gtsummary::as_flex_table() %>% 
+  # flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
 ```
 
-Completion year did not have a strong relationship with linking practices, although trials completing more recently were significantly more likely report the TRN in the abstract (`r inline_text(cor_abs, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(cor_ft, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
+Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables.
+Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(cor_abs, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(cor_ft, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
 Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.
@@
 Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link.
-<!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were significantly more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not significantly more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`).
-Futhermore, DRKS TRNs were significantly less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
-Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (62% vs.
-18%).
-Table X shows the number and proportion of trials with a TRN in the metadata given TRN in the abstract, full-text, or either, both overall and by registry.
+<!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`).
+Futhermore, DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
+Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (`r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_1", pattern = "{p}%")` vs.
+`r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_2", pattern = "{p}%")`).
+Table \@ref(tab:tbl-si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
 
-**Table X.
-Number and Proportion of trials with TRN in metadata given TRN in abstract, full-text, or either.
-Denominator indicates number of trials with TRN in abstract or full-text, from the respective registry.**
+```{r tbl-si-if-abs-ft}
+tbl_si_if_abs_ft
+```
 
-|                                         | Overall        | ClinicalTrials.gov | DRKS         |
-|-----------------------------------------|----------------|--------------------|--------------|
-| **TRN in either abstract or full-text** | 53% (723/1372) | 62% (669/1077)     | 18% (54/295) |
-| **TRN in abstract**                     | 85% (605/714)  | 97% (562/580)      | 32% (43/134) |
-| **TRN in full-text**                    | 45% (513/1131) | 54% (471/878)      | 17% (42/253) |
-
-DRKS trials were also significantly less likely than ClinicalTrials.gov trials to reference the publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
-<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)`) ClinicalTrials.gov trials referencing the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
-Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (24%, 186/783 vs.
-21%, 96/447).
+DRKS trials were also less likely than ClinicalTrials.gov trials to reference the publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
+<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r inline_text(tbl_reg_pub_link_ctgov_auto_v_manual, reference_derived, column = "stat_1", pattern = "{p}%, {n}/{N}")`) ClinicalTrials.gov trials referencing the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
+Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")` vs.
+`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_2", pattern = "{p}%, {n}/{N}")`).
 
@@
 To gauge the sensitivity and specificity of the TRN regexes, we visually inspected all PubMed secondary identifiers extracted for our sample (n = `r nrow(si)`).
-PubMed's secondary identifier field can contain identifiers not only from clinical trial registries, but also from other databanks, such as molecular sequence and gene expression identifiers or open data repository designators (i.e., figshare, Dryad) [@nationallibraryofmedicine2018].
-Most ids were clear true positive TRNs (n = `r si_tp` ) and a handful were clear true negative other ids (n = `r si_figshare`) (i.e., figshare).
-We also found `r si_drks_misformatted` ids which, based on additional PubMed data, we could determine were DRKS ids, but which were so severely misformatted that on visual inspection alone, we did not classify as TRNs (i.e., missing the preceding letters "DRKS" and just a string of numbers such as "00000711").
-As such, we consider a true negative of n = `r si_tn`.
-Our regexes correctly classified all TRNs and non-TRNs, resulting in a sensitivity and specificity of 100%.
+PubMed's secondary identifier field can contain identifiers from not only clinical trial registries, but also other databanks, such as molecular sequences and open data repositories (i.e., figshare, Dryad) [@nationallibraryofmedicine2018].
+Most ids were clear true positive TRNs (n = `r si_tp`) and a handful were clear true negative other ids (n = `r si_figshare`).
+We also found `r si_drks_misformatted` ids which we manually classified as true negative non-TRNs; however, using additional PubMed data, we determined these were severely misformatted DRKS ids (i.e., missing the preceding letters "DRKS" and just a string of numbers such as "00000711").
+Our regexes correctly classified all well formatted TRNs and non-TRNs, resulting in a sensitivity and specificity of 100%.
 If we had instead categorized the `r si_drks_misformatted` severely misformatted DRKS ids as true positives, our regexes would have sensitivity of `r style_percent(si_sens_alt, digits = 1)`% and a specificity of `r style_percent(si_spec_alt)`%.
@@
 Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research.
-This responsible practice comes at minimal costs to researchers, on the order of seconds for pasting TRNs in papers and minutes for adding a publication link to the registration.
+Poor linking poses a barrier to identifying trial publications via automated approaches and instead require researchers to perform intensive manual keyword-based searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
+This responsible practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.
 
@@
 In our sample (n = `r n_trials_ft_pdf`), `r report_n_pct_trials(n_trials_link_none)` trials had no links between registration and publication and only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication links.
+Furthermore, only `r report_n_pct_trials(n_trials_link_trn_ft_abs)` trials in our sample fully met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
+Linking practices showed at best minimal improvement over time.
+The upward trend in reporting in full-text and abstracts suggests this practice is gaining traction, however more slowly than advisable per CONSORT and ICMJE guidelines.<!--# shorten--> Trials registered in ClinicalTrials.gov were overall better linked than trials registered in DRKS.<!--# We found substantial differences between registries with ClinicalTrials.gov better linked than DRKS.
+--> These differences are in part beyond trialists' control and reliant on bibliometric databases (i.e., PubMed) and registries, namely (1) generating PubMed metadata from TRNs in the abstract or full-text, and (2) automated indexing of publications in the registry.
 
-The issue of TRN reporting in publications is also reflected in CONSORT and ICMJE guidelines.
-Only `r report_n_pct_trials(n_trials_link_trn_ft_abs)` trials in our sample fully met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
-An additional `r report_n_pct_trials(n_trials_link_trn_ft)` report in full-text only and `r report_n_pct_trials(n_trials_link_trn_abs)` did so in the abstract only.
-This situation indicates that there is much room for improvement, with involvement across stakeholders, from individual trialists and UMCs, to journals, bibliographic databases, and trial registries.
+Our findings suggest that PubMed's current approach to capturing TRNs in metadata misses TRNs from the full-text as well as DRKS trials.
+TRNs in PubMed metadata may be either provided by publishers or manually assigned by National Library of Medicine (NLM) staff who copy-and-paste TRNs found in the abstract and full-text (personal communication via NLM Helpdesk Ticket CAS-552810-T3H7V5).
+As such, we expect a TRN from any registry to be included in the metadata, if the trialist includes it in either the abstract or the full-text.
+However, we found that while trials with a TRN in the abstract were indeed more likely to include the TRN in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), TRNs in the full-text were not more likely to appear in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`).
+Furthermore, our data suggest that NLM staff are better at extracting ClinicalTrials.gov than DRKS TRNs (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).<!--# As such, it appears that PubMed metadata misses TRNs from the full-text as well as DRKS trials.-->
 
-Reporting appears to have increased over our study time in the full-text and abstracts<!--# while an increase in reporting in abstracts seems to lag behind -->.
-The upward trend suggests this practice is gaining traction, however more slowly than advisable per CONSORT and ICMJE guidelines.
-Low TRN reporting poses a barrier to identifying trial publications via automated approaches and instead require researchers to perform intensive manual keyword-based searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
-
-Trials registered in ClinicalTrials.gov showed better linking than trials registered in DRKS<!--# or be explicit that not significant for full-text -->.
-Notably, TRNs in PubMed metadata and publication references in registrations were almost three-fold more frequent in ClinicalTrials.gov than in DRKS.
-
-<!--# n addition to incorporating TRNs provided in publisher data, NLM Indexing Data Reviewers manually assign metadata tags -->
-
-TRNs in PubMed metadata may be either provided by publishers or manually assigned by National Library of Medicine (NLM) staff who copy-and-paste TRNs found in the abstract and full-text (personal communication via NLM Helpdesk Ticket CAS-552810-T3H7V5).
-As such, we would expect similar rates of TRNs in PubMed metadata across registries for trials with a TRN in abstract or full-text.
-However, even when limiting to trials with a TRN in abstract or full-text, ClinicalTrial.gov TRNs are more likely than DRKS TRNs to be included in the PubMed metadata.
-This suggests that NLM staff are more attentive to ClinicalTrial.gov TRNs and that DRKS TRNs more often go unnoticed in their manual labor.
-<!--# This is supported by our finding that TRNs were more likely to be included in PubMed metadata for ClinicalTrials.gov than DRKS trials with their TRN in the abstract or full-text.
---> Furthermore, it appears that NLM staff may be more attentive to abstracts than to full-text, since almost all trials with ClinicalTrials.gov TRNs in the abstract, but only about half of those with ClinicalTrials.gov TRNs in the full-text, had the TRN in the metadata.
-
-The almost three-fold discrepancy in the rates of publication references in registrations is driven by automatic indexing of trial publications based on TRNs in the PubMed metadata.
-ClinicalTrials.gov uses PubMed metadata to automatically link publications, whereas DRKS currently allows only for manual submission of references by trialists.
-When limiting to manually submitted publication references in registrations, performance across registries is comparable.
+Automated indexing of publications in ClinicalTrials.gov accounts for most (`r inline_text(tbl_reg_pub_link_ctgov_auto_v_manual, reference_derived, column = "stat_1", pattern = "{p}%, {n}/{N}")`) publication references in the registry and drives the almost three-fold discrepancy with DRKS (`r inline_text(tbl_link_by_reg, has_reg_pub_link, "ClinicalTrials.gov", pattern = "{p}%, {n}/{N}")` vs.
+`r inline_text(tbl_link_by_reg, has_reg_pub_link, "DRKS", pattern = "{p}%, {n}/{N}")`).<!--# Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")`)--> Currently, DRKS allows only for manual submission of references by trialists and does not index publications.<!--# add?-->
 
@@
 While our regular expressions allow for and correct minor formatting errors (such as erroneous punctuation), egregious misformatting may make the TRN undetectable.
-For example, in our visual inspection of the PubMed metadata TRNs, we found `r si_drks_misformatted` severely misformatted DRKS TRNS (i.e., numbers only with no preceding letters), which we could only identify as DRKS TRNs using additional metadata and which prevented the regex classifying them as TRNs based on the pattern alone.
-Classifying these cases as true negative, our regexes correctly classified all TRNs and non-TRNs in our sample's the PubMed metadata secondary identifier field (sensitivity and specificity of 100%).<!--# remove since not about trn formatting alone?
+For example, in our visual inspection of the PubMed metadata TRNs, we found `r si_drks_misformatted` severely misformatted DRKS TRNS (i.e., numbers only with no preceding letters), which we could only identify as DRKS TRNs using additional metadata and which prevented the regex from classifying them as TRNs based on the pattern alone.
+<!--#Classifying these cases as true negative, our regexes correctly classified all TRNs and non-TRNs in our sample's the PubMed metadata secondary identifier field (sensitivity and specificity of 100%).
+remove since not about trn formatting alone?
 Futhermore, ClinicalTrials.gov presumably relies not only on the reported TRN but also on the registry, and as such failed to index a publication (PMID: 26417987) despite the TRN (NCT01614301) appearing in the PubMed metadata, as the associated databank was incorrect (i.e., GENBANK).
@@
 In contrast with previous studies relying on PubMed queries to identify potential (randomized) clinical trial results publications, we relied on a sample of bona fide results publications from registered trials, allowing us to evaluate the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.
-
-We used an automated approach of regular expressions with high sensitivity and specificity to identify and classify publication identifiers and TRNs which allowed for a larger sample size that manual data extraction would have permitted.
-This study demonstrates the application of an automated strategy which provides a scalable method for evaluating publication links in registrations and TRN reporting in publications and can be applied to other trial sets.
+Futhermore, we used an automated approach including regular expressions with high sensitivity and specificity to identify and classify publication identifiers and TRNs which allowed for a larger sample size that manual data extraction would have permitted.
+This automated strategy is scalable and can be applied to other trial sets.
 
 The approach also faces limitations.
-Our input IntoValue dataset was limited to German UMC-led trials found on ClinicalTrials.gov and DRKS.
-It is possible that this sample and its linking rate do not reflect the broader practices across other registries and/or other countries.
-Future projects should look at similar multi-national samples of bona fide results publication.
-
+Our input IntoValue dataset comprised German UMC-led trials in ClinicalTrials.gov and DRKS and may not reflect practices across other registries and/or countries.
+Future projects should look at multi-national samples of bona fide results publications.
 Furthermore, we relied on IntoValue for trial deduplication and bona fide results publications.
-The dataset may have had a small number of unaccounted for cross-registrations (e.g., DRKS00004156 and NCT00215683) and publications that are not trial results (e.g., DRKS00009565 associated publication is 10.1097/aap.0000000000000573, a systematic review, with many reported TRNs).
-
-Finally, the approach relies quality input text data, and since journal articles are still most commonly distributed as PDFs, parsing errors may reduce text quality.
+The dataset may have had a small number of unaccounted for cross-registrations (e.g., DRKS00004156 and NCT00215683) and publications that are not trial results (e.g., systematic reviews, conference abstract books, etc.<!--# DRKS00009565 associated publication is 10.1097/aap.0000000000000573, a systematic review, with many reported TRNs-->).
+Finally, this automated approach faces limitations of the software on which it is built.
+While PubMed updated their website in 2020, their API reflects the previous backend and has subtle differences to the web version (personal communication via NLM Helpdesk Ticket CAS-575119-Y6D6Y4).
+While GROBID algorithms are trained on academic papers and have been used in large-scale bibliometric projects [e.g., @wang2020ac], parsing PDFs to XMLs may introduce some errors.<!--#Finally, the approach relies on quality input text data, and since journal articles are still most commonly distributed as PDFs, parsing errors may reduce text quality.
 We minimized this risk by relying on GROBID to parse full-text PDFs into XML with the body separated out from the abstract, header, and back.
 While GROBID algorithms are trained on academic papers and have been used in large-scale bibliometric projects [e.g., @wang2020ac], it does not perfectly classify the full-text main body.
-As such, TRNs appearing in the main body main be mistakenly excluded and TRNs appearing in the header or back material may be mistakenly included.
+As such, TRNs appearing in the main body main be mistakenly excluded and TRNs appearing in the header or back material may be mistakenly included.-->
 
@@
 
-<!--# Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research.
-This responsible practice comes at negligible costs to researchers, on the order of seconds for pasting TRNs in papers and minutes for adding a publication link to the registration.
-Our study shows that German UMCs can improve in both TRN reporting in publications and references to publications in the registration.
-While TRN reporting has improved gradually over the past decade, the rate of change is insufficient and demands a systemic change in practices across the scientific ecosystem to improve this good scientific research practice.
--->
-
 This study reveals shortcomings in the linking of German UMC-led trials registrations and results publications and highlights several promising avenues forward.
 In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is relatively straightforward and can be improved with action across stakeholders.
-Researchers should comply with CONSORT and ICMJE and include the TRN in their publication full-text and abstract.
-Research institutions like UMCs can facilitate this behavior by educating, supporting, and incentivizing their scientists to adopt this practice.
-For example, ethics review offices or UMC-core facilities for clinical research could provide researchers a reminder sheet with trial registration and reporting guidelines when reviewing trials.
-Additionally, *Leistungsorientierte Mittel (LOM)* or bonuses could be disseminated for trial registration, reporting, and reporting of registration.
-Publishers could check for correctly formatted TRNs for all submitted trial results publications, either via manual review or through simple software such as the regular expressions used for this study.
-As some already do (see for example Taylor & Francis [@taylorfrancisgroup2020]), publishers, journals, and preprint servers could also extract TRNs from the paper and/orinclude a separate TRN metadata field during manuscript submission, display this on the publication webpage, and provide this data to PubMed, CrossRef, and other bibliographic databases [@crossref2020c].
-The National Library of Medicine relies on publisher-provided data and manual indexers to create the TRN PubMed metadata if the TRN appears in the abstract or full-text.
-PubMed and other bibliographic databases could extend and semi-automate these methods in order to detect more TRNs by integrating regex-based software at least to flag potential TRNs which can then be verified by manual indexers if desired.
-Registries can support linking by highlighting the correct TRN formatting on their websites and by reminding researchers to include their TRN in publications and link their publication in the registration.
-Furthermore, they could adopt an automated linking approach using PubMed metadata, similar to that used by ClinicalTrials.gov.
-Figure \@ref(fig:flow-discussion) shows a process diagram of potential actions to improve registration-publication links from stakeholders across clinical research.
+Drawing from the study findings and additional context, Table  \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.
+
+```{r tbl-stakeholders}
+# https://www.anthonyschmidt.co/post/2020-06-03-making-apa-tables-with-gt/
+# results = "asis"
+# tbl_stakeholders %>%
+#   as_latex()
+knitr::include_graphics(here::here("docs", "figures", "tbl-stakeholders.png"))
+```
 
 With improved TRN inclusion in bibliographic metadata and increased automatic indexing of publications based on this metadata, full linking could be achieved with negligible work by researchers: simply pasting the TRN into the publication abstract and full-text.
-In our sample alone, an additional `r report_pct_num_denom_trials(n_trials_link_trn_ft_or_abs_no_si, n_trials_link_trn_ft_or_abs)` of trials with TRN in abstract or full-text could have the TRN included in PubMed's metadata, and `r report_pct_num_denom_trials(n_trials_drks_trn_si_not_linked_pub, n_trials_drks)` of DRKS trials with TRN in the PubMed metadata could be automatically indexed in the registration.
-
-```{r flow-discussion, fig.cap = "Cross-linkage of registrations and publications across the clinical research ecosystem"}
-knitr::include_graphics(here("docs", "figures", "flow-discussion.png"))
-```
+In our sample alone, an additional `r report_pct_num_denom_trials(n_trials_link_trn_ft_or_abs_no_si, n_trials_link_trn_ft_or_abs)` of trials with TRN in abstract or full-text could have the TRN included in PubMed's metadata, and an additional `r report_pct_num_denom_trials(n_trials_drks_trn_si_not_linked_pub, n_trials_drks_trn_si)` of DRKS trials with TRN in the PubMed metadata could be automatically indexed in the registration.
+With the uptake of the recommended actions across stakeholder groups, improved trial registration-publication links will allow for more comprehensive evidence synthesis and increase the value of clinical research.
 
@@
 
-# References {.unnumbered}
+\newpage
 
-::: {#refs}
+# References
+
+```{=tex}
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+```
+::: {#refs custom-style="Bibliography"}
 :::
 
-# (APPENDIX) Appendix {.unnumbered}
+\endgroup
 
-# TRN Regular Expressions
-
-```{r tbl-trn-regexes}
-ctregistries::registries %>% 
-  select(
-    Registry = registry,
-    "Registry Title" = databank_full_name,
-    "TRN Regex" = trn_regex
-  ) %>% 
-  flextable::flextable(cwidth = 2) %>% 
-  flextable::set_caption(caption = "Trial registration number regular expressions")
-```
-
-# Sensitivity Analyses
+# ```{r echo = FALSE, results = 'asis', cache = FALSE}
+# papaja::render_appendix('appendix.Rmd')
+# ```

### 2021-07-26 R/analysis/04_link-reg-pub.R

@@ -242,6 +242,15 @@ n_trials_drks_trn_si_not_linked_pub <-
   ) %>%
   nrow()
 
+# How many drks trials have trn in si
+n_trials_drks_trn_si <-
+  trials %>%
+  filter(
+    registry == "DRKS",
+    has_iv_trn_secondary_id
+  ) %>%
+  nrow()
+
 
 # TODO
 # How many trials have linked pub, excluding derived?

### 2021-07-26 R/analysis/reg-auto-link.R

@@ -1,0 +1,57 @@
+# Manually-linked publication in registration -----------------------------
+
+tbl_reg_pub_link_manual <-
+  trials %>%
+  filter(!reference_derived | is.na(reference_derived)) %>%
+  select(registry, has_reg_pub_link) %>%
+  gtsummary::tbl_summary(
+    by = registry,
+    statistic = everything() ~ "{p}% ({n} / {N})",
+    label = everything() ~ "Publication in registration"
+  ) %>%
+
+  # Remove rowname label
+  modify_header(label = "") %>%
+
+  # Remove N from column names
+  modify_header(all_stat_cols() ~ "**{level}**") %>%
+
+  # Add overall column and remove N
+  add_overall() %>%
+  modify_header(stat_0 ~ "**Overall**") %>%
+
+  # Remove footnote
+  modify_footnote(everything() ~ NA) %>%
+  bold_labels()
+
+
+# ClinicalTrials.gov publication in registration: auto vs. manual --------
+
+tbl_reg_pub_link_ctgov_auto_v_manual <-
+
+  trials %>%
+  filter(
+    registry == "ClinicalTrials.gov",
+    has_reg_pub_link
+  ) %>%
+  # tbl_cross(
+  #   has_reg_pub_link, reference_derived,
+  #   percent = "row",
+  #   margin = "row"
+  # )
+  select(registry, reference_derived) %>%
+  gtsummary::tbl_summary(
+    by = registry,
+    statistic = everything() ~ "{p}% ({n} / {N})",
+    label = everything() ~ "Automated link"
+  ) %>%
+
+  # Remove rowname label
+  modify_header(label = "") %>%
+
+  # Remove N from column names
+  modify_header(all_stat_cols() ~ "**{level}**") %>%
+
+  # Remove footnote
+  modify_footnote(everything() ~ NA) %>%
+  bold_labels()

### 2021-07-26 R/analysis/si-if-abs-ft.R

@@ -1,0 +1,40 @@
+# Function to create tables for si by registry ----------------------------
+
+tbl_si_registry <- function(var, var_name) {
+  trials %>%
+    mutate(has_iv_trn_abs_or_ft = has_iv_trn_abstract | has_iv_trn_ft_pdf) %>%
+    filter({{var}}) %>%
+    select(registry, has_iv_trn_secondary_id) %>%
+    gtsummary::tbl_summary(
+      by = registry,
+      statistic = has_iv_trn_secondary_id ~ "{p}% ({n} / {N})",
+      label = has_iv_trn_secondary_id ~ var_name
+    ) %>%
+
+    # Remove rowname label
+    modify_header(label = "") %>%
+
+    # Remove N from column names
+    modify_header(all_stat_cols() ~ "**{level}**") %>%
+
+    # Add overall column and remove N
+    add_overall() %>%
+    modify_header(stat_0 ~ "**Overall**") %>%
+
+    # Remove footnote
+    modify_footnote(everything() ~ NA) %>%
+    bold_labels()
+}
+
+
+# Prepare abstract and full-text tables -----------------------------------
+tbl_si_abs_or_ft <- tbl_si_registry(has_iv_trn_abs_or_ft, "TRN in abstract or full-text")
+tbl_si_abs <- tbl_si_registry(has_iv_trn_abstract, "TRN in abstract")
+tbl_si_ft <- tbl_si_registry(has_iv_trn_ft_pdf, "TRN in full-text")
+
+
+# Display si table --------------------------------------------------------
+
+tbl_si_if_abs_ft <-
+  tbl_stack(list(tbl_si_abs_or_ft, tbl_si_abs, tbl_si_ft)) %>%
+  modify_caption("Number and proportion of trials with TRN in metadata given TRN in abstract, full-text, or either. Denominator indicates number of trials with TRN in abstract and/or full-text, from the respective registry.")

### 2021-07-26 R/analysis/tbl-stakeholders.R

@@ -1,0 +1,84 @@
+# Prepare actions and examples --------------------------------------------
+
+
+action_researchers <- "
+- Include TRN in abstract and full-text per CONSORT and ICMJE guidelines
+- Link publication in registry
+"
+
+example_researchers <- "—"
+
+action_registries <- "
+- Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registry
+- Link publications automatically using TRNs in bibliographic database metadata
+"
+
+example_registries <- "ClinicalTrials.gov uses PubMed TRN metadata to automatically link publications in the registration  [@nationallibraryofmedicine2017c]. Our study shows this automated publication linking is responsible for the majority of linked publications in ClinicalTrials.gov and implies similar potential for DRKS."
+
+action_institutions <- "
+- Educate, support, and incentivize researchers on linking
+"
+
+example_institutions <- "
+Ethics review offices or UMC-core facilities for clinical research could give researchers a info sheet with trial registration and reporting guidelines when reviewing trial. Bonuses, i.e., *Leistungsorientierte Mittel (LOM)*, could be disseminated for trial registration, reporting, and linking.
+"
+
+action_journals <- "
+- Request TRN in specialized metadata field
+- Review abstract and full-text for TRN inclusion, using automated regexes and/or manual strategies
+- Provide TRN as metadata to bibliographic databases
+"
+
+example_journals <- "
+Taylor & Francis extracts TRNs from the abstract and full-text, submit this metadata to CrossRef and Pubmed, and displays the linked trial via Crossmark on the article page [@taylorfrancisgroup2020; @crossref2020c])
+"
+
+action_pubmed <- "
+- Integrate publisher-provided TRNs into metadata
+- Extract TRN as metadata from abstract and full-text, using automated regexes and/or manual strategies
+"
+
+example_pubmed <- "
+The National Library of Medicine relies on publisher-provided data and manual indexers to create the TRN PubMed metadata if the TRN appears in the abstract or full-text. Our study shows this manual-only strategy misses TRNs and could be semi-automated to detect more TRNs for manual verfication.
+"
+
+
+# Prepare stakeholder table
+tbl_stakeholders <-
+  dplyr::tribble(
+    ~Stakeholder, ~Recommendation, ~Example,
+    "Researchers", action_researchers, example_researchers,
[…]

### 2021-08-02 docs/manuscript_papaja.Rmd
Update manuscript

@@
 abstract: |
-  **Objective** To evaluate, using a novel automated approach, the links between registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
-The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make both clinical trial publications and registrations findable and thus increase the value of clinical research.
+  **Objective** To evaluate, using a novel automated approach, links between registration and publication across clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make clinical trial publications and registrations findable and increase the value of clinical research.
 
@@
 
-  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials comprehensively link their registrations and publications, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata.
+  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata.
 In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links.
-The most common link was TRN reported in the full-text, which occurred for `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_trn_ft")` trials.
+The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_trn_ft")`).
 ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists.
-Of trials reporting a TRN in the abstract, ClinicalTrials.gov trials were more likely than DRKS trials to have the TRN captured in the PubMed metadata.
-Additionally, most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_ctgov_link_reg_trn_si_derived")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
+Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata.
+Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_ctgov_link_reg_trn_si_derived")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
 
@@
 keywords          : "clinical trials, registration, reporting, meta-research"
-wordcount         : "3714"
+wordcount         : "3551"
 
@@
   - \usepackage{orcidlink}
+
+always_allow_html: true
 ---
@@
 library(papaja)
+# library(knitr)
+# library(kableExtra)
+
+library(rentrez)
+library(roadoi)
+theme_gtsummary_compact()
 
@@
 r_refs("r-references.bib")
-r_citations <- 
-  cite_r(file = "r-references.bib", footnote = TRUE) # pkgs, withhold #r_citations$r and r_citations$pkgs
+r_citation <- cite_r() # file = "r-references.bib", footnote = TRUE, pkgs, withhold #r_citations$r and r_citations$pkgs
 
@@
 abs_n_trials_link_none <- report_n_pct_trials(n_trials_link_none)
-abs_n_trials_link_trn_ft <- report_n_pct_trials(n_trials_link_trn_ft)
+abs_n_trials_link_trn_ft <- "1131, 60% of trials"
+  # Change to gtsummary once updated: https://stackoverflow.com/questions/68502977/gtsummaryinline-text-for-stat-0-with-pattern
+  # inline_text(tbl_link_by_reg, has_iv_trn_ft_pdf, stat_0, pattern = "{n}, {p}% of trials")
+  # report_n_pct_trials(n_trials_link_trn_ft)
 
@@
 
-Clinical trial results risk being left out of evidence synthesis and health policy decision making if publications cannot be found.
+Clinical trial evidence informs clinical practice and health policy decision making through reviews and meta-analyses.
+The validity of health evidence synthesis is jeopardized if relevant trial data are left out to do inadequate or incomplete trial metadata, or inability to find trial results through automated means.
 Incomplete evidence threatens healthcare and the lives and welfare of human volunteers.
-Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp].
+Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp; @chalmers2013b].
 Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing clinical research value [@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->.
@@
 
-Trial evidence appears to be insufficiently linked in both registries and publications.
-Previous studies on trial registrations found links to results publications in as few as 13% to as many as 42% of trials [@bashir2017sr; @huser2013po].
-TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b].
-For example, some studies included trial publications limited to specific journals or fields, whereas others relied on PubMed "clinical trial" publication type queries.
-
-In order to draw robust conclusions about registration-publication linking, the analysis population must be limited to registered clinical trials with published results.
-Previous projects faced limitations due to the unclear publication cohort and unknown true registration rates.
-PubMed queries currently do not suffice to identify the true populations of clinical trials or even RCTs, both excluding true positives (i.e., sensitivity) and including false positives (i.e., specificity) [@glanville2020jmla].
-Therefore not all PubMed-indexed clinical trials or RCTs should be registered <!--# as they may not reflect true clinical trials or RCTs -->, and not all that should be are.
-Furthermore, not all reported TRNs correspond to bona fide registrations of the trial whose results are reported in the publication, i.e., TRNs may be reported for related trials in the background section.
-As such, conclusions drawn by these previous studies relying on PubMed-query based samples, about registration reporting rates must be qualified as the rate of reported trial registration numbers in PubMed-indexed (randomized) clinical trials with unknown registration rates.
-<!--# remove previous sentence?
---> Similarly, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link.
-Furthermore, these studies of publication links in registrations have focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of previous studies investigating publication links in DRKS.
+Trial evidence appears to be insufficiently linked in both registrations and publications.
+Previous studies on trial registrations found links to results publications in as few as 13% to as many as 60% of trials [@bashir2017sr; @huser2013po; @wetering2012po].
+TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b; @huser2013jamia; @wetering2012po].
+The sampling strategies used in these previous studies limit our ability to draw conclusions about registration-publication linking.
+Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included non-trial-result publications, and trials with unknown registration rates (as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo]); registration-based cohorts included included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b].
+Furthermore, these studies of publication links in registrations have primarily focused on ClinicalTrials.gov, and we are not aware of previous studies investigating publication links in DRKS.
+In this study, we limit our sample to registered clinical trials with published results which can be expected to link the registration and publication, and which allows us to draw conclusions about the prevalence of this responsible research practices.
 
@@
 
-We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset.
-In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches [ADD DATASET REFERENCE].
+We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset [@riedel2021a].
+In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.
 
@@
 
-We queried the PubMed Entrez Programming Utilities (E-utilities) API<!--#on DAY MONTH YEEAR--> for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter].
+We queried the PubMed Entrez Programming Utilities (E-utilities) API<!--#on DAY MONTH YEAR--> for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @R-rentrez].
 From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->
@@
 We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies?
--->[@ourresearch; @jahn].
+-->[@ourresearch; @R-roadoi].
 We then used the GROBID machine learning library for technical and scientific publications (v0.6.1) via Python to parse the PDFs into machine-readable XMLs [@2008; @lopez2021].
@@
 
-Figure \@ref(fig:flow-methods) show the methods flow using automated techniques<!--# with the exception of some full-text PDFs -->.
-
-```{r flow-methods, fig.cap = "Data collection sources and methods", out.width="50%"}
-knitr::include_graphics(here("docs", "figures", "flow-methods.png"))
-```
-
 ### Inclusion and exclusion criteria
 
-After updating registry data, we reapplied the IntoValue exclusion criteria to exclude trials that were completed before 2009 or after 2017, were non-interventional, or did not meet the recruitment or study status inclusion criteria (see @wieschowski2019joce and @riedel2021 for further details on these criteria).
-We also excluded trials without a German UMC lead, as only one of the two IntoValue cohorts included trials not led by German UMCs.
-For duplicate trials (i.e., appearing in both IntoValue cohorts), we excluded the earlier record.
-We also excluded trials without results, and trials with results for which we could not find a PMID and full-text publication.
+After updating registry data, we reapplied the IntoValue exclusion criteria (completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead; see @riedel2021 for further details on these criteria) and limited our sample to trials with results with a PMID and full-text publication.
 
@@
 
-As reference data is available in difference formats depending on the registry, we developed parallel approaches to extract publication identifiers for each.
+As reference data is available in different formats depending on the registry, we developed parallel approaches to extract publication identifiers for each.
 For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT.
@@
 
-We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
+We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (available at <https://github.com/maia-sh/ctregistries/blob/master/inst/extdata/registries.csv>) and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
 To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata<!--# for true positives and true negatives-->.
@@
 We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source uncovered through the cleaning process.
-We also coalesced across sources to deduce a list of unique TRNs by publication along with the source(s) in which each was reported.
+We also merged our sources to produce a list of unique TRNs by publication along with the source(s) in which each was reported.
 
-The output of this TRN extraction, cleaning, and deduplication included TRNs beyond those of the known registrations.
+The output of TRN extraction, cleaning, and deduplication included TRNs beyond those of the known registrations.
 These additional TRNs could be cross-registrations of the same trial or provided as background or discussion.
@@
 We generated descriptive statistics on trial and publication characteristics, overall and by registry.
-We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran exploratory logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables.
-In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata.
+We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables.
+Regressions were performed as exploratory analyses, and all variables were included in each regression model.<!--#no model selection or fitting was conducted; the explanatory variables were tested a priori to verify there was no violation of the assumption of the linearity of the logit TODO, ask Murph--> In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata.
 Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs.
@@
 
-Data collection, preparation, and analysis was performed in `r r_citations$r`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> All code and the final analysis data for this study are available at <https://github.com/maia-sh/reg-pub-link>.
+Data collection, preparation, and analysis was performed in `r r_citation`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> All code, the final analysis data, and a list of all packages used for this study are available at <https://github.com/maia-sh/reg-pub-link>.
 Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>.
-`r r_citations$pkgs`
 
@@
 Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.
-<!--# QUESTION: what summary info about trials and publications?
-I also include a long version of demographic data in case we want to include any of these, ideally with collapsed categories, in Table @ref(tab:tbl-trials-full) -->
 
@@
 
-```{r tbl-trials}
+<!-- (ref:tbl-trials-caption) **Characteristics of German UMC-led trials with published results** A trial was considered randomized if allocation included randomization.
+A trial was considered prospectively registered if registered in the same or previous months to start date.
+Summary results were taken from a structured data field in ClinicalTrials.gov, and determined based on manual inspection for terms such as *Ergebnisbericht* or *Abschlussbericht* in DRKS.
+*Top* journals refer to the journals with the greatest number of trial publications in our sample.
+-->
+
+```{r tbl-trials, results = "asis"}
+# tbl <- read_rds(here("docs", "figures", "tbl-trials.rds"))
+# tbl %>% 
+#   as_kable_extra(booktabs = TRUE)
 tbl_trials %>%
-  modify_caption("Characteristics of included trials with results publications")
+
+  as_kable_extra()
+  # as_tibble() %>%
+  # apa_table(
+  #   caption = "This is a table about trials",
+  #   format.args = list(na_string = ""),
+  #   format = "pandoc"
+  # )
+# modify_caption("Characteristics of included trials with results publications")
   # gtsummary::as_flex_table() %>% 
@@
 tbl_link_by_reg %>% 
-  modify_caption("Registration-publication links by registry")
-  # gtsummary::as_flex_table() %>% 
+  modify_caption("Registration-publication links by registry") #%>% 
+  # gtsummary::as_flex_table() #%>%
   # flextable::set_caption(caption = "Registration-publication links by registry")
@@
 tbl_trials_regression %>% 
-  modify_caption("Crude and adjuusted odds ratios for factors associated with publication-registration links")
+  modify_caption("Crude and adjusted odds ratios for factors associated with publication-registration links")
   # gtsummary::as_flex_table() %>% 
@@
 Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables.
-Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(cor_abs, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(cor_ft, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
+Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
 Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.
@@
 `r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_2", pattern = "{p}%")`).
-Table \@ref(tab:tbl-si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
-
-```{r tbl-si-if-abs-ft}
-tbl_si_if_abs_ft
-```
+Appendix \@ref(si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
 
@@
 Poor linking poses a barrier to identifying trial publications via automated approaches and instead require researchers to perform intensive manual keyword-based searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
-This responsible practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.
+This responsible research practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.
 
@@
 In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is relatively straightforward and can be improved with action across stakeholders.
-Drawing from the study findings and additional context, Table  \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.
+<!--#Drawing from the study findings and additional context, -->Table \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.
 
@@
 knitr::include_graphics(here::here("docs", "figures", "tbl-stakeholders.png"))
+# tbl_stakeholders
 ```
@@
 
-We would like to thank Kerstin Rubarth from the Charité Institut für Biometrie und Klinische Epidemiologie for her consultation on the statistical analysis.
+We would like to thank Kerstin Rubarth from the Charité Institut für Biometrie und Klinische Epidemiologie for her consultation on the statistical analysis, and Nicholas Devito for earlier versions of the trial registration number regular expressions.
 
@@
 
-# ```{r echo = FALSE, results = 'asis', cache = FALSE}
-# papaja::render_appendix('appendix.Rmd')
-# ```
+```{r render_appendix, include=FALSE}
+render_appendix("appendix.Rmd")
+```

### 2021-08-02 R/analysis/03_tbl-trials.R
Update trials table 1 with additional vars

@@ -61,27 +61,56 @@
 
 # Summarize selected trial info -------------------------------------------
 
+# ClinicalTrials.gov and DRKS use different phase names
+# We use IntoValue's lookup table to coalesce names
+# Repo is currently private so using URL and later will switch to repo
+# registries_lookup_path <- "https://github.com/quest-bih/IntoValue2/blob/master/data/2_dataset_cleaning/final_dataset/iv_data_lookup_registries.csv"
+
+registries_lookup_path <- fs::path(fs::path_dir(here()), "IntoValue2", "data/2_dataset_cleaning/final_dataset/iv_data_lookup_registries.csv" )
+
+phase_lookup <-
+  read_csv(registries_lookup_path) %>%
+  filter(name == "phase") %>%
+  select(phase = level_registry, phase_unified = level_unified)
+
+N_JOURNALS <- 6
+
 tbl_trials <-
   trials %>%
 
-  # Add pubmed journal info and munge so display top 5
+  # Add pubmed journal info and munge so display top N_JOURNALS
   left_join(select(pubmed_all, journal, pmid), by = "pmid") %>%
-  # group_by(registry) %>%
   mutate(
     journal = if_else(journal == "Lancet (London, England)", "Lancet", journal),
-    journal = forcats::fct_lump_n(journal, n = 5, other_level = NA),
+
+
+    # journal = stringi::stri_trans_totitle(journal),
+    journal = gsub("\\b([a-z])", "\\U\\1", journal, perl=TRUE),
+    journal = str_replace(journal, "Of", "of"),
+
+    journal = forcats::fct_lump_n(journal, n = N_JOURNALS, other_level = NA),
     journal = forcats::fct_infreq(journal)
   ) %>%
+
+
+  # Prepare industry sponsor
+  mutate(industry_sponsor = ifelse(main_sponsor == "Industry", TRUE, FALSE)) %>%
+
+  # Prepare phase
+  left_join(phase_lookup, by = "phase") %>%
 
   select(
     registry,
     days_cd_to_publication,
-    # days_cd_to_summary,
-    # has_summary_results,
-    # is_multicentric,
-    # enrollment,
-    is_randomized,
+    days_cd_to_summary,
+    has_summary_results,
     is_prospective,
+    is_randomized,
+    is_multicentric,
+    industry_sponsor,
+    enrollment,
+    phase_unified,
+    # center_size
     # n_crossreg_reg,
[…]

### 2021-08-02 R/analysis/04_link-reg-pub.R
Remove differences from link prop table

@@ -278,30 +278,34 @@ tbl_link_by_reg <-
   trials %>%
   select(
     registry,
-    has_reg_pub_link,
-    has_iv_trn_secondary_id,
+    has_iv_trn_ft_pdf,
     has_iv_trn_abstract,
-    has_iv_trn_ft_pdf
+    has_iv_trn_secondary_id,
+    has_reg_pub_link
   ) %>%
   gtsummary::tbl_summary(
     by = registry,
     label = list(
-      has_reg_pub_link ~ "Reference in registration",
-      has_iv_trn_secondary_id ~ "TRN (PubMed metadata)",
+      has_iv_trn_ft_pdf ~ "TRN (Full-text)",
       has_iv_trn_abstract ~ "TRN (Abstract)",
-      has_iv_trn_ft_pdf ~ "TRN (Full-text)"
+      has_iv_trn_secondary_id ~ "TRN (PubMed metadata)",
+      has_reg_pub_link ~ "Reference in registration"
     )
   ) %>%
 
   #TODO: figure out what test/stats to include
-  add_difference() %>%
+  # add_difference() %>%
   # add_p() %>%
   add_overall() %>% #show_header_names()
   # Move stats legend to each line
   # add_stat_label(location = NULL) %>%
-  modify_header(label = "**Registration-Publication Linkage**") %>%
+  # modify_header(label = NULL) #%>%
+  modify_header(label = "") %>%
+  # modify_header(label = "**Registration-Publication Linkage**") %>%
   # modify_caption("**German UMC-led trials with published result** (N = {N})") %>%
-  bold_labels()
+  bold_labels() %>%
+  modify_footnote(everything() ~ NA)
+
 
 # TODO: Continuity correct set to FALSE. is this ok?
 # chisq_link_by_reg_pdf_tidy <-

### 2021-08-02 R/analysis/06_plot-change-over-time.R
Update change over time to facet by link type

@@ -1,78 +1,6 @@
 # Percentage of trials reg-pub linkage, by year
 # Denominator n = 1902 (n trials)
 
-
-# Prepare data: reg-pub linkage per completion year -----------------------
-
-links_by_year <-
-  trials %>%
-  select(
-    id,
-    # registry, # could also group by registry
-    completion_year,
-    has_reg_pub_link,
-    has_iv_trn_secondary_id,
-    has_iv_trn_abstract,
-    has_iv_trn_ft_pdf
-  ) %>%
-  rename(
-    "Reference in registration" = has_reg_pub_link,
-    "TRN (PubMed metadata)" = has_iv_trn_secondary_id,
-    "TRN (Abstract)" = has_iv_trn_abstract,
-    "TRN (Full-text)" = has_iv_trn_ft_pdf
-  ) %>%
-  pivot_longer(c(-id, -completion_year),
-               names_to = "link_type",
-               values_to = "link",
-               values_drop_na = TRUE
-  ) %>%
-  group_by(link_type, completion_year) %>%
-  summarise(
-    n_link_type= n(),
-    n_link = sum(link),
-    prop_link = n_link/n_link_type,
-    .groups = "drop"
-  )
-
-
-# Plot reg-pub linkage per completion year --------------------------------
-
-plot_links_year <-
-  links_by_year %>%
-  ggplot(aes(x = completion_year, y = prop_link, color = link_type)) +
-  geom_point() +
-  geom_line() +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_x_continuous(breaks = unique(links_by_year$completion_year)) +
-  scale_colour_viridis_d() +
-  #TODO: show ticks for all years
-  labs(
-    y = "Registration-Publication Linkage (%)",
-    x = "Completion Year"
-  ) +
-  guides(colour = guide_legend(title = "Registration-Publication Link Type")) #+
-  # theme(legend.position= "top"
-        # legend.justification="top"
-        # )
-
-# ggsave(here::here("docs", "figures", "plot-links-year.png"), plot_links_year)
-
-
-# Plot reg-pub linkage per completion year,  with glm ---------------------
-
-
-# TODO: Decide whether to add trend line
-plot_links_year_glm <-
-  plot_links_year +
-  geom_smooth(
-    method = glm, #lm probably not appropriate with binary output
-    se = TRUE, #not readable with SE but need to show I think
-    alpha = 0.2
-  )
-
[…]

### 2021-08-02 R/analysis/logistic-regression.R
Change reference lines in log reg

@@ -30,17 +30,28 @@ tbl_glm_mv_trials <- function(trials_formula) {
     data = trials_model,
     family = "binomial"
   ) %>%
+
     gtsummary::tbl_regression(
       exponentiate = TRUE,
       show_single_row = c(-completion_year, -registry),
       # add_estimate_to_reference_rows = TRUE # Change reference row
     ) %>%
+
     modify_table_styling(
       column = estimate,
       rows = !is.na(estimate),
       cols_merge_pattern = "{estimate} ({ci}; {p.value})",
       label = "**aOR (95% CI; p-value)**"
     ) %>%
+
+    # Change reference label
+    # Thanks to gtsummary dev: https://github.com/ddsjoberg/gtsummary/issues/942
+    modify_table_styling(
+      column = estimate,
+      rows = reference_row %in% TRUE,
+      missing_symbol = "1 (Ref)"
+    ) %>%
+
     modify_header(update = list(label ~ ""))
 }
 
@@ -49,7 +60,9 @@ tbl_glm_mv_trials <- function(trials_formula) {
 tbl_glm_uv_trials <- function(outcome) {
 
   trials_model %>%
+
     select(completion_year, registry, has_iv_trn_secondary_id, has_iv_trn_ft_pdf, has_reg_pub_link, has_iv_trn_abstract) %>%
+
     tbl_uvregression(
       method = glm,
       y = {{outcome}},
@@ -67,6 +80,14 @@ tbl_glm_uv_trials <- function(outcome) {
       label = "**cOR (95% CI; p value)**",
       # footnote_abbrev = "cOR = Crude Odds Ratio,  CI = Confidence Interval "
     ) %>%
+
+    # Change reference label
+    modify_table_styling(
+      column = estimate,
+      rows = reference_row %in% TRUE,
+      missing_symbol = "1 (Ref)"
+    ) %>%
+
     modify_header(update = list(label ~ ""))
   # modify_footnote(or ~ "cOR = Crude Odds Ratio")
 }
@@ -79,6 +100,7 @@ aor_reg <-
     "has_reg_pub_link ~ completion_year + registry +
       has_iv_trn_secondary_id + has_iv_trn_abstract + has_iv_trn_ft_pdf"
   )
+
 
 aor_si <-
   tbl_glm_mv_trials(
@@ -124,9 +146,17 @@ tbl_trials_regression <-
[…]

### 2021-08-19 R/09_combine-ctgov-drks-trials.R
Dump script updates

@@ -93,7 +93,8 @@ ctgov_references <-
   )
 
 registry_references <-
-  bind_rows(ctgov_references, drks_references) %>%
-  filter(!(is.na(doi) & is.na(pmid)))
+  bind_rows(ctgov_references, drks_references) #%>%
+  # filter(!(is.na(doi) & is.na(pmid))) %>%
+  # distinct()
 
 write_rds(registry_references, here::here("data", "processed", "registry-references.rds"))

### 2021-08-19 R/11_prepare-trials.R

@@ -1,4 +1,5 @@
 library(dplyr)
+library(tidyr)
 library(readr)
 library(here)
 library(fs)
@@ -82,25 +83,49 @@ trials <-
     preconditions = ~ . %>% filter(has_pubmed & has_ft_pdf)
   )
 
-# Add info about registry reference links
+# Add info about registry reference links: any, any with doi or pmid, intovalue
+
+# Number of references in registration
+references_any <-
+  registry_references %>%
+  count(id, name = "n_reg_pub_any")
+
+# Limit references to unique with doi or pmid
+registry_references <-
+  registry_references %>%
+  filter(!(is.na(doi) & is.na(pmid))) %>%
+  distinct()
+
+# Number of references with doi or pmid in registration
+references_doi_or_pmid <-
+  registry_references %>%
+  count(id, name = "n_reg_pub_doi_or_pmid")
+
 pmid_references <-
   registry_references %>%
   select(-doi, -reference_type) %>%
-  tidyr::drop_na(pmid) %>%
+  drop_na(pmid) %>%
   mutate(pmid_link = TRUE) %>%
   distinct()
 
 doi_references <-
   registry_references %>%
   select(-pmid, -reference_type) %>%
-  tidyr::drop_na(doi) %>%
+  drop_na(doi) %>%
   mutate(doi_link = TRUE) %>%
   distinct()
 
 trials <-
   trials %>%
 
-  # Join in registry references by pmid
+  # Join in registry references (any, and pmid or doi)
+  left_join(references_any, by = "id") %>%
+  left_join(references_doi_or_pmid, by = "id") %>%
+  mutate(
+    n_reg_pub_any = replace_na(n_reg_pub_any, as.integer(0)),
+    n_reg_pub_doi_or_pmid = replace_na(n_reg_pub_doi_or_pmid, as.integer(0))
+  ) %>%
+  # Join in intovalue registry references by pmid
   left_join(pmid_references, by = c("id", "pmid")) %>%
   left_join(doi_references, by = c("id", "doi")) %>%
 
@@ -235,7 +260,9 @@ trials <-
     "n_crossreg_secondary_id",
     "n_crossreg_abstract",
     "n_crossreg_ft_pdf",
-    "n_crossreg_reg"
+    "n_crossreg_reg",
+    "n_reg_pub_any",
+    "n_reg_pub_doi_or_pmid"
   )
 
 # Check that all intovalue columns in trials

### 2021-08-19 R/analysis/04_link-reg-pub.R

@@ -334,3 +334,16 @@ chisq_link_by_reg_pdf <-
 # https://calerovaldez.com/post/using-the-apastats-package-to-write-reproducible-reports/
 # http://frederikaust.com/papaja_man/reporting.html
 # https://benwhalley.github.io/just-enough-r/apa-output.html
+
+# How many publications (including iv) are listed in registrations?
+
+reg_pub_max_drks <-
+  trials %>%
+  filter(registry == "DRKS") %>%
+  slice_max(n_reg_pub_any)
+
+reg_pub_max_ctgov <-
+  trials %>%
+  filter(registry == "ClinicalTrials.gov") %>%
+  slice_max(n_reg_pub_any)
+

### 2021-08-19 R/analysis/05_upset-link-reg-pub.R

@@ -41,25 +41,25 @@ trials_no_links <-
   ) %>%
   mutate(links = list(NULL))
 
-plot_upset_links_reg_pub <-
-  bind_rows(trials_links, trials_no_links) %>%
-  ggplot(aes(x = links)) +
-  geom_bar() +
-  geom_text(stat='count', aes(label = after_stat(count)), vjust = -.5) +
-  ggupset::scale_x_upset() +
-  ylab("Number of trials") + #not number of TRNs since 1 reg could have multi-trn
-  xlab(NULL) +
-  ggupset::theme_combmatrix(
-    # combmatrix.label.make_space = TRUE,
-    combmatrix.panel.line.size = 0,
-    combmatrix.label.text = element_text(size=11)
-  )
+# plot_upset_links_reg_pub <-
+#   bind_rows(trials_links, trials_no_links) %>%
+#   ggplot(aes(x = links)) +
+#   geom_bar() +
+#   geom_text(stat='count', aes(label = after_stat(count)), vjust = -.5) +
+#   ggupset::scale_x_upset() +
+#   ylab("Number of trials") + #not number of TRNs since 1 reg could have multi-trn
+#   xlab(NULL) +
+#   ggupset::theme_combmatrix(
+#     # combmatrix.label.make_space = TRUE,
+#     combmatrix.panel.line.size = 0,
+#     combmatrix.label.text = element_text(size=11)
+#   )
 
-ggsave(
-  here("docs", "figures", "plot-upset-links-reg-pub.png"),
-  plot_upset_links_reg_pub,
-  scale = 1.25
-)
+# ggsave(
+#   here("docs", "figures", "plot-upset-links-reg-pub.png"),
+#   plot_upset_links_reg_pub,
+#   scale = 1.25
+# )
 
 plot_upset_links_reg_pub_registry <-
   bind_rows(trials_links, trials_no_links) %>%
@@ -80,6 +80,18 @@ plot_upset_links_reg_pub_registry <-
   ) +
 
   theme(legend.position = c(.85, .9))
+
+# plot_upset_links_reg_pub_registry +
+#
+#   # Move y axis label closer to plot
+#   # Thanks to https://stackoverflow.com/questions/68593982
+#   theme(axis.title.y=element_blank()) +
+#   annotate(geom = "text", x = -0.2, y = 6500, label = "count", angle = 90, size=4) +
+#   coord_cartesian(xlim = c(1, 8), clip = "off")
+
+
[…]

### 2021-08-19 R/analysis/06_plot-change-over-time.R

@@ -35,28 +35,33 @@ links_by_year_reg <-
 
 # Prepare plot ------------------------------------------------------------
 library(showtext)
-font_add_google("Poppins", "Poppins")
+# font_add_google("Poppins", "Poppins")
+font_add_google("Roboto", "Roboto")
 font_add_google("Roboto Mono", "Roboto Mono")
 showtext_auto()
 
-theme_set(theme_light(base_size = 18, base_family = "Poppins"))
-theme_set(ggthemes::theme_fivethirtyeight(base_size = 18, base_family = "Poppins"))
+theme_set(theme_light(base_size = 18, base_family = "Roboto"))
+# theme_set(theme_light(base_size = 18, base_family = "Poppins"))
+# theme_set(ggthemes::theme_fivethirtyeight(base_size = 18, base_family = "Poppins"))
 
-# links_by_year_reg_separated <-
-links_by_year_reg %>%
-  ggplot(aes(x = completion_year, y = prop_link, color = registry, linetype = registry)) +
+plot_links_year_reg <-
+  links_by_year_reg %>%
+  rename(Registry = registry) %>%
+  ggplot(aes(x = completion_year, y = prop_link, color = Registry, linetype = Registry)) +
   facet_wrap(~link_type, ncol = 1) +
   geom_point() +
   geom_line() +
   scale_y_continuous(labels = scales::percent, limits = c(0, 1), expand = c(0.005, 0.005)) +
   scale_x_continuous(breaks = unique(links_by_year_reg$completion_year)) +
-  scale_color_grey() +
+  scale_color_manual(values = c("gray60", "gray10")) +
+  # scale_color_grey(start = 0.5, end = 0.8) +
   # scale_colour_viridis_d() +
   #TODO: show ticks for all years
   labs(
     # y = "Registration-Publication Linkage (%)",
     y = NULL,
-    x = "Trial Completion Year"
+    x = "Trial Completion Year",
+    title = "Percentage of linked registrations and publication by trial completion year"
   ) +
   guides(
     # colour = guide_legend(title = "Registration-Publication Link Type"),
@@ -66,15 +71,26 @@ links_by_year_reg %>%
   ) +
   theme(
     legend.position = "top",
-    legend.box = "vertical",
+    legend.justification = "left",
+    # legend.box.just = "left",
+    # legend.box = "vertical",
     legend.margin = margin(),
     axis.title = element_text(size = 16),
-    axis.text = element_text(family = "Roboto Mono", size = 12),
[…]

### 2021-08-19 R/analysis/alt-plot-change-over-time.R

@@ -1,63 +1,0 @@
-# STACKED -----------------------------------------------------------------
-
-plot_links_year_reg_stacked <-
-links_by_year_reg %>%
-  ggplot(aes(x = completion_year, y = prop_link, color = link_type)) +
-  facet_wrap(~registry, ncol = 1) +
-  geom_point() +
-  geom_line() +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_x_continuous(breaks = unique(links_by_year$completion_year)) +
-  scale_colour_viridis_d() +
-  #TODO: show ticks for all years
-  labs(
-    y = "Registration-Publication Linkage (%)",
-    x = "Completion Year"
-  ) +
-  guides(colour = guide_legend(title = "Registration-Publication Link Type")) +
-  theme(legend.position= "top")
-
-
-# LINETYPE ----------------------------------------------------------------
-
-plot_links_year_reg_linetype <-
-  links_by_year_reg %>%
-  ggplot(aes(x = completion_year, y = prop_link, color = link_type, linetype = registry)) +
-  # facet_wrap(~registry, ncol = 1) +
-  geom_point() +
-  geom_line() +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_x_continuous(breaks = unique(links_by_year$completion_year)) +
-  scale_colour_viridis_d() +
-  #TODO: show ticks for all years
-  labs(
-    y = "Registration-Publication Linkage (%)",
-    x = "Completion Year"
-  ) +
-  guides(
-    colour = guide_legend(title = "Registration-Publication Link Type"),
-    linetype = guide_legend(title = "Registry")
-    ) +
-  theme(legend.position= "bottom", legend.box="vertical", legend.margin=margin())
-
-
-# SEPARATED
-links_by_year_reg_separated <-
-  links_by_year_reg %>%
-  ggplot(aes(x = completion_year, y = prop_link, color = registry)) +
-  facet_wrap(~link_type, ncol = 1) +
-  geom_point() +
-  geom_line() +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_x_continuous(breaks = unique(links_by_year$completion_year)) +
-  scale_colour_viridis_d() +
-  #TODO: show ticks for all years
-  labs(
-    y = "Registration-Publication Linkage (%)",
-    x = "Completion Year"
-  ) +
-  guides(
[…]

### 2021-08-19 R/analysis/non-iv-links.R

@@ -1,0 +1,36 @@
+# How many (additional, non-iv) publications are listed in registrations?
+# How many (additional, non-iv) trns are in the registration, full-text, abstract, metadata?
+
+trials %>%
+  # filter(n_reg_pub_any>0) %>%
+  select(registry,
+         starts_with("n_reg_pub"),
+         starts_with("n_crossreg"
+         )) %>%
+  mutate(across(as.integer())) %>%
+  tbl_summary(
+    by = registry,
+    type = list(everything() ~ "continuous"),
+    statistic = list(everything() ~ "{median} ({min}, {max})")
+  ) %>%
+  add_overall()
+
+trials %>%
+  # count(n_crossreg_secondary_id)
+  count(n_reg_pub_any)
+  count(n_reg_pub_doi_or_pmid)
+
+trials %>%
+  filter(n_reg_pub_any==13)
+
+n_reg_pub_max_drks <-
+  trials %>%
+  filter(registry == "DRKS") %>%
+  summarise(max(n_reg_pub_any)) %>%
+  pull()
+
+n_reg_pub_max_ctgov <-
+  trials %>%
+  filter(registry == "ClinicalTrials.gov") %>%
+  summarise(max(n_reg_pub_any)) %>%
+  pull()

### 2021-08-19 R/analysis/si-if-abs-ft.R

@@ -37,4 +37,5 @@ tbl_si_ft <- tbl_si_registry(has_iv_trn_ft_pdf, "TRN in full-text")
 
 tbl_si_if_abs_ft <-
   tbl_stack(list(tbl_si_abs_or_ft, tbl_si_abs, tbl_si_ft)) %>%
-  modify_caption("Number and proportion of trials with TRN in metadata given TRN in abstract, full-text, or either. Denominator indicates number of trials with TRN in abstract and/or full-text, from the respective registry.")
+  modify_caption("Number and proportion of trials with TRN in metadata given TRN in abstract, full-text, or either. Denominator in each cell indicates number of trials with TRN in abstract, full-text, or either, overall or from the respective registry.") %>%
+  modify_spanning_header(all_stat_cols() ~ "**TRN in Metadata**")

### 2021-08-19 R/analysis/tbl-stakeholders.R

@@ -11,6 +11,7 @@ example_researchers <- "—"
 action_registries <- "
 - Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registry
 - Link publications automatically using TRNs in bibliographic database metadata
+- Provide metadata on linked publication type (i.e., result, background)
 "
 
 example_registries <- "ClinicalTrials.gov uses PubMed TRN metadata to automatically link publications in the registration  [@nationallibraryofmedicine2017c]. Our study shows this automated publication linking is responsible for the majority of linked publications in ClinicalTrials.gov and implies similar potential for DRKS."
@@ -76,9 +77,26 @@ tbl_stakeholders <-
   )
   # tab_options(table.width = px(400))
 
-# gtsave(
-#   tbl_stakeholders,
-#   here::here("docs", "figures", "tbl-stakeholders.png"),
-#   expand = 10,
-#   zoom = 3
-# )
+gtsave(
+  tbl_stakeholders,
+  here::here("docs", "figures", "tbl-stakeholders.png"),
+  expand = 10,
+  zoom = 3
+)
+
+# tbl_stakeholders <-
+#   dplyr::tribble(
+#     ~Stakeholder, ~Recommendation, ~Example,
+#     "Researchers", action_researchers, example_researchers,
+#     "Registries", action_registries, example_registries,
+#     "Research Institutions", action_institutions, example_institutions,
+#     "Publishers/Journals", action_journals, example_journals,
+#     "Bibliographic Databases (e.g., PubMed)", action_pubmed, example_pubmed,
+#   ) %>%
+#   knitr::kable(
+#     format = "markdown",
+#     caption = "Recommended stakeholder actions to improve links between trial registrations and publications"
+#   ) %>%
+#   kable_styling(latex_options = "scale_down")
+# https://community.rstudio.com/t/references-in-tables-kable-vs-gt/66780
+# https://haozhu233.github.io/kableExtra/

### 2021-08-19 R/01_get-intovalue.R
Moved dataset creation to https://github.com/maia-sh/intovalue-data

@@ -1,33 +1,0 @@
-library(dplyr)
-
-dir <- fs::dir_create(here::here("data", "raw"))
-
-# IntoValue dataset
-# Versions 1 and 2 combined and cleaned, enhanced with PMIDs/DOIs
-# NOTE: access not working so manually downloaded on 2021-07-06
-# "https://github.com/quest-bih/IntoValue2/blob/msh-dataset-checks/data/iv_main_dataset.csv"
-# "https://github.com/quest-bih/IntoValue2/blob/master/data/iv_main_dataset.csv"
-# "https://raw.githubusercontent.com/quest-bih/IntoValue2/master/data/iv_main_dataset.csv?token=AJJJ2UEAXMXNWNBD7NUBHMDA4SG2A"
-
-intovalue <-
-
-  readr::read_csv(fs::path(dir, "iv_main_dataset.csv")) %>%
-
-  # Create shorter names for publication ids
-  rename(
-    doi = publication_doi,
-    pmid = publication_pmid,
-    url = publication_url
-  )
-
-readr::write_csv(intovalue, fs::path(dir, "intovalue.csv"))
-
-# intovalue <- readr::read_csv(
-#   here::here("data", "raw", "intovalue.csv"),
-#   col_types = cols(
-#     publication_pmid = col_number(),
-#     facility_cities = col_character()
-#   )
-# )
-
-

### 2021-08-19 R/02_get-pubmed.R

@@ -1,51 +1,0 @@
-library(dplyr)
-library(fs)
-library(here)
-
-source(here("R", "functions", "download_pubmed.R"))
-
-dir <- dir_create(here("data", "raw", "pubmed-xml"))
-intovalue <- readr::read_csv(here("data", "raw", "intovalue.csv"))
-
-pmids <-
-  intovalue %>%
-  tidyr::drop_na(pmid) %>%
-  distinct(pmid) %>%
-  pull()
-
-# If pmids already downloaded, remove those from list to download
-if (dir_exists(dir)){
-
-  pmids_downloaded <-
-    dir_ls(dir) %>%
-    path_file() %>%
-    path_ext_remove() %>%
-    as.numeric()
-
-  # Check whether pmids downloaded which aren't needed and manually review and remove
-  pmids_downloaded_unused <- setdiff(pmids_downloaded, pmids)
-  if (length(pmids_downloaded_unused) > 0) {
-    rlang::warn(glue::glue("Unused pmid downloaded: {pmids_downloaded_unused}"))
-  }
-
-  pmids <- setdiff(pmids, pmids_downloaded)
-}
-
-# Download remaining pmids, if any
-if (length(pmids) > 0) {
-
-  # Use pubmed api key locally stored as "ncbi-pubmed", if available
-  # Else ask user and store
-  pubmed_api_key <-
-    ifelse(
-      nrow(keyring::key_list("ncbi-pubmed")) == 1,
-      keyring::key_get("ncbi-pubmed"),
-      keyring::key_set("ncbi-pubmed")
-    )
-
-  pmids %>%
-    purrr::walk(download_pubmed,
-                dir = dir,
-                api_key = pubmed_api_key
-    )
-}

### 2021-08-19 R/03_get-ft-pdf.R

@@ -1,104 +1,0 @@
-library(dplyr)
-library(fs)
-library(readr)
-library(here)
-
-# Full-text acquired via a combination of automated and manual techniques
-# Full-text searched for via DOI, if available, else PMID
-# Converted to XML via GROBID
-# Converted XML manually added to `dir_doi` and `dir_pmid`
-
-# PDFs
-dir_doi_raw <- here("data", "raw", "ft-pdf-doi")
-dir_pmid_raw <- here("data", "raw", "ft-pdf-pmid")
-
-# PDFs converted to XML
-dir_doi <- here("data", "raw", "ft-pdf-doi-grobid-xml")
-dir_pmid <- here("data", "raw", "ft-pdf-pmid-grobid-xml")
-
-intovalue <- readr::read_csv(here("data", "raw", "intovalue.csv"))
-
-source(here("R", "functions", "retrieve_pdf.R"))
-source(here("R", "functions", "resolve_doi.R"))
-source(here("R", "functions", "correct_pdf_url.R"))
-
-# Get pdfs from dois ------------------------------------------------------
-
-# Limit to dois for pubmed records
-dois <-
-  intovalue %>%
-  filter(!is.na(pmid) & !is.na(doi)) %>%
-  distinct(doi) %>%
-  pull()
-
-# If dois already downloaded/converted, remove those from list to download
-# Note: Currently only already converted to xml, but could be as pdf
-if (fs::dir_exists(dir_doi)){
-
-  ft_doi_retrieved <-
-    fs::dir_ls(dir_doi) %>%
-    fs::path_file() %>%
-    stringr::str_remove(".tei.xml$") %>%
-    stringr::str_replace_all("\\+", "/")
-
-
-  # Check whether unneeded dois retrieved, and manually review and remove
-  dois_downloaded_unused <- setdiff(ft_doi_retrieved, dois)
-  if (length(dois_downloaded_unused) > 0){
-    rlang::warn(glue::glue("Unused dois downloaded:{dois_downloaded_unused}"))
-  }
-
-  # Limit to missing dois
-  dois <- setdiff(dois, ft_doi_retrieved) %>% sort()
-}
-
-# Download remaining dois, if any
-if (length(dois) > 0) {
-
-  # Use email locally stored as "rm-email", if available
-  # Else ask user and store
-  email <-
-    ifelse(
-      nrow(keyring::key_list("rm-email")) == 1,
-      keyring::key_get("rm-email"),
[…]

### 2021-08-19 R/04_extract-trns.R

@@ -1,250 +1,0 @@
-library(dplyr)
-library(readr)
-library(fs)
-library(ctregistries)
-library(tidypubmed)
-
-source(here::here("R", "functions", "extract_pubmed.R"))
-source(here::here("R", "functions", "get_grobid_ft_trn.R"))
-
-intovalue <- readr::read_csv(here::here("data", "raw", "intovalue.csv"))
-
-# Get all unique pmid/doi combinations (for consistent pub ids as pmids across trns)
-pmids_dois <-
-  intovalue %>%
-  distinct(pmid, doi) %>%
-  tidyr::drop_na(pmid, doi)
-
-# Extract TRNs from: PubMed secondary identifier, PubMed abstract, and PDF full-text
-
-pm_dir <- here::here("data", "raw", "pubmed-xml")
-
-pm_xmls <-
-  pm_dir %>%
-  dir_ls()
-
-
-# Secondary identifier ----------------------------------------------------
-
-si <-
-  pm_xmls %>%
-  purrr::map_dfr(extract_pubmed, datatype = "databanks", quiet = FALSE) %>%
-  ctregistries::mutate_trn_registry(accession_number)
-
-write_rds(si, path_wd("data", "processed", "pubmed-si", ext = "rds"))
-# si <- read_rds(path_wd("data", "processed", "pubmed-si", ext = "rds"))
-
-# Visually inspect mismatching trns and registries
-# 1 trn is eudract with preceeding letters (detected, preceeding letter removed)
-# 1 trn is eudract with country code (detected, country code removed)
-# 5 trns are incorrectly formatted drks with no preceding "drks" (n = 4) or only "s" (n = 1) (not detected)
-# 3 accession numbers are figshare
-# 1 registry is attributed to genbank but drks
-# 1 registry is attributed to genbank but ct.gov
-# For now leave and later could discuss/analyze,
-# Or, combine databank/accession_number and check for trn match
-si_trn_mismatches <-
-  si %>%
-  filter(!accession_number %in% trn |
-           !databank %in% registry)
-# write_rds(si_trn_mismatches, path_wd("data", "processed", "si-trn-mismatches", ext = "rds"))
-
-si <-
[…]

### 2021-08-19 R/05_get-aact.R

@@ -1,95 +1,0 @@
-library(dplyr)
-library(dbplyr)
-library(DBI)
-library(RPostgreSQL)
-
-source(here::here("R", "functions", "aact-functions.R"))
-
-dir <- fs::dir_create(here::here("data", "raw", "ctgov"))
-
-# Get clinicaltrials.gov data via https://aact.ctti-clinicaltrials.org/
-# aact is a relational database of clinicaltrials.gov
-# some variables of interest (such as registration_date and completion_date) not in intovalue
-# also fetch additional ids and publications
-# also fetch other dates since possibly changed and for intovalue 1/2 parity
-# do no re-fetch variables that are either less likely to change or unclear how to clean like intovalue, e.g.: allocation (shouldn't change), main_sponsor (unclear source), is_multicentric (unclear source), lead_cities (unclear source and lots of intovalue munging)
-
-# Get clinicaltrial.gov trns
-ct_trns <-
-  readr::read_rds(here::here("data", "processed", "trn-all.rds")) %>%
-  filter(registry == "ClinicalTrials.gov") %>%
-  pull(trn)
-
-# ct_trns <- ct_trns[1:5]
-
-# Connect aact database
-AACT_USER <- "respmetrics"
-con <- connect_aact(AACT_USER)
-
-# Query aact database and write output to files
-studies <- query_aact(
-  "studies", ct_trns,
-  con, filepath = fs::path(dir, "studies", ext = "csv"),
-  last_update_submitted_date,
-  start_month_year, completion_month_year, primary_completion_month_year,
-  study_first_submitted_date, results_first_submitted_date,
-  study_type, phase, enrollment, overall_status
-)
-
-designs <- query_aact(
-  "designs", ct_trns,
-  con, filepath = fs::path(dir, "designs", ext = "csv"),
-  allocation, masking
-)
-
-interventions <- query_aact(
-  "interventions", ct_trns,
-  con, filepath = fs::path(dir, "interventions", ext = "csv"),
-  intervention_type
-)
-
-references <- query_aact(
-  "study_references", ct_trns,
-  con, filepath = fs::path(dir, "references", ext = "csv"),
-  pmid, reference_type, citation
-)
-
-ids <- query_aact(
-  "id_information", ct_trns,
-  con, filepath = fs::path(dir, "ids", ext = "csv"),
-  id_type, id_value
-)
-
-centers <- query_aact(
[…]

### 2021-08-19 R/06_process-aact.R

@@ -1,182 +1,0 @@
-library(dplyr)
-library(readr)
-library(fs)
-library(assertr)
-library(ctregistries)
-
-source(here::here("R", "functions", "duration_days.R"))
-
-input_dir <- here::here("data", "raw", "ctgov")
-output_dir <- dir_create(here::here("data", "processed", "ctgov"))
-
-
-# Get raw aact data -------------------------------------------------------
-
-studies <- read_csv(path(input_dir, "studies", ext = "csv"))
-designs <- read_csv(path(input_dir, "designs", ext = "csv"))
-interventions <- read_csv(path(input_dir, "interventions", ext = "csv"))
-centers <- read_csv(path(input_dir, "centers", ext = "csv"))
-references <- read_csv(path(input_dir, "references", ext = "csv"))
-ids <- read_csv(path(input_dir, "ids", ext = "csv"))
-resp_parties <- read_csv(path(input_dir, "responsible-parties", ext = "csv"))
-officials <- read_csv(path(input_dir, "officials", ext = "csv"))
-sponsors <- read_csv(path(input_dir, "sponsors", ext = "csv"))
-facilities <- read_csv(path(input_dir, "facilities", ext = "csv"))
-
-
-# Process lead and facility affiliations ----------------------------------
-
-
-# Since we don't distinguish between sponsors/official/responsible parties, combine affiliations
-# Note: For now, using intovalue umc affiliations, so simply save output
-
-resp_parties <-
-  resp_parties %>%
-  filter(!is.na(affiliation) | !is.na(organization)) %>%
-
-  # Check that each study has only affiliation OR organization, and then merge
-  assert_rows(num_row_NAs, in_set(1), c(affiliation, organization)) %>%
-
-  mutate(affiliation = coalesce(affiliation, organization), .keep = "unused") %>%
-  distinct() %>%
-  mutate(affiliation_type = "Responsible Party")
-
-officials <-
-  officials %>%
-  distinct() %>%
-  filter(!is.na(affiliation)) %>%
-  mutate(affiliation_type = "Study Official")
-
-sponsors <-
-  sponsors %>%
-  filter(lead_or_collaborator == "lead") %>%
-  select(-lead_or_collaborator) %>%
-
-  rename(
-    main_sponsor = agency_class,
-    affiliation = name
-  ) %>%
-  assert(is_uniq, nct_id) %>%
-  mutate(affiliation_type = "Sponsor")
-
-affiliations <-
[…]

### 2021-08-19 R/07_get-drks.R

@@ -1,20 +1,0 @@
-library(dplyr)
-library(purrr)
-
-source(here::here("R", "functions", "drks-functions.R"))
-
-dir <- fs::dir_create(here::here("data", "raw", "drks"))
-
-# Get drks trns
-drks_trns <-
-  readr::read_rds(here::here("data", "processed", "trn-all.rds")) %>%
-  filter(registry == "DRKS") %>%
-  pull(trn)
-
-# Download drks records
-drks_trns <-
-  readr::read_rds(here::here("data", "processed", "trn-all.rds")) %>%
-  filter(registry == "DRKS") %>%
-  pull(trn)
-
-purrr::walk(drks_trns, download_drks, dir = dir)

### 2021-08-19 R/08_process-drks.R

@@ -1,234 +1,0 @@
-library(dplyr)
-library(purrr)
-library(fs)
-library(readr)
-library(stringr)
-library(ctregistries)
-
-source(here::here("R", "functions", "drks-functions.R"))
-source(here::here("R", "functions", "duration_days.R"))
-
-input_dir <- here::here("data", "raw", "drks")
-output_dir <- dir_create(here::here("data", "processed", "drks"))
-
-# Parse all records in directory, alternatively could use drks_trns
-drks_htmls <-
-  dir_ls(input_dir)
-
-# drks_htmls <- drks_htmls[1:50]
-
-# Parse secondary ids -----------------------------------------------------
-
-drks_ids <-
-  drks_htmls %>%
-  map_dfr(parse_drks_ids) %>%
-  ctregistries::mutate_trn_registry(id) %>%
-
-  # Several ISRCTN are misformatted so capture here
-  mutate(
-
-    trn = if_else(
-      str_detect(id, "\\d{8}[[:blank:][:punct:]]*ISRCTN"),
-      str_c("ISRCTN", str_extract(id, "\\d{8}")),
-      trn
-    ),
-
-    registry = if_else(
-      str_detect(id, "\\d{8}[[:blank:][:punct:]]*ISRCTN"),
-      "ISRCTN",
-      registry
-    )
-  ) %>%
-
-  # Clean trns and collapse EudraCT entries
-  mutate(
-    raw_trn = trn,
-    trn = purrr::map_chr(raw_trn, ctregistries::clean_trn),
-    id_type = if_else(stringr::str_detect(id_type, "EudraCT"),"EudraCT", id_type)
-  )
-
-write_rds(drks_ids, path(output_dir, "drks-ids", ext = "rds"))
-
-drks_crossreg <-
-  drks_ids %>%
-  filter(!is.na(trn)) %>%
-  select(drks_id, crossreg_registry = registry, crossreg_trn = trn)
-
-write_rds(drks_crossreg, path(output_dir, "drks-crossreg", ext = "rds"))
-
-# Parse references --------------------------------------------------------
-
-drks_references <-
-  drks_htmls %>%
-
-  map_dfr(parse_drks_references) %>%
-
-  # Extract dois and pmids, from both `citation` and `link`
-  # Note: regexes adapted for edge cases as described in comments
-  mutate(
-
-    # Some (e.g., DRKS00004841) have >1 doi --> extract all and unnest)
-    # One (i.e., DRKS00000005) has has extra space in doi --> extract and clean
[…]

### 2021-08-19 R/09_combine-ctgov-drks-trials.R

@@ -1,100 +1,0 @@
-library(dplyr)
-library(readr)
-library(fs)
-library(lubridate)
-
-source(here::here("R", "functions", "duration_days.R"))
-
-# Get data ----------------------------------------------------------------
-
-ctgov_dir <- here::here("data", "processed", "ctgov")
-drks_dir <- here::here("data", "processed", "drks")
-
-ctgov_studies <- read_rds(path(ctgov_dir, "ctgov-studies.rds"))
-ctgov_references <- read_rds(path(ctgov_dir, "ctgov-references.rds"))
-
-drks_studies <- read_rds(path(drks_dir, "drks-studies.rds"))
-drks_references <- read_rds(path(drks_dir, "drks-references.rds"))
-
-
-# Combine registry studies ------------------------------------------------
-
-ctgov_studies <-
-  ctgov_studies %>%
-  rename(id = nct_id) %>%
-  select(-last_update_submitted_date) %>%
-  mutate(registry = "ClinicalTrials.gov")
-
-drks_studies <-
-  drks_studies %>%
-  rename(id = drks_id) %>%
-  mutate(registry = "DRKS")
-
-# Check registry column names:
-# Expect none in drks not in ctgov
-# Expect pcd and summary results date in ctgov and not in drks
-if (
-  !rlang::is_empty(setdiff(colnames(drks_studies), colnames(ctgov_studies))) |
-  rlang::is_empty(setdiff(colnames(ctgov_studies), colnames(drks_studies))) |
-  !all(setdiff(colnames(ctgov_studies), colnames(drks_studies)) %in% c("summary_results_date", "primary_completion_date"))
-) {rlang::warn("There are unexpected drks and/or ctgov column names!")}
-
-registry_studies <-
-  bind_rows(ctgov_studies, drks_studies) %>%
-
-
-  mutate(
-
-    # Trials from registries are resolved
-    is_resolved = TRUE,
-
-    # Trials are randomized if allocation includes randomized
-    # Otherwise, not randomized (unless no allocation, then NA)
-    is_randomized =
-      if_else(stringr::str_detect(allocation, "(?i)(?<!non-)randomized"), TRUE, FALSE),
-
-    # Registration is prospective if registered in same or prior month to start
-    is_prospective =
-      (floor_date(registration_date, unit = "month") <=
-      floor_date(start_date, unit = "month")),
-
-    # Update registry date variables to match intovalue
-    completion_year = year(completion_date),
-    primary_completion_year = year(primary_completion_date),
[…]

### 2021-08-19 R/10_prepare-cross-registrations.R

@@ -1,174 +1,0 @@
-library(dplyr)
-library(readr)
-library(here)
-
-
-# Get data ----------------------------------------------------------------
-
-intovalue <- read_csv(here("data", "raw", "intovalue.csv"))
-ctgov_crossreg <- read_rds(here("data", "processed", "ctgov", "ctgov-crossreg.rds"))
-drks_crossreg <- read_rds(here("data", "processed", "drks", "drks-crossreg.rds"))
-# registry_crossreg <- read_rds(here("data", "processed", "registry-crossreg.rds"))
-trn_reported_wide <- read_rds(here("data", "processed", "trn-reported-wide.rds"))
-trn_reported_long <- read_rds(here("data", "processed", "trn-reported-long.rds"))
-pubmed_ft_pdf_retrieved <- read_rds(here("data", "processed", "pubmed-ft-pdf-retrieved.rds"))
-
-# Combine registry cross-registrations ------------------------------------
-
-registry_crossreg <-
-  bind_rows(
-    rename(ctgov_crossreg, id = nct_id),
-    rename(drks_crossreg, id = drks_id)
-  ) %>%
-  group_by(id) %>%
-  mutate(n_crossreg = row_number()) %>%
-  ungroup()
-
-write_rds(registry_crossreg, here::here("data", "processed", "registry-crossreg.rds"))
-
-# There are up to 4 cross-registrations
-# max(registry_crossreg$n_crossreg)
-
-# To join with trials, could pivot wider
-# registry_crossreg %>%
-#   tidyr::pivot_wider(
-#     id_cols = id,
-#     names_from = n_crossreg,
-#     values_from = c(crossreg_registry, crossreg_trn)
-#   )
-
-
-# Prepare all cross-registrations -----------------------------------------
-
-# Cross-registrations per pubmed/publication
-crossreg_pub <-
-
-  # Remove intovalue primary registrations (since already included in `trials`)
-  anti_join(trn_reported_wide, intovalue, by = c("trn" = "id", "pmid")) %>%
-  rename(crossreg_registry = registry, crossreg_trn = trn) %>%
-  rename_with(~stringr::str_replace(., "has_trn", "is_crossreg")) %>%
-
-  # Add in intovalue primary trn
-  # Also cause duplication of some publications associated with multiple trials
[…]

### 2021-08-19 R/11_prepare-trials.R

@@ -1,273 +1,0 @@
-library(dplyr)
-library(tidyr)
-library(readr)
-library(here)
-library(fs)
-library(lubridate)
-
-source(here("R", "functions", "duration_days.R"))
-
-# Get data ----------------------------------------------------------------
-
-intovalue <- read_csv(here("data", "raw", "intovalue.csv"))
-registry_studies <- read_rds(here("data", "processed", "registry-studies.rds"))
-registry_references <- read_rds(here("data", "processed", "registry-references.rds"))
-n_cross_registrations <- read_rds(here("data", "processed", "n-cross-registrations.rds"))
-trn_reported_wide <- read_rds(here("data", "processed", "trn-reported-wide.rds"))
-pubmed_ft_pdf_retrieved <- read_rds(here("data", "processed", "pubmed-ft-pdf-retrieved.rds"))
-
-
-# Prepare trials ----------------------------------------------------------
-
-# Prepare columns to keep from intovalue
-# Exclude "days_DATE_to_publication" columns since will recalculate from registries
-iv_cols_to_keep <-
-  setdiff(colnames(intovalue), colnames(registry_studies)) %>%
-  setdiff(stringr::str_subset(., "^days_")) %>%
-  c("id", .)
-
-trials <-
-
-  # Select intovalue columns to keep
-  intovalue %>%
-
-  select(all_of(iv_cols_to_keep)) %>%
-
-  # Add updated registry data
-  left_join(registry_studies, by = "id") %>%
-
-  mutate(
-
-    days_cd_to_publication = duration_days(completion_date, publication_date),
-    days_pcd_to_publication = duration_days(primary_completion_date, publication_date),
-    days_reg_to_publication = duration_days(registration_date, publication_date),
-
-    # Note: some (n = 3) trials !has_publication do have pmids (abstract only)
-    has_pmid = if_else(!is.na(pmid), TRUE, FALSE)
-  ) %>%
-
-  # Add info about pubmed and ft (pdf) retrieval
-  left_join(pubmed_ft_pdf_retrieved, by = c("id", "doi", "pmid")) %>%
-
-  # Add info about whether intovalue trn in secondary id, abstract, ft pdf
-  left_join(trn_reported_wide, by = c("id" = "trn", "registry", "doi", "pmid")) %>%
-
[…]

### 2021-08-19 R/12_validate-trials.R

@@ -1,89 +1,0 @@
-library(dplyr)
-library(assertr)
-library(assertive)
-library(pointblank)
-
-trials <- readr::read_rds(here::here("data", "processed", "trials.rds"))
-intovalue <- readr::read_csv(here::here("data", "raw", "intovalue.csv"))
-
-# Check that all intovalue columns in trials
-if (!rlang::is_empty(setdiff(colnames(intovalue), colnames(trials)))) {
-  rlang::warn("There are intovalue columns missing from trials!")
-}
-
-trials %>%
-
-  # Validate publication dois
-  pointblank::col_vals_regex(
-    columns = vars(doi),
-    regex = "^10\\.\\d{4,9}/[-.;()/:_[:alnum:]]+$",
-    na_pass = TRUE
-  ) %>%
-
-  # Validate publication pmids
-  pointblank::col_vals_regex(
-    columns = vars(pmid),
-    regex = "^[0-9]{8}$",
-    na_pass = TRUE
-  ) %>%
-
-  # Validate publication urls
-  pointblank::col_vals_regex(
-    columns = vars(url),
-    regex = "^http",
-    na_pass = TRUE
-  ) %>%
-
-  # Publication urls should not go to registries
-  assertr::assert(
-    function(url) !stringr::str_detect(url, "clinicaltrials.gov|drks.de")|is.na(url),
-    url
-  ) %>%
-
-  # Check that all trials with publication (including abstract) has pmid, doi, OR url
-  assertr::verify(nrow(filter(., identification_step != "No publ" & is.na(url) & is.na(doi) & is.na(pmid))) == 0) %>%
-
-  # Check that if identification_step is no publication or abstract only, not has_publication
-  pointblank::col_vals_equal(
-    vars(has_publication),
-    value = FALSE,
-    preconditions = ~ . %>% filter(identification_step %in% c("No publ", "Abstract only"))
-  ) %>%
-
-  # Check that if not has_publication, no publication date and identification_step is no publication or abstract only
-  pointblank::conjointly(
-    ~ col_vals_in_set(.,
-                      columns = vars(identification_step),
-                      set = c("No publ", "Abstract only")
-    ),
-    ~ col_vals_null(., vars(publication_date)),
-    preconditions = ~ . %>% filter(!has_publication)
-  ) %>%
-
-  # Check that if no publication, no publication ids
-  pointblank::col_vals_null(
-    vars(url, doi, pmid),
[…]

### 2021-08-19 R/functions/aact-functions.R

@@ -1,70 +1,0 @@
-# Get clinicaltrials.gov data via https://aact.ctti-clinicaltrials.org/
-# AACT is a relational database of clinicaltrials.gov
-
-# You must first create an account with AACT: https://aact.ctti-clinicaltrials.org/users/sign_up
-
-#' Connect to AACT database
-#'
-#' Based on https://aact.ctti-clinicaltrials.org/r.
-#' Use account credentials created via https://aact.ctti-clinicaltrials.org/users/sign_up
-#'
-#' @param user Character. AACT username.
-#' @param password Character. Default is NULL. If not provided, password will be searched for in \code{keyring} under the \code{aact} service with the provided \code{username}. If no password found, user will be interactively asked and input will be stored in keyring.
-#'
-#' @return <PostgreSQLConnection> object from \code{dbConnect}
-#'
-
-connect_aact <- function(user, password = NULL) {
-  password <-
-    ifelse(
-      rlang::is_true(stringr::str_detect(keyring::key_list("aact")$username, user)),
-      keyring::key_get("aact", user),
-      keyring::key_set("aact", user)
-    )
-
-  drv <- DBI::dbDriver('PostgreSQL')
-  con <- RPostgreSQL::dbConnect(drv,
-                   dbname = "aact",
-                   host = "aact-db.ctti-clinicaltrials.org",
-                   port = 5432,
-                   user = user,
-                   password = password
-  )
-
-  con
-}
-
-
-#' Query AACT table
-#'
-#' AACT documentation available in the [database schema](https://aact.ctti-clinicaltrials.org/schema) and [data dictionary](https://aact.ctti-clinicaltrials.org/data_dictionary)
-#'
-#' @param table Character. AACT table.
-#' @param ids Character. Vector of NCT ids to query.
-#' @param con Connection to AACT database as returned by \code{dbConnect}.
-#' @param filepath Character. Filepath including directory portion and ending in ".csv". If no filepath provided, output will be only returned and not written to file.
-#' @param ... One or more unquoted expressions separated by commas. Column names from AACT table. `nct_id` is always returned and should not be included in this.
-#'
-#' @return Dataframe with specified columns from \code{table}
-#'
-
-query_aact <- function(table, ids, con, filepath = NULL, ...){
-
-  query <-
-    dplyr::tbl(con, table) %>%
-    dplyr::filter(nct_id %in% ids) %>%
[…]

### 2021-08-19 R/functions/correct_pdf_url.R

@@ -1,66 +1,0 @@
-#' Correct PDF urls based on publisher
-#'
-#' @param url Character. A publication URL. Usually generated from \code{resolve_doi}.
-#'
-#' @return A URL
-#' @export
-#'
-correct_pdf_url <- function(url) {
-
-  parsed <- httr::parse_url(url)
-
-  # Determine publisher
-  publisher <-  stringr::str_extract(parsed$hostname, c(
-    "onlinelibrary.wiley.com",
-    "ascopubs.org",
-    "linkinghub.elsevier.com"
-  )) %>%
-    purrr::discard(is.na)
-
-  # If there was more than 1 publisher match, treat as "other" and flag in info
-  # Unspecified publishers are categorized as "other"
-  if (length(publisher) > 1) {
-    info <- "There seems to be multiple publishers"
-    publisher <- "other"
-  } else if (rlang::is_empty(publisher)) {
-    publisher <- "other"
-  }
-
-  parsed <-
-    switch(publisher,
-           "onlinelibrary.wiley.com" = build_wiley(parsed),
-           "ascopubs.org" = build_asco(parsed),
-           "linkinghub.elsevier.com" = build_elsevier(parsed),
-           "other" = parsed
-    )
-
-  if (publisher != "other"){
-    url <- httr::build_url(parsed)
-  }
-
-  url
-
-}
-
-build_elsevier <- function(parsed){
-  parsed$hostname <- "www.sciencedirect.com"
-  parsed$path <-
-    stringr::str_c(
-      "science/article",
-      stringr::str_extract(parsed$path, "/pii/.+$"),
-      "/pdfft"
-    )
-  parsed$query <- list(isDTMRedir = "true", download = "true")
-
-  parsed
-}
-
-build_wiley <- function(parsed){
-  parsed$path <- stringr::str_replace(parsed$path, "full", "pdfdirect")
-  parsed
-}
-
-build_asco <- function(parsed){
-  parsed$path <- stringr::str_replace(parsed$path, "doi", "doi/pdfdirect")
-  parsed
-}

### 2021-08-19 R/functions/download_pubmed.R

@@ -1,99 +1,0 @@
-#' Download PubMed record
-#'
-#' This is a wrapper around rentrez::entrez_fetch() that allows to purrr::safely() fetch and save a PubMed record.
-#'
-#' This function is *not* vectorized so wrap in `purrr::walk()` to run on multiple pmids.
-#'
-#' @param pmid 	character, publication identifier. PMID or PMCID.
-#' @param dir character, director in which to save XMLs. Defaults to working directory with `here::here()`.
-#' @param overwrite logical, if file exists, should it be overwritten? Defaults to `FALSE`.
-#' @param api_key character. See {rentrez} documentation on [Rate-limiting and API Keys](https://cran.r-project.org/web/packages/rentrez/vignettes/rentrez_tutorial.html#rate-limiting-and-api-keys) for info on \code(api_key). Defaults to NULL.
-#'
-#'@param sleep delay in seconds. Defaults to NULL. If \code(api_key) provided, \code(sleep) set to 1\10 seconds, else to 1\3 seconds per [Entrez API limits](https://ncbiinsights.ncbi.nlm.nih.gov/2017/11/02/new-api-keys-for-the-e-utilities/).
-#' @return If \code{pmid} successfully downloaded, returns XML. If \code{pmid} did not resolve, or already downloaded in \code{dir} (and \code{overwrite} is FALSE), returns NULL.
-#'
-#' @export
-#'
-#' @examples
-#' \dontrun{
-#' download_pubmed(99999999,
-#' dir = here("data", "raw", "pubmed-xml"),
-#' api_key = keyring::key_get("ncbi-pubmed")
-#' )
-#' }
-#' \dontrun{
-#' c(99999999, 12345678) %>%
-#'   walk(download_pubmed,
-#'        dir = here("data", "raw", "pubmed-xml"),
-#'        api_key = keyring::key_get("ncbi-pubmed")
-#'   )
-#' }
-
-download_pubmed <- function(pmid,
-                            dir = here::here(),
-                            overwrite = FALSE,
-                            api_key = NULL,
-                            sleep = NULL){
-
-  # Check input
-  if (!stringr::str_detect(pmid, "[0-9]{8}")) {
-    rlang::abort(glue::glue("{pmid} is not a valid pmid"))
-  }
-
-  # Create directory, if it doesn't exist
-  fs::dir_create(dir)
-
-  # Construct filepath
[…]

### 2021-08-19 R/functions/drks-functions.R

@@ -1,481 +1,0 @@
-#' Get DRKS record
-#'
-#' Fetch DRKS record based on \code{drks_id} and check that it resolved to a valid, matching record.
-#'
-#' @param drks_id
-#' @param quiet
-#'
-#' @return DRKS record as html_document, if valid, otherwise, NULL
-#'
-#' @examples
-#' fetch_drks(drks_id)
-
-fetch_drks <- function(drks_id, quiet = FALSE){
-
-  drks_url <- "https://www.drks.de/drks_web/navigate.do?navigationId=trial.HTML&TRIAL_ID="
-
-  doc <-
-    rvest::read_html(glue::glue("{drks_url}{drks_id}"))
-
-  doc_id <- get_drks_element(doc, ".drks_id p")
-
-  # If drks_id does not resolve to valid record with same id, inform and return early
-  if (rlang::is_empty(doc_id) || doc_id != drks_id) {
-    if (!quiet){rlang::inform(glue::glue("{drks_id}: invalid drks id"))}
-    return(NULL)
-  }
-
-  if (!quiet){rlang::inform(glue::glue("{drks_id}: fetched"))}
-  return(doc)
-
-}
-
-#' Download DRKS
-#'
-#' Save a local copy of DRKS record. Calls \code{fetch_drks()} to fetch DRKS record, if needed.
-#'
-#' @param drks_id
-#' @param quiet
-#' @param overwrite
-#' @param quiet
-#'
-#' @return DRKS record as html_document, if valid, otherwise, NULL
-#'
-#' @examples
-#' \dontrun{download_drks(drks_id, dir, overwrite = TRUE)}
-
-download_drks <- function(drks_id,
-                          dir = here::here(),
-                          overwrite = FALSE,
-                          quiet = FALSE){
-
-  drks_url <- "https://www.drks.de/drks_web/navigate.do?navigationId=trial.HTML&TRIAL_ID="
-
-  # Create directory (if doesn't exist) and filepath
-  dir_files <-
-    fs::dir_create(dir) %>%
-    fs::dir_ls()
-
-  filepath <- glue::glue("{dir}/{drks_id}")
-
-  # If record already downloaded and no rewrite, inform and return early
-  if (filepath %in% dir_files & !overwrite){
-    if (!quiet){rlang::inform(glue::glue("{drks_id}: already downloaded, not overwriting"))}
-    doc <- rvest::read_html(filepath)
-    return(doc)
-  }
[…]

### 2021-08-19 R/functions/duration_days.R

@@ -1,4 +1,0 @@
-# Helper function to calculate duration in days
-duration_days <- function(start, end){
-  as.numeric(lubridate::as.duration(lubridate::interval(start, end)), "days")
-}

### 2021-08-19 R/functions/extract_pubmed.R

@@ -1,67 +1,0 @@
-#' Extract dataypes from PubMed XML records
-#'
-#' @param filepath Character. Filepath including directory portion. Filepath should end in ".xml". If \code{pmid} is not provided, filename should be PMID.
-#' @param pmid Character. Default = NULL in which case \code{pmid} parsed from the filename.
-#' @param datatype Type of data to extract from xml for which there is a corresponding "pubmed_" function ("main", "abstract", "databanks", "authors", "mesh", "keywords", "pubtypes"). "main" used `pubmed_table()`
-#' @param processed_dir Character. Default = NULL. Filepath of a directory to which processed file should be moved. By default, file will not be moved.
-#' @param quiet Logical. Default = FALSE. Whether to inform user.
-#'
-#' @return Dataframe with output from corresponding "pubmed_" function
-#'
-
-extract_pubmed <- function(filepath,
-                           datatype = NULL,
-                           pmid = NULL,
-                           processed_dir = NULL,
-                           quiet = FALSE){
-
-
-  # Check for correct filetype
-  if (fs::path_ext(filepath) != "xml"){
-    rlang::abort("`filepath` must end in '.xml'")
-  }
-
-  datatypes <- c("main", "abstract", "databanks", "authors", "mesh", "keywords", "pubtypes")
-
-  datatypes_txt <-
-    glue::glue_collapse(glue::glue("'{datatypes}'"), sep = ", ", last = ", or ")
-
-  # Check for valid datatype
-  if (!datatype %in% datatypes){
-    rlang::abort(glue::glue("`datatype` must be in {datatypes_txt}"))
-  }
-
-  # Convert "main" to table" (use "main" for user since clearer than "table")
-  function_datatype <- ifelse(datatype == "main", "table", datatype)
-
-  # Extract id from filepath if not user-provided
-  if (purrr::is_null(pmid)){
-    pmid <-
-      filepath %>%
-      fs::path_file() %>%
-      fs::path_ext_remove()
-  }
-  pmid = as.numeric(pmid)
-
-  article <-
-    filepath %>%
-    xml2::read_xml() %>%
-    xml2::xml_find_all("PubmedArticle")
-
-  # Move file if user specified
-  if (!rlang::is_null(processed_dir)) {
-    fs::dir_create(processed_dir)
-    fs::file_move(filepath, processed_dir)
-  }
-
-
-  out <-
[…]

### 2021-08-19 R/functions/get_grobid_ft_trn.R

@@ -1,182 +1,0 @@
-#' Extract TRNs from full text publications.
-#'
-#' Publications should have been parsed by GROBID from PDF to TEI XML.
-#'
-#' @param filepath Character. Filepath including directory portion. Filepath should end in ".tei.xml" extensions as generated by GROBID. If \code{doi} or \code{pmid} is not provided, filename should be either pmid or doi with "/" replaced by "+".
-#' @param doi Character. Default = NULL in which case \code{doi} parsed from the filename. Provide only \code{doi} or \code{pmid}.
-#' @param pmid Character or numeric. Default = NULL in which case \code{pmid} parsed from the filename. Provide only \code{doi} or \code{pmid}.
-#' @param where Character. Default = "body" What part of the GROBID XML should be searched for TRNs? This paramater currently supports only "body" but may be further developed to allow to other sections or even arbitrary xpaths.
-#' @param processed_dir Character. Default = NULL. Filepath of a directory to which processed file should be moved. By default, file will not be moved.
-#' @param dedupe Logical. Default = TRUE. Whether to deduplicate repeated TRNs from the same \code{where}. Two instances of the same NCT from within "body" would be reduced to one, but two instances of the same NCT from different sections (if this becomes an option) would be maintained.
-#' @param quiet Logical. Default = FALSE. Whether to inform user about whether TRNs found for each DOI.
-#'
-#' @return Tibble with 6 columns: `doi`, `pmid`, `n` (trn count from `source`), `trn`, `registry` (generated from TRN, not extracted from full-text), `body`
-#'
-#' @examples
-#' \dontrun{get_grobid_ft_trn(dir_ls(here::here("data-raw", "fulltext-xml"))[1])}
-#' \dontrun{purrr::map_dfr(dir_ls(here::here("data-raw", "fulltext-xml")), get_grobid_ft_trn}
-
-get_grobid_ft_trn <- function(filepath,
-                       doi = NULL,
-                       pmid = NULL,
-                       where = "body",
-                       processed_dir = NULL,
-                       dedupe = TRUE,
-                       quiet = FALSE) {
[…]

### 2021-08-19 R/functions/report-results.R

@@ -1,17 +1,0 @@
-# # Report the % (num/denom) of links in a given year by type ---------------
-#
-# report_link_year <- function(link_type_in, year) {
-#   links_by_year %>%
-#     filter(link_type == link_type_in & completion_year == year) %>%
-#     mutate(report = glue("{style_percent(prop_link, symbol =  TRUE)} ({n_link}/{n_link_type})")) %>%
-#     pull(report)
-# }
-#
-# report_link_year_reg <- function(link_type_in, year, registry_in) {
-#   links_by_year_reg %>%
-#     filter(
-#       link_type == link_type_in & completion_year == year & registry == registry_in
-#     ) %>%
-#     mutate(report = glue("{style_percent(prop_link, symbol =  TRUE)} ({n_link}/{n_link_type})")) %>%
-#     pull(report)
-# }

### 2021-08-19 R/functions/resolve_doi.R

@@ -1,52 +1,0 @@
-#' Resolve a DOI using doi.org
-#'
-#' Resolve a DOI to a URL via [doi.org](https://doi.org/). This is slower than using the [REST API](https://www.doi.org/factsheets/DOIProxy.html#rest-api), however returns the final, redirected URL.
-#'
-#' @param doi Character. A publication doi with or without a preceding URL.
-#'
-#'#' @return S3 Class object \code{doi} with \code{doi} and \code{url}, as well as \code{success} (binary) and \code{error} (details on error if not \code{success}), and raw \code{response}.
-#'
-#' @export
-#'
-#' @examples
-#' \dontrun{
-#' resolve_doi("10.1002/bjs.9387")
-#' }
-#'
-resolve_doi <- function(doi) {
-
-  # Remove scheme and hostname, if supplied
-  doi <- httr::parse_url(doi)$path
-
-  resp <- httr::HEAD("https://doi.org", path = doi)
-
-  # Check for errors (status code)
-  if (httr::http_error(resp)){
-    error <-  resp$status_code
-    success <-  FALSE
-    url <-  NA
-  } else {
-    # If no errors, get url
-    url <- resp$url
-    success <- TRUE
-    error <- NA
-  }
-
-  structure(
-    list(
-      doi = doi,
-      url = url,
-      success = success,
-      error = error,
-      response = resp
-    ),
-    class = "doi"
-  )
-}
-
-print.doi <- function(x, ...) {
-  cat("<DOI ", x$doi, ">\n", sep = "")
-  if (x$success) str(x$url)
-  if (!x$success) cat("Error:", x$error)
-  invisible(x)
-}

### 2021-08-19 R/functions/retrieve_pdf.R

@@ -1,155 +1,0 @@
-# doi <- dois_missing[1]
-# doi <- "10.1371/journal.pone.0199776"
-# doi <- "doi"
-# dir <- here("data", "raw", "ft-pdf-iv2")
-
-# retrieve_pdf(doi, dir)
-
-retrieve_pdf <- function(doi,
-                         email,
-                         dir = here::here(),
-                         overwrite = FALSE,
-                         source = "unpaywall",
-                         sleep = 1/3){
-
-  # Check for valid DOI
-  if (!stringr::str_detect(doi, "^10\\.\\d{4,9}/[-.;()/:\\w]+$")){
-    rlang::abort(glue::glue("`doi` must be a properly formated doi, not {doi}"))
-  }
-
-  # Check for valid source(s)
-  sources <- c("unpaywall", "publisher", "magic")
-
-  sources_txt <-
-    glue::glue_collapse(glue::glue("'{sources}'"), sep = ", ", last = ", or ")
-
-  if (!all(source %in% sources)){
-    rlang::abort(glue::glue("Each `source` must be in {sources_txt}"))
-  }
-
-  # Create directory, if it doesn't exist
-  fs::dir_create(dir)
-
-  # Construct filepath
-  filepath <- fs::path(dir, stringr::str_replace_all(doi, "/", "\\+"), ext = "pdf")
-
-  # If file exists and not overwriting, don't download, and return NULL
-  if (!overwrite && fs::file_exists(filepath)){
-    rlang::inform(glue::glue("{doi}: Already downloaded"))
-    return(NULL)
-  }
-
-  Sys.sleep(sleep)
-
-  # Try unpaywall
-  if ("unpaywall" %in% source){
-
-    oa_url <- get_pdf_url_unpaywall(doi, email)
-
-    # If link, download and check valid pdf, and if good, then return
-    if(!is.na(oa_url)){
-      download.file(oa_url, filepath, mode = "wb", quiet = TRUE)
-
-      if(check_exists_length(filepath) > 0){
-        rlang::inform(glue::glue("{doi}: Successfully downloaded via unpaywall"))
-        return()
-      } else (fs::file_delete(filepath))
-    }
-  }
-
-  # Try publisher (via doi.org)
-  if ("publisher" %in% source){
-    doi_url <- resolve_doi(doi)$url
-
-    if(!is.na(doi_url)){
-
-      # Edit pdf link per manual rules
[…]

### 2021-08-19 R/functions/screening.R

@@ -1,39 +1,0 @@
-# Apply screening criteria ------------------------------------------------
-# Returns list of inclusion counts and filtered dataframe
-
-count_filter <- function(data, vars) {
-
-  counts <-
-    tibble(name = as.character(),
-           value = as.logical(),
-           n = as.integer()
-    )
-
-  for (var in vars) {
-    counts <-
-      data %>%
-      count(.data[[var]]) %>%
-      pivot_longer(-n) %>%
-      add_row(counts, .)
-
-    data <- filter(data, .data[[var]])
-  }
-
-  list(data = data, counts = counts)
-
-}
-
-
-# Report screening summary counts -----------------------------------------
-
-report_n <- function(counts, var, condition) {
-  n <-
-    counts %>%
-    filter(name == var & value == condition) %>%
-    pull(n)
-
-  # If empty, count is 0
-  if (rlang::is_empty(n)){n <- 0}
-
-  n
-}

### 2021-08-21 R/analysis/01_screen-trials.R
Remove earlier version of analysis and manuscript

@@ -1,96 +1,0 @@
-trials_all <-
-  trials_all %>%
-
-  # Create flag for dupes in IV1 for exclusion
-  mutate(is_not_iv1_dupe = if_else(!(is_dupe & iv_version == 1), TRUE, FALSE))
-
-# Screen trials -----------------------------------------------------------
-
-# IntoValue 1 and 2
-iv12_counts <-
-  trials_all %>%
-  count(iv_version) %>%
-  mutate(iv_version = str_c("iv", iv_version)) %>%
-  rename(name = iv_version) %>%
-  mutate(value =  TRUE, .after = "name")
-
-# IntoValue 1 and 2 German UMC leads
-# iv12_lead_counts <-
-#   trials_all %>%
-#   count(iv_version, has_german_umc_lead) %>%
-#   mutate(iv_version = str_c("has_german_umc_lead_iv", iv_version)) %>%
-#   rename(name = iv_version, value = has_german_umc_lead)
-
-screening_criteria <- c(
-  "iv_completion",
-  "iv_status",
-  "iv_interventional",
-  "has_german_umc_lead",
-  "is_not_iv1_dupe",
-  "has_publication",
-  "has_pmid",
-  "has_pubmed",
-  "has_ft_pdf"
-)
-
-trials_screened <-
-  trials_all %>%
-
-  # Remove doi, to be replace with more complete dois from publications
-  select(-doi) %>%
-
-  # is_dupe needs to be is_not_dupe, since screening keeps TRUE
-  mutate(is_not_dupe_trial = if_else(is_dupe, FALSE, TRUE, missing = TRUE)) %>%
-
-  count_filter(screening_criteria)
-
-trials <- trials_screened$data
-
-# Tabularize trial screening counts
-
-trial_screening <-
-  bind_rows(
-    iv12_counts,
-    # iv12_lead_counts,
-    trials_screened$counts
-  ) %>%
-  add_row(
-    name = "is_unique_pmid",
-    value = c(TRUE, FALSE),
-    n = c(n_distinct(trials$pmid),
-          nrow(trials) - n_distinct(trials$pmid))
-  )
-
-# Report trial screening counts
-
-n_trials_iv1 <- report_n(trial_screening, "iv1", TRUE) #n_iv1
-n_trials_iv2 <- report_n(trial_screening, "iv2", TRUE) #n_iv2
[…]

### 2021-08-21 R/analysis/02_flowchart-trials.R

@@ -1,110 +1,0 @@
-# Prepare labels
-label_iv_all <- glue('IntoValue trials\n(n = {n_trials_iv})\nIV1 = {n_trials_iv1} & IV2 = {n_trials_iv2}')
-label_iv_screened <- glue('Trials meeting IntoValue criteria\nwith new registry data\n(n = {n_trials_iv_interventional})')
-label_trials_lead <- glue('IntoValue German UMC Lead Trials\n(n = {n_trials_lead})')
-label_trials_deduped <- glue('Trials with duplicates removed\n(n = {n_trials_deduped})')
-label_pub <- glue('Trials with published results\n(n = {n_trials_pub})')
-label_pmid <- glue('Trials with published results\nwith PMID\n(n = {n_trials_pmid})')
-# label_pubmed <- glue('Trials with published results\nwith resolved PubMed record\n(n = {n_trials_pubmed})')
-label_fulltext <- glue('Trials with published results\nas full-text (pdf)\n(n = {n_trials_ft_pdf})')
-
-label_iv_screened_ex <- glue('Trials not meeting IntoValue criteria\nwith new registry data\n(n = {n_trials_iv_completion_date_ex + n_trials_iv_status_ex + n_trials_iv_interventional_ex})\nNot completed 2009-2017 (n = {n_trials_iv_completion_date_ex})\nNot included status (n = {n_trials_iv_status_ex})\nNot interventional (n = {n_trials_iv_interventional_ex})')
-label_iv_lead_ex <- glue('No German UMC Lead\n(n = {n_trials_lead_ex})')
-label_trials_dupes_ex <- glue('Duplicate trials\n(n = {n_trials_dupes_ex})')
-label_pub_ex <- glue('No published results\n(n = {n_trials_pub_ex})')
-label_pmid_ex <- glue('No PMID\n(n = {n_trials_pmid_ex})')
-# label_pubmed_ex <- glue('PMID unresolved in PubMed\n(n = {n_trials_pubmed_ex})')
-label_fulltext_ex <- glue('No full-text (pdf) found\n(n = {n_trials_ft_pdf_ex})')
-
-# Prepare flowchart
-flow_trials <- DiagrammeR::grViz("digraph trials {
-
-# GRAPH
-graph [layout = dot, rankdir = LR, splines = false]
[…]

### 2021-08-21 R/analysis/02_flowchart-trials_dep.R

@@ -1,116 +1,0 @@
-# Prepare labels
-label_iv_all <- glue('IntoValue trials\n(n = {n_trials_iv})\nIV1 = {n_trials_iv1} & IV2 = {n_trials_iv2}')
-label_iv_screened <- glue('Trials meeting IntoValue criteria\nwith new registry data\n(n = {n_trials_iv_interventional})')
-label_trials_lead <- glue('IntoValue German UMC Lead Trials\n(n = {n_trials_lead})')
-label_trials_deduped <- glue('Trials with duplicates removed\n(n = {n_trials_deduped})')
-label_pub <- glue('Trials with published results\n(n = {n_trials_pub})')
-label_pmid <- glue('Trials with published results\nwith PMID\n(n = {n_trials_pmid})')
-label_pubmed <- glue('Trials with published results\nwith resolved PubMed record\n(n = {n_trials_pubmed})')
-label_fulltext <- glue('Trials with published results\nas full-text (pdf)\n(n = {n_trials_ft_pdf})')
-
-label_iv_screened_ex <- glue('Trials not meeting IntoValue criteria\nwith new registry data\n(n = {n_trials_iv_completion_date_ex + n_trials_iv_status_ex + n_trials_iv_interventional_ex})\nNot completed 2009-2017 (n = {n_trials_iv_completion_date_ex})\nNot included status (n = {n_trials_iv_status_ex})\nNot interventional (n = {n_trials_iv_interventional_ex})')
-label_iv_lead_ex <- glue('No German UMC Lead\n(n = {n_trials_lead_ex})')
-label_trials_dupes_ex <- glue('Duplicate trials\n(n = {n_trials_dupes_ex})')
-label_pub_ex <- glue('No published results\n(n = {n_trials_pub_ex})')
-label_pmid_ex <- glue('No PMID\n(n = {n_trials_pmid_ex})')
-label_pubmed_ex <- glue('PMID unresolved in PubMed\n(n = {n_trials_pubmed_ex})')
-label_fulltext_ex <- glue('No fulltext (pdf) found\n(n = {n_trials_ft_pdf_ex})')
-
-# Prepare flowchart
-flow_trials <- DiagrammeR::grViz("digraph trials {
-
-# GRAPH
-graph [layout = dot, rankdir = LR, splines = false]
[…]

### 2021-08-21 R/analysis/03_tbl-trials.R

@@ -1,175 +1,0 @@
-# Summarize characteristics of trials and publications included in sample
-
-# Summarize all trial info ------------------------------------------------
-
-# tbl_trials_full <-
-#   trials %>%
-#
-#   # Add indicators for none and all
-#   mutate(
-#     has_trn_all = case_when(
-#       has_trn_secondary_id & has_trn_abstract & (has_trn_ft_pmc | has_trn_ft_pdf) ~ TRUE,
-#       is.na(has_trn_secondary_id) | is.na(has_trn_abstract) | (is.na(has_trn_ft_pmc) & is.na(has_trn_ft_pdf)) ~ NA,
-#       TRUE ~ FALSE
-#     ),
-#     has_trn_none = if_else(
-#       !has_trn_secondary_id & !has_trn_abstract & (!has_trn_ft_pmc | !has_trn_ft_pdf),
-#       TRUE, FALSE, missing = FALSE
-#     )
-#   ) %>%
-#
-#   select(
-#     has_summary_results,
-#     is_multicentric,
-#     enrollment,
-#     main_sponsor,
-#     # recruitment_status, # could have changed
-#     phase,
-#     masking,
-#     allocation,
-#     intervention_type,
-#     # lead_cities, # separate table/analysis?
-#     starts_with("days_"),
-#     starts_with("has_trn") & !ends_with("intovalue"),
-#   ) %>%
-#   gtsummary::tbl_summary(
-#     label = list(
-#       has_trn_any ~ "Anywhere",
-#       has_trn_all ~ "Metadata + Abstract + Full-text (PMC or PDF)",
-#       has_trn_none ~ "Nowhere",
-#       has_trn_secondary_id ~ "PubMed Metadata" ,
-#       has_trn_abstract ~ "Abstract",
-#       has_trn_ft_pmc ~ "PMC Full-Text",
-#       has_trn_ft_pdf ~ "PDF Full-Text"
-#     ),
-#     statistic =
-#       # all_logical() ~
-#       c("has_trn_any",
-#         "has_trn_all",
-#         "has_trn_none",
-#         "has_trn_secondary_id",
-#         "has_trn_abstract",
-#         "has_trn_ft_pmc",
-#         "has_trn_ft_pdf") ~
-#       "{n}/{N_nonmiss} ({p}%)"#,
-#     # missing = "no"
-#   ) %>%
-#     # modify_caption("**Trial registration number (TRN) reporting across published PubMed-indexed results from German UMC-led registered trials**")
[…]

### 2021-08-21 R/analysis/04_link-reg-pub.R

@@ -1,349 +1,0 @@
-# Report results as: N (%)
-# Most results are relative to `n_trials_ft_pdf`
-# However, some have different denominator, so optional parameter
-report_n_pct_trials <- function(numerator, denominator = n_trials_ft_pdf){
-  pct <- gtsummary::style_percent(numerator/denominator, digits = 0, symbol =  TRUE)
-  glue::glue("{numerator} ({pct})")
-}
-
-# Report results as % (n = NUM/DENOM)
-report_pct_num_denom_trials <- function(numerator, denominator = n_trials_ft_pdf){
-  pct <- gtsummary::style_percent(numerator/denominator, digits = 0, symbol =  TRUE)
-  glue::glue("{pct} ({numerator}/{denominator})")
-}
-
-# report_n_pct_trials(n_trials_link_reg_drks, n_trials_drks)
-# report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)
-
-# Full linkage ------------------------------------------------------------
-
-n_trials_link_full <-
-  trials %>%
-  filter(has_reg_pub_link &
-           has_iv_trn_secondary_id &
-           has_iv_trn_abstract &
-           has_iv_trn_ft_pdf
-  ) %>%
-  nrow()
-
-n_trials_link_none <-
-  trials %>%
-  filter(
-    !has_reg_pub_link &
-      !has_iv_trn_secondary_id &
-      !has_iv_trn_abstract &
-      !has_iv_trn_ft_pdf
-  ) %>%
-  nrow()
-
-n_trials_link_ft_abs_reg <-
-  trials %>%
-  filter(has_reg_pub_link &
-           # has_iv_trn_secondary_id &
-           has_iv_trn_abstract &
-           has_iv_trn_ft_pdf
-  ) %>%
-  nrow()
-
-
-# Publication reference in registration -----------------------------------
-
-n_trials_link_reg <-
-  trials %>%
-  filter(has_reg_pub_link) %>%
-  nrow()
-
-n_trials_link_reg_ctgov <-
-  trials %>%
-  filter(has_reg_pub_link & registry == "ClinicalTrials.gov") %>%
-  nrow()
-
-n_trials_link_reg_drks <-
-  trials %>%
-  filter(has_reg_pub_link & registry == "DRKS") %>%
-  nrow()
-
-n_trials_ctgov <-
-  trials %>%
[…]

### 2021-08-21 R/analysis/05_upset-link-reg-pub.R

@@ -1,100 +1,0 @@
-# Prepare trials with links
-# Transform links into list column of intersection sets
-trials_links <-
-  trials %>%
-  select(id,
-         registry,
-         has_reg_pub_link,
-         has_iv_trn_secondary_id,
-         has_iv_trn_abstract,
-         has_iv_trn_ft_pdf
-  ) %>%
-  rename(
-    "Reference in registration" = has_reg_pub_link,
-    "TRN (PubMed metadata)" = has_iv_trn_secondary_id,
-    "TRN (Abstract)" = has_iv_trn_abstract,
-    "TRN (Full-text)" = has_iv_trn_ft_pdf
-  ) %>%
-  pivot_longer(
-    cols = c(-id, -registry),
-    names_to = "link"
-  ) %>%
-  filter(value == TRUE) %>%
-  group_by(id) %>%
-  mutate(links = list(link)) %>%
-  ungroup() %>%
-  select(-value, -link) %>%
-  distinct()
-
-# Prepare trials without links
-# Create dummy links list column
-trials_no_links <-
-  trials %>%
-  filter(
-    !has_reg_pub_link &
-      !has_iv_trn_secondary_id &
-      !has_iv_trn_abstract &
-      !has_iv_trn_ft_pdf
-  ) %>%
-  select(id,
-         registry
-  ) %>%
-  mutate(links = list(NULL))
-
-# plot_upset_links_reg_pub <-
-#   bind_rows(trials_links, trials_no_links) %>%
-#   ggplot(aes(x = links)) +
-#   geom_bar() +
-#   geom_text(stat='count', aes(label = after_stat(count)), vjust = -.5) +
-#   ggupset::scale_x_upset() +
-#   ylab("Number of trials") + #not number of TRNs since 1 reg could have multi-trn
-#   xlab(NULL) +
-#   ggupset::theme_combmatrix(
-#     # combmatrix.label.make_space = TRUE,
-#     combmatrix.panel.line.size = 0,
-#     combmatrix.label.text = element_text(size=11)
-#   )
-
-# ggsave(
-#   here("docs", "figures", "plot-upset-links-reg-pub.png"),
-#   plot_upset_links_reg_pub,
-#   scale = 1.25
-# )
-
-plot_upset_links_reg_pub_registry <-
-  bind_rows(trials_links, trials_no_links) %>%
-  rename(Registry = registry) %>%
-  ggplot(aes(x = links)) +
-
-  # TODO: switch to "dodge" and fix numbers
[…]

### 2021-08-21 R/analysis/06_plot-change-over-time.R

@@ -1,212 +1,0 @@
-# Percentage of trials reg-pub linkage, by year
-# Denominator n = 1902 (n trials)
-
-# By registry -------------------------------------------------------------
-
-links_by_year_reg <-
-  trials %>%
-  select(
-    id,
-    registry, # could also group by registry
-    completion_year,
-    has_reg_pub_link,
-    has_iv_trn_secondary_id,
-    has_iv_trn_abstract,
-    has_iv_trn_ft_pdf
-  ) %>%
-  rename(
-    "Reference in registration" = has_reg_pub_link,
-    "TRN (PubMed metadata)" = has_iv_trn_secondary_id,
-    "TRN (Abstract)" = has_iv_trn_abstract,
-    "TRN (Full-text)" = has_iv_trn_ft_pdf
-  ) %>%
-  pivot_longer(c(-id, -registry, -completion_year),
-               names_to = "link_type",
-               values_to = "link",
-               values_drop_na = TRUE
-  ) %>%
-  group_by(link_type, registry, completion_year) %>%
-  summarise(
-    n_link_type= n(),
-    n_link = sum(link),
-    prop_link = n_link/n_link_type,
-    .groups = "drop"
-  )
-
-# Prepare plot ------------------------------------------------------------
-library(showtext)
-# font_add_google("Poppins", "Poppins")
-font_add_google("Roboto", "Roboto")
-font_add_google("Roboto Mono", "Roboto Mono")
-showtext_auto()
-
-theme_set(theme_light(base_size = 18, base_family = "Roboto"))
-# theme_set(theme_light(base_size = 18, base_family = "Poppins"))
-# theme_set(ggthemes::theme_fivethirtyeight(base_size = 18, base_family = "Poppins"))
-
-plot_links_year_reg <-
-  links_by_year_reg %>%
-  rename(Registry = registry) %>%
-  ggplot(aes(x = completion_year, y = prop_link, color = Registry, linetype = Registry)) +
-  facet_wrap(~link_type, ncol = 1) +
-  geom_point() +
-  geom_line() +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1), expand = c(0.005, 0.005)) +
-  scale_x_continuous(breaks = unique(links_by_year_reg$completion_year)) +
-  scale_color_manual(values = c("gray60", "gray10")) +
-  # scale_color_grey(start = 0.5, end = 0.8) +
-  # scale_colour_viridis_d() +
[…]

### 2021-08-21 R/analysis/cross-registrations.R

@@ -1,37 +1,0 @@
-n_trials_crossreg_reg <-
-  trials %>%
-  filter(n_crossreg_reg > 0) %>%
-  nrow()
-
-n_crossreg_reg <- sum(trials$n_crossreg_reg)
-
-# # by registry
-# trials %>% count(registry, n_crossreg_reg > 0)
-# trials %>%
-#   mutate(has_crossreg_reg = n_crossreg_reg > 0) %>%
-#   select(registry, has_crossreg_reg) %>%
-#   group_by(registry, has_crossreg_reg) %>%
-#   summarise(n = n()) %>%
-#   mutate(prop = n / sum(n))
-
-tbl_crossreg <-
-  trials %>%
-  select(
-    registry,
-    n_crossreg_reg,
-    n_crossreg_abstract, n_crossreg_secondary_id, n_crossreg_ft_pdf
-  ) %>%
-  gtsummary::tbl_summary(
-    by = registry,
-    label = list(
-      n_crossreg_reg ~ "Cross-registrations in registry",
-      n_crossreg_abstract ~ "Additional TRNs in abstract",
-      n_crossreg_secondary_id ~ "Additional TRNs in PubMed metadata",
-      n_crossreg_ft_pdf ~ "Additional TRNs in full-text"
-    )
-  ) %>%
-  add_stat_label(location = NULL) %>%
-
-  modify_header(label = "**Cross-registrations and additional TRNs**") %>%
-  # modify_caption("**German UMC-led trials with published result** (N = {N})") %>%
-  bold_labels()

### 2021-08-21 R/analysis/logistic-regression.R

@@ -1,196 +1,0 @@
-# http://www.danieldsjoberg.com/gt-and-gtsummary-presentation/#34
-library(dplyr)
-library(labelled)
-library(gtsummary)
-
-
-# Prepare trials and regression functions ---------------------------------
-
-# Convert variable to labelled factors for regressions
-trials_model <-
-  trials %>%
-  mutate(
-    across(c(registry, has_reg_pub_link, starts_with("has_iv_trn_")),
-           as.factor)
-  ) %>%
-  labelled::set_variable_labels(
-    has_reg_pub_link = "Publication in Registry",
-    has_iv_trn_secondary_id = "TRN in PubMed Metadata",
-    has_iv_trn_abstract = "TRN in Abstract",
-    has_iv_trn_ft_pdf = "TRN in Full-text",
-    completion_year = "Completion year",
-    registry = "Registry"
-  )
-
-# Multivariate regressions (adjusted odds ratio)
-tbl_glm_mv_trials <- function(trials_formula) {
-
-  glm(
-    formula = trials_formula,
-    data = trials_model,
-    family = "binomial"
-  ) %>%
-
-    gtsummary::tbl_regression(
-      exponentiate = TRUE,
-      show_single_row = c(-completion_year, -registry),
-      # add_estimate_to_reference_rows = TRUE # Change reference row
-    ) %>%
-
-    modify_table_styling(
-      column = estimate,
-      rows = !is.na(estimate),
-      cols_merge_pattern = "{estimate} ({ci}; {p.value})",
-      label = "**aOR (95% CI; p-value)**"
-    ) %>%
-
-    # Change reference label
-    # Thanks to gtsummary dev: https://github.com/ddsjoberg/gtsummary/issues/942
-    modify_table_styling(
-      column = estimate,
-      rows = reference_row %in% TRUE,
-      missing_symbol = "1 (Ref)"
-    ) %>%
-
-    modify_header(update = list(label ~ ""))
-}
-
-# Univariate regressions (crude odds ratio)
-
-tbl_glm_uv_trials <- function(outcome) {
-
-  trials_model %>%
-
-    select(completion_year, registry, has_iv_trn_secondary_id, has_iv_trn_ft_pdf, has_reg_pub_link, has_iv_trn_abstract) %>%
-
-    tbl_uvregression(
[…]

### 2021-08-21 R/analysis/non-iv-links.R

@@ -1,36 +1,0 @@
-# How many (additional, non-iv) publications are listed in registrations?
-# How many (additional, non-iv) trns are in the registration, full-text, abstract, metadata?
-
-trials %>%
-  # filter(n_reg_pub_any>0) %>%
-  select(registry,
-         starts_with("n_reg_pub"),
-         starts_with("n_crossreg"
-         )) %>%
-  mutate(across(as.integer())) %>%
-  tbl_summary(
-    by = registry,
-    type = list(everything() ~ "continuous"),
-    statistic = list(everything() ~ "{median} ({min}, {max})")
-  ) %>%
-  add_overall()
-
-trials %>%
-  # count(n_crossreg_secondary_id)
-  count(n_reg_pub_any)
-  count(n_reg_pub_doi_or_pmid)
-
-trials %>%
-  filter(n_reg_pub_any==13)
-
-n_reg_pub_max_drks <-
-  trials %>%
-  filter(registry == "DRKS") %>%
-  summarise(max(n_reg_pub_any)) %>%
-  pull()
-
-n_reg_pub_max_ctgov <-
-  trials %>%
-  filter(registry == "ClinicalTrials.gov") %>%
-  summarise(max(n_reg_pub_any)) %>%
-  pull()

### 2021-08-21 R/analysis/reg-auto-link.R

@@ -1,57 +1,0 @@
-# Manually-linked publication in registration -----------------------------
-
-tbl_reg_pub_link_manual <-
-  trials %>%
-  filter(!reference_derived | is.na(reference_derived)) %>%
-  select(registry, has_reg_pub_link) %>%
-  gtsummary::tbl_summary(
-    by = registry,
-    statistic = everything() ~ "{p}% ({n} / {N})",
-    label = everything() ~ "Publication in registration"
-  ) %>%
-
-  # Remove rowname label
-  modify_header(label = "") %>%
-
-  # Remove N from column names
-  modify_header(all_stat_cols() ~ "**{level}**") %>%
-
-  # Add overall column and remove N
-  add_overall() %>%
-  modify_header(stat_0 ~ "**Overall**") %>%
-
-  # Remove footnote
-  modify_footnote(everything() ~ NA) %>%
-  bold_labels()
-
-
-# ClinicalTrials.gov publication in registration: auto vs. manual --------
-
-tbl_reg_pub_link_ctgov_auto_v_manual <-
-
-  trials %>%
-  filter(
-    registry == "ClinicalTrials.gov",
-    has_reg_pub_link
-  ) %>%
-  # tbl_cross(
-  #   has_reg_pub_link, reference_derived,
-  #   percent = "row",
-  #   margin = "row"
-  # )
-  select(registry, reference_derived) %>%
-  gtsummary::tbl_summary(
-    by = registry,
-    statistic = everything() ~ "{p}% ({n} / {N})",
-    label = everything() ~ "Automated link"
-  ) %>%
-
-  # Remove rowname label
-  modify_header(label = "") %>%
-
-  # Remove N from column names
-  modify_header(all_stat_cols() ~ "**{level}**") %>%
-
-  # Remove footnote
-  modify_footnote(everything() ~ NA) %>%
-  bold_labels()

### 2021-08-21 R/analysis/si-if-abs-ft.R

@@ -1,41 +1,0 @@
-# Function to create tables for si by registry ----------------------------
-
-tbl_si_registry <- function(var, var_name) {
-  trials %>%
-    mutate(has_iv_trn_abs_or_ft = has_iv_trn_abstract | has_iv_trn_ft_pdf) %>%
-    filter({{var}}) %>%
-    select(registry, has_iv_trn_secondary_id) %>%
-    gtsummary::tbl_summary(
-      by = registry,
-      statistic = has_iv_trn_secondary_id ~ "{p}% ({n} / {N})",
-      label = has_iv_trn_secondary_id ~ var_name
-    ) %>%
-
-    # Remove rowname label
-    modify_header(label = "") %>%
-
-    # Remove N from column names
-    modify_header(all_stat_cols() ~ "**{level}**") %>%
-
-    # Add overall column and remove N
-    add_overall() %>%
-    modify_header(stat_0 ~ "**Overall**") %>%
-
-    # Remove footnote
-    modify_footnote(everything() ~ NA) %>%
-    bold_labels()
-}
-
-
-# Prepare abstract and full-text tables -----------------------------------
-tbl_si_abs_or_ft <- tbl_si_registry(has_iv_trn_abs_or_ft, "TRN in abstract or full-text")
-tbl_si_abs <- tbl_si_registry(has_iv_trn_abstract, "TRN in abstract")
-tbl_si_ft <- tbl_si_registry(has_iv_trn_ft_pdf, "TRN in full-text")
-
-
-# Display si table --------------------------------------------------------
-
-tbl_si_if_abs_ft <-
-  tbl_stack(list(tbl_si_abs_or_ft, tbl_si_abs, tbl_si_ft)) %>%
-  modify_caption("Number and proportion of trials with TRN in metadata given TRN in abstract, full-text, or either. Denominator in each cell indicates number of trials with TRN in abstract, full-text, or either, overall or from the respective registry.") %>%
-  modify_spanning_header(all_stat_cols() ~ "**TRN in Metadata**")

### 2021-08-21 R/analysis/tbl-stakeholders.R

@@ -1,102 +1,0 @@
-# Prepare actions and examples --------------------------------------------
-
-
-action_researchers <- "
-- Include TRN in abstract and full-text per CONSORT and ICMJE guidelines
-- Link publication in registry
-"
-
-example_researchers <- "—"
-
-action_registries <- "
-- Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registry
-- Link publications automatically using TRNs in bibliographic database metadata
-- Provide metadata on linked publication type (i.e., result, background)
-"
-
-example_registries <- "ClinicalTrials.gov uses PubMed TRN metadata to automatically link publications in the registration  [@nationallibraryofmedicine2017c]. Our study shows this automated publication linking is responsible for the majority of linked publications in ClinicalTrials.gov and implies similar potential for DRKS."
-
-action_institutions <- "
-- Educate, support, and incentivize researchers on linking
-"
-
-example_institutions <- "
-Ethics review offices or UMC-core facilities for clinical research could give researchers a info sheet with trial registration and reporting guidelines when reviewing trial. Bonuses, i.e., *Leistungsorientierte Mittel (LOM)*, could be disseminated for trial registration, reporting, and linking.
-"
-
-action_journals <- "
-- Request TRN in specialized metadata field
-- Review abstract and full-text for TRN inclusion, using automated regexes and/or manual strategies
-- Provide TRN as metadata to bibliographic databases
-"
-
-example_journals <- "
-Taylor & Francis extracts TRNs from the abstract and full-text, submit this metadata to CrossRef and Pubmed, and displays the linked trial via Crossmark on the article page [@taylorfrancisgroup2020; @crossref2020c])
-"
-
-action_pubmed <- "
-- Integrate publisher-provided TRNs into metadata
-- Extract TRN as metadata from abstract and full-text, using automated regexes and/or manual strategies
-"
-
-example_pubmed <- "
-The National Library of Medicine relies on publisher-provided data and manual indexers to create the TRN PubMed metadata if the TRN appears in the abstract or full-text. Our study shows this manual-only strategy misses TRNs and could be semi-automated to detect more TRNs for manual verfication.
-"
-
-
-# Prepare stakeholder table
-tbl_stakeholders <-
-  dplyr::tribble(
[…]

### 2021-08-21 R/analysis/trn-regex-analysis.R

@@ -1,43 +1,0 @@
-si <- read_rds(here("data", "processed", "pubmed-si.rds"))
-
-# Visually inspected all detected TRNs and are all TP (no FP)
-si_tp <- nrow(filter(si, !is.na(trn)))
-si_fp <- 0
-
-# Visually inspect mismatching trns and registries
-
-# NOT CATEGORIZED AS TRNs:
-# 5 trns are severely misformatted drks with no preceding "drks" (n = 4) or only "s" (n = 1) (not detected)
-# 3 accession numbers are figshare
-
-# CATEGORIZED AS TRNs:
-# MINOR MISFORMATTING CLEANED:
-# 1 trn is eudract with preceeding letters (detected, preceeding letter removed)
-# 1 trn is eudract with country code (detected, country code removed)
-# INCORRECT REGISTRY:
-# 1 registry is attributed to genbank but drks
-# 1 registry is attributed to genbank but ct.gov
-si_trn_mismatches <-
-  si %>%
-  filter(!accession_number %in% trn |
-           !databank %in% registry)
-
-# Encode si not categorized as trns
-si_figshare <- 3
-si_drks_misformatted <- 5
-
-# Consider DRKS misformatting TN since severe and visually not classifiable
-si_tn <- si_figshare + si_drks_misformatted
-si_fn <- 0
-
-# Calculate sens/spec
-si_sens <- si_tp/(si_tp + si_fn)
-si_spec <- si_tn/(si_tn + si_fp)
-
-# Recalculate with alternate assumption that DRKS are TP
-si_tp_alt <- si_tp + si_drks_misformatted
-si_tn_alt <- si_figshare
-si_fn_alt <- si_drks_misformatted
-
-si_sens_alt <- si_tp_alt/(si_tp_alt + si_fn_alt)
-si_spec_alt <- si_tn_alt/(si_tn_alt + si_fp)

### 2021-08-21 manuscript/manuscript.Rmd
Add manuscript Rmd plus appendix and bib

--- 2021-08-02 docs/manuscript_papaja.Rmd
@@
 
-  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata.
+  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata, and only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft_abs")` trials met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
 In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links.
-The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_trn_ft")`).
+The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft")`).
 ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists.
 Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata.
-Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_ctgov_link_reg_trn_si_derived")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
+Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
 
@@
 
-keywords          : "clinical trials, registration, reporting, meta-research"
+keywords          : "clinical trials, registration, reporting, meta-research, research transparency"
 wordcount         : "3551"
@@
 bibliography      : ["references.bib", "r-references.bib"]
-# appendix          : "appendix.Rmd"
-
+appendix          : "appendix.Rmd"
 numbersections    : yes
@@
 always_allow_html: true
+
+nocite: |
+  @taylorfrancisgroup2020, @crossref2020c, @nationallibraryofmedicine2017c 
 ---
 
-```{r setup, include = FALSE}
+```{r rmd-setup, include = FALSE}
+knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
+```
 
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
-ggplot2::theme_set(ggplot2::theme_minimal())
-
+```{r packages}
 library(dplyr)
@@
 library(readr)
+library(fs)
 
 library(ggplot2)
+library(showtext)
+
 library(DiagrammeR)
@@
 library(papaja)
-# library(knitr)
-# library(kableExtra)
-
-library(rentrez)
-library(roadoi)
-theme_gtsummary_compact()
-
-# Get data
-trials_all <- read_rds(here("data", "processed", "trials.rds"))
-pubmed_all <- read_rds(here("data", "processed", "pubmed-main.rds"))
-
-# Prepare references
-r_refs("r-references.bib")
-r_citation <- cite_r() # file = "r-references.bib", footnote = TRUE, pkgs, withhold #r_citations$r and r_citations$pkgs
-
-# Source analyses
-source(here("R", "functions", "screening.R"))
-# # source(here("R", "functions", "report-results.R"))
-source(here("R", "analysis", "01_screen-trials.R"))
-# # source(here("R", "analysis", "02_flowchart-trials.R"))
-source(here("R", "analysis", "03_tbl-trials.R"))
-source(here("R", "analysis", "04_link-reg-pub.R"))
-# # source(here("R", "analysis", "05_upset-link-reg-pub.R"))
-source(here("R", "analysis", "06_plot-change-over-time.R"))
-source(here("R", "analysis", "trn-regex-analysis.R"))
-source(here("R", "analysis", "cross-registrations.R"))
-source(here("R", "analysis", "logistic-regression.R"))
-source(here("R", "analysis", "si-if-abs-ft.R"))
-source(here("R", "analysis", "reg-auto-link.R"))
-source(here("R", "analysis", "tbl-stakeholders.R"))
-# # fs::dir_ls(here("R", "analysis")) %>% 
-# #   purrr::walk(source)
+library(knitr)
+library(kableExtra)
 ```
 
-```{r abstract-vars, cache = TRUE}
-abs_n_trials_link_full <- report_pct_num_denom_trials(n_trials_link_full)
-abs_n_trials_link_none <- report_n_pct_trials(n_trials_link_none)
-abs_n_trials_link_trn_ft <- "1131, 60% of trials"
-  # Change to gtsummary once updated: https://stackoverflow.com/questions/68502977/gtsummaryinline-text-for-stat-0-with-pattern
-  # inline_text(tbl_link_by_reg, has_iv_trn_ft_pdf, stat_0, pattern = "{n}, {p}% of trials")
-  # report_n_pct_trials(n_trials_link_trn_ft)
+```{r r-citations}
+r_citation <- cite_r()                              # base R
+r_refs("r-references.bib", append = FALSE)          # references to loaded packages
+create_bib(
+  c("loggit", "rentrez", "roadoi"),
+  "r-references.bib", append = TRUE) # references to additional packages
+```
 
-abs_n_trials_ctgov_link_reg_trn_si_derived <-
-  inline_text(tbl_reg_pub_link_ctgov_auto_v_manual, reference_derived, column = "stat_1", pattern = "{p}%")
+```{r figures-setup}
+# figures (plots and tables)
+
+# Prepare directory
+dir_figures <- dir_create(here("manuscript", "figures"))
+
+# theme_gtsummary_compact()
+
+# Set themes
+lightgray <- "gray75" #"coral3" # #"gray60"
+darkgray <- "gray35" #"gray10"
+registry_colors <- c("DRKS" = darkgray, "ClinicalTrials.gov" = lightgray)
+# registry_linetype <- c("DRKS" = "longdash", "ClinicalTrials.gov" = "solid")
+
+# font_add_google("Poppins", "Poppins")
+font_add_google("Roboto", "Roboto")
+font_add_google("Roboto Mono", "Roboto Mono")
+showtext_auto()
+
+theme_set(theme_light(
+  # base_size = 18,
+  base_family = "Roboto"
+))
+
+theme_update(
+  text = element_text(family = "Roboto"),
+  # axis.title = element_text(size = 16),
+  axis.text = element_text(family = "Roboto Mono",
+                           size = 8),
+)
+```
+
+```{r raw-data, cache = TRUE}
+# Dataset created in https://github.com/maia-sh/intovalue-data
+
+# TODO: create a github release and change link to version; for now using commit link: https://docs.github.com/en/github/administering-a-repository/releasing-projects-on-github/managing-releases-in-a-repository
+# "https://github.com/maia-sh/intovalue-data/raw/main/data/processed/trials.csv?raw=true"
+
+dir <- dir_create(here("data"))
+
+# Trials dataset
+trials_raw <- 
+  read_csv("https://github.com/maia-sh/intovalue-data/blob/b7909d4059e1435695e6e0be56d4da468b1f5a96/data/processed/trials.csv?raw=true") %>% 
+  write_csv(fs::path(dir, "trials-raw.csv"))
+
+# PubMed metadata
+pubmed_si <- read_rds("https://github.com/maia-sh/intovalue-data/blob/a95fd7e93f1838a595e747f7f756a13a49fb66bd/data/processed/pubmed/pubmed-si.rds?raw=true") %>% 
+  write_rds(fs::path(dir, "pubmed-si.rds"))
+
+# Query dates
+download.file(
+  "https://github.com/maia-sh/intovalue-data/raw/6b97fb62804aae1bab6c85bd2817baa027207bab/queries.log",
+  fs::path(dir, "queries.log")
+)
+
+query_logs <- loggit::read_logs(fs::path(dir, "queries.log"))
+
+# Extract most recent query dates
+get_latest_query <- function(query, logs) {
+  logs %>%
+    filter(log_msg == query) %>%
+    arrange(desc(timestamp)) %>%
+    slice_head(n = 1) %>%
+    pull(timestamp) %>%
+    as.Date.character() %>%
+    format("%d %B %Y")
+}
+
+pubmed_query_date <- get_latest_query("PubMed", query_logs)
+aact_query_date <- get_latest_query("AACT", query_logs)
+drks_query_date <- get_latest_query("DRKS", query_logs)
+```
+
+```{r analysis, cache = TRUE}
+dir_ls(here("analysis")) %>%
+  walk(source)
+```
+
+```{r abstract-vars, cache = TRUE, dependson = "analysis"}
+abs_n_trials_link_full <- 
+  report_trials(n_trials_link_full, pattern = "{p}% ({n}/{N})")
+abs_n_trials_link_ft_abs <- 
+  report_trials(n_trials_link_ft_abs, pattern = "{p}% ({n})")
+abs_n_trials_link_none <- 
+  report_trials(n_trials_link_none)
+abs_n_trials_link_ft <- 
+  inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0, pattern = "{n}, {p}% of trials")
+abs_n_trials_link_reg_ctgov_auto <- 
+  report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")
 ```
@@
 Clinical trial evidence informs clinical practice and health policy decision making through reviews and meta-analyses.
-The validity of health evidence synthesis is jeopardized if relevant trial data are left out to do inadequate or incomplete trial metadata, or inability to find trial results through automated means.
-Incomplete evidence threatens healthcare and the lives and welfare of human volunteers.
+The validity of health evidence synthesis is jeopardized if relevant trial data are either left out due to an inability to find results through automated means based on inadequate trial metadata, or duplicated due to an inability to connect publications and their registrations.
+Incomplete or duplicated evidence threatens the validity of clinical decision-making and potentially misinforms decisions regarding the launch of subsequent clinical trials, which may affect the wellbeing of patients enrolled in them [@chan2014tl; @councilforinternationalorganizationsofmedicalsciences2016]<!--# undermines the justification for insurance reimbursements, drives the misallocation of healthcare resources-->.
 Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp; @chalmers2013b].
-Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing clinical research value [@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->.
+Linking trial registrations and results publications makes each more findable, reducing the risk of information loss or duplication and increasing clinical research value [@kubben2019]<!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->.
 Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].
@@
 <!--# add policy background on why pub in reg --> Results references may also be entered manually in many registries, with varying degrees of structure.
-ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas DRKS offers an unstructured free-text field.
+ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas the German Clinical Trials Registry (DRKS) offers an unstructured free-text field.
 <!--# add some background on ctgov vs.
@@
 TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b; @huser2013jamia; @wetering2012po].
-The sampling strategies used in these previous studies limit our ability to draw conclusions about registration-publication linking.
+The sampling strategies used in previous studies limit our ability to draw conclusions about registration-publication linking.
 Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included non-trial-result publications, and trials with unknown registration rates (as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo]); registration-based cohorts included included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b].
@@
 
-The exploratory study evaluates linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) with completion dates between 2009 and 2017.
+The exploratory study evaluates linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or DRKS with completion dates between 2009 and 2017.
 We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs, which may be applied to other trial cohorts.
@@
 
-We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset [@riedel2021a].
-In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.
+We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021c and referred to as the "IntoValue" dataset [@riedel2021a].
+The dataset consists of clinical trials registered on ClinicalTrials.gov or DRKS and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.
 
-We <!--# programmatically -->downloaded data from both registries on 15 May 2021.
+We <!--# programmatically -->downloaded data from both registries on `r ifelse(aact_query_date == drks_query_date, aact_query_date, stop())`.
 <!--# We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data.
@@
 
-We queried the PubMed Entrez Programming Utilities (E-utilities) API<!--#on DAY MONTH YEAR--> for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @R-rentrez].
+We queried the PubMed Entrez Programming Utilities (E-utilities) API on `r pubmed_query_date` for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @R-rentrez].
 From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->
@@
 
-After updating registry data, we reapplied the IntoValue exclusion criteria (completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead; see @riedel2021 for further details on these criteria) and limited our sample to trials with results with a PMID and full-text publication.
+After updating registry data, we reapplied the IntoValue exclusion criteria (completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead; see @riedel2021c for further details on these criteria) and limited our sample to trials with results with a PMID and full-text publication.
 
@@
 
-As reference data is available in different formats depending on the registry, we developed parallel approaches to extract publication identifiers for each.
+As publication data is available in different formats depending on the registry, we developed parallel approaches to extract publication identifiers for each.
 For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT.
@@
 We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables.
-Regressions were performed as exploratory analyses, and all variables were included in each regression model.<!--#no model selection or fitting was conducted; the explanatory variables were tested a priori to verify there was no violation of the assumption of the linearity of the logit TODO, ask Murph--> In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata.
+Regressions were performed as exploratory analyses, and all variables were included in each regression model.
+There was no model selection or fitting, or correction for multiple testing.
+In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata.
 Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs.
 manually linked publications in ClinicalTrials.gov.
-To explore registry differences for only manually linked references, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration.
+To explore registry differences for only manually linked publications, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration.
 <!--#add automatic links to regression for pub in registry?
@@
 The IntoValue dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
-After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
-Figure \@ref(fig:flow-trials) details the trial and publication screening.
+After applying our exclusion criteria, our sample included `r n_trials_ft` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
+Appendix \@ref(app-flow-trials) provides a flow diagram of the trial and publication screening.
 Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.
-
-```{r flow-trials, fig.cap = "Trials and publications screening", out.width="50%", out.height="50%"}
-knitr::include_graphics(here("docs", "figures", "flow-trials.png"))
-```
 
@@
 ```{r tbl-trials, results = "asis"}
-# tbl <- read_rds(here("docs", "figures", "tbl-trials.rds"))
-# tbl %>% 
-#   as_kable_extra(booktabs = TRUE)
-tbl_trials %>%
-
-  as_kable_extra()
-  # as_tibble() %>%
-  # apa_table(
-  #   caption = "This is a table about trials",
-  #   format.args = list(na_string = ""),
-  #   format = "pandoc"
-  # )
-# modify_caption("Characteristics of included trials with results publications")
-  # gtsummary::as_flex_table() %>% 
-  # flextable::set_caption(caption = "Characteristics of included trials with results publications")
+as_kable(tbl_trials, format = 'pipe')
 ```
 
-While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata.
-Disregarding metadata, which is largely beyond trialists' control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_n_pct_trials(n_trials_link_ft_abs_reg)`.
-In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication.
-The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_n_pct_trials(n_trials_link_trn_ft_only)` trials.
-The inclusion of the various link types ranged from `r report_pct_num_denom_trials(n_trials_link_trn_abs)` in abstracts to `r report_pct_num_denom_trials(n_trials_link_trn_ft)` in full-text.
+Per our inclusion criteria all trials in our sample (n = `r n_trials_ft`) were registered and had a publication, however only `r report_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning the publication linked in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata.
+Disregarding metadata, which is largely beyond trialists' control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_trials(n_trials_link_ft_abs_reg)`.
+An additional `r report_trials(n_trials_link_ft_abs - n_trials_link_ft_abs_reg)` met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
+In contrast, we found `r report_trials(n_trials_link_none)` trials with no links in either the registration or the publication.
+The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_trials(n_trials_link_ft_only)` trials.
+The inclusion of the various link types ranged from `r inline_text(tbl_link_by_reg, has_iv_trn_abstract, stat_0, pattern = "{p}% ({n}/{N})")` in abstracts to `r inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0, pattern = "{p}% ({n}/{N})")` in full-text.
 Table \@ref(tab:tbl-link-by-reg) shows registration-publication links overall and by registry.
@@
 ```{r tbl-link-by-reg}
-tbl_link_by_reg %>% 
-  modify_caption("Registration-publication links by registry") #%>% 
-  # gtsummary::as_flex_table() #%>%
-  # flextable::set_caption(caption = "Registration-publication links by registry")
+as_kable(tbl_link_by_reg, format = "pipe")
 ```
 
-```{r plot-upset-links-reg-pub-registries, fig.cap = "Registration-publication links in ClinicalTrials.gov and DRKS", out.width="80%"}
-knitr::include_graphics(here("docs", "figures", "plot-upset-links-reg-pub-registry.png"))
+```{r plot-upset-links-reg-pub-registries, fig.cap = "Trials with all combinations of links between registration and publication in ClinicalTrials.gov and DRKS", out.width="80%"}
+knitr::include_graphics(path(dir_figures, "plot-upset-link-type.pdf"))
 ```
@@
 ```{r tbl-trials-regression}
-tbl_trials_regression %>% 
-  modify_caption("Crude and adjusted odds ratios for factors associated with publication-registration links")
-  # gtsummary::as_flex_table() %>% 
-  # flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
+# options(knitr.kable.NA = "**")
+as_kable(tbl_trials_regression, format = "pipe") #%>% 
+  # kable_styling(latex_options = "scale_down")
+  # landscape()
 ```
@@
 Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
-Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.
+Figure \@ref(fig:plot-link-year) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.
 
-```{r plot-links-year-reg, fig.cap = "Change over time in registration-publication links in ClinicalTrials.gov and DRKS. Completion year from registry."}
-knitr::include_graphics(here("docs", "figures", "plot-links-year-reg.png"))
+```{r plot-link-year, fig.cap = "Percentage of trials with linked registrations and publications by trial completion year in ClinicalTrials.gov and DRKS.
+Completion year from registry.", out.width = "95%"}
+knitr::include_graphics(path(dir_figures, "plot-link-year.pdf"))
 ```
 
-Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link.
-<!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`).
+Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with link type were generally more likely to have another type of link.
+<!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")`).
 Futhermore, DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
@@
 `r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_2", pattern = "{p}%")`).
-Appendix \@ref(si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
+Appendix \@ref(app-si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
 
-DRKS trials were also less likely than ClinicalTrials.gov trials to reference the publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
-<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r inline_text(tbl_reg_pub_link_ctgov_auto_v_manual, reference_derived, column = "stat_1", pattern = "{p}%, {n}/{N}")`) ClinicalTrials.gov trials referencing the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
+Trials registered in DRKS were also less likely than trials registered in ClinicalTrials.gov to reference the full-text primary outcome publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
+<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) ClinicalTrials.gov trials referencing the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
 Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")` vs.
 `r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_2", pattern = "{p}%, {n}/{N}")`).
+Trials also had additional publications linked in the registrations, up to `r reg_pub_max_ctgov$n_reg_pub_any` in ClinicalTrials.gov (`r reg_pub_max_ctgov$id`) and `r reg_pub_max_drks$n_reg_pub_any` in DRKS (`r reg_pub_max_drks$id`).
 
-## Evaluation of TRN regular expressions
-
-<!--# TODO: add to methods if this is staying? -->
-
-To gauge the sensitivity and specificity of the TRN regexes, we visually inspected all PubMed secondary identifiers extracted for our sample (n = `r nrow(si)`).
-PubMed's secondary identifier field can contain identifiers from not only clinical trial registries, but also other databanks, such as molecular sequences and open data repositories (i.e., figshare, Dryad) [@nationallibraryofmedicine2018].
-Most ids were clear true positive TRNs (n = `r si_tp`) and a handful were clear true negative other ids (n = `r si_figshare`).
+In our manual evaluation of the TRN regular expressions for classifying PubMed secondary identifiers (n = `r nrow(pubmed_si)`), we found that most ids were clear true positive TRNs (n = `r si_tp`).
+A handful (n = `r si_figshare`) were clear true negative other ids, meaning ids from non-registry databanks, such as molecular sequences and open data repositories (i.e., figshare, Dryad) [@nationallibraryofmedicine2018].
 We also found `r si_drks_misformatted` ids which we manually classified as true negative non-TRNs; however, using additional PubMed data, we determined these were severely misformatted DRKS ids (i.e., missing the preceding letters "DRKS" and just a string of numbers such as "00000711").
-Our regexes correctly classified all well formatted TRNs and non-TRNs, resulting in a sensitivity and specificity of 100%.
+Our regexes correctly classified all well formatted TRNs and non-TRNs, resulting in a sensitivity and specificity of `r ifelse(si_sens == si_spec, style_percent(si_sens), stop())`%.
 If we had instead categorized the `r si_drks_misformatted` severely misformatted DRKS ids as true positives, our regexes would have sensitivity of `r style_percent(si_sens_alt, digits = 1)`% and a specificity of `r style_percent(si_spec_alt)`%.
@@
 
-Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research.
-Poor linking poses a barrier to identifying trial publications via automated approaches and instead require researchers to perform intensive manual keyword-based searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
+Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of research.
+Poor linking poses a barrier to identifying trial publications via automated approaches and instead requires researchers to perform intensive manual searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
 This responsible research practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.
@@
 Our study shows that German UMCs can improve in both TRN reporting in publications and references to publications in the registration.
-In our sample (n = `r n_trials_ft_pdf`), `r report_n_pct_trials(n_trials_link_none)` trials had no links between registration and publication and only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication links.
-Furthermore, only `r report_n_pct_trials(n_trials_link_trn_ft_abs)` trials in our sample fully met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
+In our sample (n = `r n_trials_ft`), `r report_trials(n_trials_link_none)` trials had no links between registration and publication and only `r report_trials(n_trials_link_full)` trials had the most comprehensive registration-publication links.
+Furthermore, only `r report_trials(n_trials_link_ft_abs)` trials in our sample fully met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
 Linking practices showed at best minimal improvement over time.
@@
 As such, we expect a TRN from any registry to be included in the metadata, if the trialist includes it in either the abstract or the full-text.
-However, we found that while trials with a TRN in the abstract were indeed more likely to include the TRN in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), TRNs in the full-text were not more likely to appear in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`).
+However, we found that while trials with a TRN in the abstract were indeed more likely to include the TRN in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), TRNs in the full-text were not more likely to appear in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")`).
 Furthermore, our data suggest that NLM staff are better at extracting ClinicalTrials.gov than DRKS TRNs (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).<!--# As such, it appears that PubMed metadata misses TRNs from the full-text as well as DRKS trials.-->
 
-Automated indexing of publications in ClinicalTrials.gov accounts for most (`r inline_text(tbl_reg_pub_link_ctgov_auto_v_manual, reference_derived, column = "stat_1", pattern = "{p}%, {n}/{N}")`) publication references in the registry and drives the almost three-fold discrepancy with DRKS (`r inline_text(tbl_link_by_reg, has_reg_pub_link, "ClinicalTrials.gov", pattern = "{p}%, {n}/{N}")` vs.
-`r inline_text(tbl_link_by_reg, has_reg_pub_link, "DRKS", pattern = "{p}%, {n}/{N}")`).<!--# Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")`)--> Currently, DRKS allows only for manual submission of references by trialists and does not index publications.<!--# add?-->
+Automated indexing of publications in ClinicalTrials.gov accounts for most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) publication references in the registry and drives the almost three-fold discrepancy with DRKS (`r inline_text(tbl_link_by_reg, has_reg_pub_link, "ClinicalTrials.gov", pattern = "{p}%, {n}/{N}")` vs.
+`r inline_text(tbl_link_by_reg, has_reg_pub_link, "DRKS", pattern = "{p}%, {n}/{N}")`).<!--# Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")`)--> Currently, DRKS allows  for only manual submission of references by trialists and does not index publications.<!--# add?--> Furthermore, registrations may refer to additional publications beyond results (e.g., up to 72 publications for one ClinicalTrials.gov registration in our sample), so registry metadata should encode publication type to support quick identification of results; while such publication type metadata is currently available in both registries, it is not systematically used, and many publications are categorized generically as a "paper" (DRKS) or "reference" (ClinicalTrials.gov).
 
@@
 This study reveals shortcomings in the linking of German UMC-led trials registrations and results publications and highlights several promising avenues forward.
-In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is relatively straightforward and can be improved with action across stakeholders.
+In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is straightforward and can be improved with action across stakeholders.
 <!--#Drawing from the study findings and additional context, -->Table \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.
@@
 # https://www.anthonyschmidt.co/post/2020-06-03-making-apa-tables-with-gt/
-# results = "asis"
-# tbl_stakeholders %>%
-#   as_latex()
-knitr::include_graphics(here::here("docs", "figures", "tbl-stakeholders.png"))
-# tbl_stakeholders
+knitr::include_graphics(path(dir_figures, "tbl-stakeholders.pdf"))
 ```
@@
 With improved TRN inclusion in bibliographic metadata and increased automatic indexing of publications based on this metadata, full linking could be achieved with negligible work by researchers: simply pasting the TRN into the publication abstract and full-text.
-In our sample alone, an additional `r report_pct_num_denom_trials(n_trials_link_trn_ft_or_abs_no_si, n_trials_link_trn_ft_or_abs)` of trials with TRN in abstract or full-text could have the TRN included in PubMed's metadata, and an additional `r report_pct_num_denom_trials(n_trials_drks_trn_si_not_linked_pub, n_trials_drks_trn_si)` of DRKS trials with TRN in the PubMed metadata could be automatically indexed in the registration.
+In our sample alone, an additional `r report_trials(n_trials_link_ft_or_abs_no_si, n_trials_link_ft_or_abs, pattern = "{p}% ({n}/{N})")` of trials with TRN in abstract or full-text could have the TRN included in PubMed's metadata, and an additional `r report_trials(n_trials_drks_link_si_not_reg, n_trials_drks_link_si, pattern = "{p}% ({n}/{N})")` of DRKS trials with TRN in the PubMed metadata could be automatically indexed in the registration.
 With the uptake of the recommended actions across stakeholder groups, improved trial registration-publication links will allow for more comprehensive evidence synthesis and increase the value of clinical research.
@@
 
-This work was funded under a grant from the Federal Ministry of Education and Research of Germany (Bundesministerium für Bildung und Forschung - BMBF) [insert grant number].
+This work was funded under a grant from the Federal Ministry of Education and Research of Germany (Bundesministerium für Bildung und Forschung - BMBF) [01PW18012].
 The funder had no involvement in the study design, data collection, analysis, or interpretation, writing of the manuscript, or the decision to submit for publication.
@@
 
-\endgroup
-
-```{r render_appendix, include=FALSE}
-render_appendix("appendix.Rmd")
-```
+\newpage

### 2021-08-21 analysis/00_functions.R
Add analysis scripts

@@ -1,0 +1,19 @@
+# Generic reporting function
+# Based on gtsummary::inline_text
+# Default denominator (N) is analysis population (`n_trials_ft`)
+# report_trials(1137")
+# report_trials(1137, 5000")
+# report_trials(1137, pattern = "{p}% ({n}/{N})")
+report_trials <- function(n,
+                          N = n_trials_ft,
+                          pattern = "{n} ({p}%)",
+                          digits = 0){
+  # Format numbers
+  p <- gtsummary::style_percent(n/N, digits = digits, symbol =  FALSE)
+  n <- gtsummary::style_number(n)
+  N <- gtsummary::style_number(N)
+
+  glue::glue(pattern)
+}
+
+

### 2021-08-21 analysis/01_screen-trials.R

@@ -1,0 +1,130 @@
+# Prepare trials ----------------------------------------------------------
+
+trials_raw <-
+
+  readr::read_csv(here::here("data", "trials-raw.csv")) %>%
+
+  mutate(
+
+    # Create flag for dupes in IV1 for exclusion
+    is_not_iv1_dupe = if_else(!(is_dupe & iv_version == 1), TRUE, FALSE),
+
+    # Trials with a journal article have a publication (disregard dissertations and abstracts)
+    has_publication = if_else(publication_type == "journal publication", TRUE, FALSE, missing = FALSE),
+
+    has_pmid = if_else(!is.na(pmid), TRUE, FALSE)
+  )
+
+
+# Prepare screening functions ---------------------------------------------
+
+# Apply screening criteria
+# Returns list of inclusion counts and filtered dataframe
+count_filter <- function(data, vars) {
+
+  counts <-
+    tibble(name = as.character(),
+           value = as.logical(),
+           n = as.integer()
+    )
+
+  for (var in vars) {
+    counts <-
+      data %>%
+      count(.data[[var]]) %>%
+      pivot_longer(-n) %>%
+      add_row(counts, .)
+
+    data <- filter(data, .data[[var]])
+  }
+
+  list(data = data, counts = counts)
+
+}
+
+# Report screening summary counts
+report_n <- function(counts, var, condition) {
+  n <-
+    counts %>%
+    filter(name == var & value == condition) %>%
+    pull(n)
+
+  # If empty, count is 0
+  if (rlang::is_empty(n)){n <- 0}
+
+  n
+}
+
+
+# Screen trials -----------------------------------------------------------
+
+screening_criteria <- c(
+  "iv_completion",
+  "iv_status",
+  "iv_interventional",
+  "has_german_umc_lead",
+  "is_not_iv1_dupe",
+  "has_publication",
+  "has_pmid",
+  "has_pubmed",
+  "has_ft"
+)
+
+trials_screened <-  count_filter(trials_raw, screening_criteria)
+
+trials <- trials_screened$data
+
+write_csv(trials, here("data", "trials.csv"))
+
+# Tabularize trial screening counts ---------------------------------------
+
+trial_screening <-
+
+  # IntoValue 1 and 2
+  trials_raw %>%
[…]

### 2021-08-21 analysis/analysis_logistic-regression.R

@@ -1,0 +1,154 @@
+# http://www.danieldsjoberg.com/gt-and-gtsummary-presentation/#34
+
+# Prepare trials and regression functions ---------------------------------
+
+# Convert variable to labelled factors for regressions
+trials_model <-
+  trials %>%
+  mutate(
+    across(c(registry, has_reg_pub_link, starts_with("has_iv_trn_")),
+           as.factor)
+  ) %>%
+  labelled::set_variable_labels(
+    has_iv_trn_ft = "TRN in full-text",
+    has_iv_trn_abstract = "TRN in abstract",
+    has_iv_trn_secondary_id = "TRN in PubMed metadata",
+    has_reg_pub_link = "Publication in registration",
+    completion_year = "Completion year",
+    registry = "Registry"
+  )
+
+# Multivariate regressions (adjusted odds ratio)
+tbl_glm_mv_trials <- function(trials_formula) {
+
+  glm(
+    formula = trials_formula,
+    data = trials_model,
+    family = "binomial"
+  ) %>%
+
+    gtsummary::tbl_regression(
+      exponentiate = TRUE,
+      show_single_row = c(-completion_year, -registry),
+      # add_estimate_to_reference_rows = TRUE # Change reference row
+    ) %>%
+
+    modify_table_styling(
+      column = estimate,
+      rows = !is.na(estimate),
+      cols_merge_pattern = "{estimate} ({ci}; {p.value})",
+      label = "**aOR (95% CI; p-value)**"
+    ) %>%
+
+    # Change reference label
+    # Thanks to gtsummary dev: https://github.com/ddsjoberg/gtsummary/issues/942
+    modify_table_styling(
+      column = estimate,
+      rows = reference_row %in% TRUE,
+      missing_symbol = "1 (Ref)"
+    ) %>%
+
+    modify_header(update = list(label ~ ""))
+}
+
+# Univariate regressions (crude odds ratio)
+
+tbl_glm_uv_trials <- function(outcome) {
+
+  trials_model %>%
+
+    select(completion_year, registry, has_iv_trn_ft, has_iv_trn_abstract, has_iv_trn_secondary_id,  has_reg_pub_link) %>%
+
+    tbl_uvregression(
+      method = glm,
+      y = {{outcome}},
+      method.args = list(family = binomial),
+      exponentiate = TRUE,
[…]

### 2021-08-21 analysis/analysis_reg-pub-link.R

@@ -1,0 +1,185 @@
+# Individual results ------------------------------------------------------
+
+# Trials with all links
+n_trials_link_full <-
+  trials %>%
+  filter(
+    has_reg_pub_link &
+      has_iv_trn_secondary_id &
+      has_iv_trn_abstract &
+      has_iv_trn_ft
+  ) %>%
+  nrow()
+
+# Trials with no links
+n_trials_link_none <-
+  trials %>%
+  filter(
+    !has_reg_pub_link &
+      !has_iv_trn_secondary_id &
+      !has_iv_trn_abstract &
+      !has_iv_trn_ft
+  ) %>%
+  nrow()
+
+# Trials with full-text and abstract links (CONSORT/ICMJE compliant)
+n_trials_link_ft_abs <-
+  trials %>%
+  filter(has_iv_trn_ft & has_iv_trn_abstract) %>%
+  nrow()
+
+# Trials with full-text, abstract and registry links (PubMed metadata not in researchers control)
+n_trials_link_ft_abs_reg <-
+  trials %>%
+  filter(
+    has_reg_pub_link &
+      has_iv_trn_abstract &
+      has_iv_trn_ft
+  ) %>%
+  nrow()
+
+# Trials with only full-text link
+n_trials_link_ft_only <-
+  trials %>%
+  filter(
+    has_iv_trn_ft &
+      !has_iv_trn_abstract & !has_iv_trn_secondary_id & !has_reg_pub_link
+  ) %>%
+  nrow()
+
+# Trials with trn in ft or abs
+n_trials_link_ft_or_abs <-
+  trials %>%
+  filter(has_iv_trn_ft | has_iv_trn_abstract) %>%
+  nrow()
+
+# Trials with trn in ft or abs and NOT in si
+n_trials_link_ft_or_abs_no_si <-
+  trials %>%
+  filter(
+    has_iv_trn_ft | has_iv_trn_abstract,
+    !has_iv_trn_secondary_id
+  ) %>%
+  nrow()
+
+# DRKS trials with trn in si
+n_trials_drks_link_si <-
+  trials %>%
+  filter(
+    registry == "DRKS",
+    has_iv_trn_secondary_id
+  ) %>%
+  nrow()
+
+# DRKS trials with trn in si and NOT linked pub in registry
+n_trials_drks_link_si_not_reg <-
+  trials %>%
+  filter(
+    registry == "DRKS",
+    has_iv_trn_secondary_id,
+    !has_reg_pub_link
+  ) %>%
+  nrow()
+
+
+
[…]

### 2021-08-21 analysis/analysis_si-if-abs-ft.R

@@ -1,0 +1,42 @@
+# Analysis of TRN in PubMed metadata given TRN in abstract/full-text
+# Function to create tables for si by registry ----------------------------
+
+tbl_si_registry <- function(var, var_name) {
+  trials %>%
+    mutate(has_iv_trn_abs_or_ft = has_iv_trn_abstract | has_iv_trn_ft) %>%
+    filter({{var}}) %>%
+    select(registry, has_iv_trn_secondary_id) %>%
+    gtsummary::tbl_summary(
+      by = registry,
+      statistic = has_iv_trn_secondary_id ~ "{p}% ({n} / {N})",
+      label = has_iv_trn_secondary_id ~ var_name
+    ) %>%
+
+    # Remove rowname label
+    modify_header(label = "") %>%
+
+    # Remove N from column names
+    modify_header(all_stat_cols() ~ "**{level}**") %>%
+
+    # Add overall column and remove N
+    add_overall() %>%
+    modify_header(stat_0 ~ "**Overall**") %>%
+
+    # Remove footnote
+    modify_footnote(everything() ~ NA) %>%
+    bold_labels()
+}
+
+
+# Prepare abstract and full-text tables -----------------------------------
+tbl_si_abs_or_ft <- tbl_si_registry(has_iv_trn_abs_or_ft, "TRN in abstract or full-text")
+tbl_si_abs <- tbl_si_registry(has_iv_trn_abstract, "TRN in abstract")
+tbl_si_ft <- tbl_si_registry(has_iv_trn_ft, "TRN in full-text")
+
+
+# Display si table --------------------------------------------------------
+
+tbl_si_if_abs_ft <-
+  tbl_stack(list(tbl_si_abs_or_ft, tbl_si_abs, tbl_si_ft)) %>%
+  modify_caption("Number and proportion of trials with TRN in metadata given TRN in abstract, full-text, or either. Denominator in each cell indicates number of trials with TRN in abstract, full-text, or either, overall or from the respective registry.") %>%
+  modify_spanning_header(all_stat_cols() ~ "**TRN in Metadata**")

### 2021-08-21 analysis/analysis_trn-regex-si.R

@@ -1,0 +1,48 @@
+# Visually inspected all detected TRNs and are all TP (no FP)
+si_tp <- nrow(filter(pubmed_si, !is.na(trn)))
+si_fp <- 0
+
+# Visually inspect mismatching trns and registries
+
+# NOT CATEGORIZED AS TRNs:
+# 5 trns are severely misformatted drks with no preceding "drks" (n = 4) or only "s" (n = 1) (not detected)
+# 3 accession numbers are figshare
+
+# CATEGORIZED AS TRNs:
+# MINOR MISFORMATTING CLEANED:
+# 1 trn is eudract with preceeding letters (detected, preceeding letter removed)
+# 1 trn is eudract with country code (detected, country code removed)
+# INCORRECT REGISTRY:
+# 1 registry is attributed to genbank but drks
+# 1 registry is attributed to genbank but ct.gov
+si_trn_mismatches <-
+  pubmed_si %>%
+  filter(!accession_number %in% trn |
+           !databank %in% registry)
+
+# Encode si not categorized as trns
+si_figshare <-
+  si_trn_mismatches %>%
+  filter(databank == "figshare") %>%
+  nrow()
+
+si_drks_misformatted <-
+  si_trn_mismatches %>%
+  filter(databank == "DRKS") %>%
+  nrow()
+
+# Consider DRKS misformatting TN since severe and visually not classifiable
+si_tn <- si_figshare + si_drks_misformatted
+si_fn <- 0
+
+# Calculate sens/spec
+si_sens <- si_tp/(si_tp + si_fn)
+si_spec <- si_tn/(si_tn + si_fp)
+
+# Recalculate with alternate assumption that DRKS are TP
+si_tp_alt <- si_tp + si_drks_misformatted
+si_tn_alt <- si_figshare
+si_fn_alt <- si_drks_misformatted
+
+si_sens_alt <- si_tp_alt/(si_tp_alt + si_fn_alt)
+si_spec_alt <- si_tn_alt/(si_tn_alt + si_fp)

### 2021-08-21 analysis/fig_flowchart-trials.R

@@ -1,0 +1,138 @@
+# Prepare labels
+label_iv_all <- glue('IntoValue trials\n(n = {n_trials_iv})\nIV1 = {n_trials_iv1} & IV2 = {n_trials_iv2}')
+label_iv_screened <- glue('Trials meeting IntoValue criteria\nwith new registry data\n(n = {n_trials_iv_interventional})')
+label_trials_lead <- glue('IntoValue German UMC Lead Trials\n(n = {n_trials_lead})')
+label_trials_deduped <- glue('Trials with duplicates removed\n(n = {n_trials_deduped})')
+label_pub <- glue('Trials with published results\n(n = {n_trials_pub})')
+label_pmid <- glue('Trials with published results\nwith PMID\n(n = {n_trials_pmid})')
+# label_pubmed <- glue('Trials with published results\nwith resolved PubMed record\n(n = {n_trials_pubmed})')
+label_fulltext <- glue('Trials with published results\nas full-text (pdf)\n(n = {n_trials_ft})')
+
+label_iv_screened_ex <- glue('Trials not meeting IntoValue criteria\nwith new registry data\n(n = {n_trials_iv_completion_date_ex + n_trials_iv_status_ex + n_trials_iv_interventional_ex})\nNot completed 2009-2017 (n = {n_trials_iv_completion_date_ex})\nNot included status (n = {n_trials_iv_status_ex})\nNot interventional (n = {n_trials_iv_interventional_ex})')
+label_iv_lead_ex <- glue('No German UMC Lead\n(n = {n_trials_lead_ex})')
+label_trials_dupes_ex <- glue('Duplicate trials\n(n = {n_trials_dupes_ex})')
+label_pub_ex <- glue('No published results\n(n = {n_trials_pub_ex})')
+label_pmid_ex <- glue('No PMID\n(n = {n_trials_pmid_ex})')
+# label_pubmed_ex <- glue('PMID unresolved in PubMed\n(n = {n_trials_pubmed_ex})')
+label_fulltext_ex <- glue('No full-text (pdf) found\n(n = {n_trials_ft_ex})')
+
+# Prepare flowchart
+flow_trials <- DiagrammeR::grViz("digraph trials {
+
+# GRAPH
+graph [layout = dot, rankdir = LR, splines = false]
[…]

### 2021-08-21 analysis/fig_plot-change-over-time.R

@@ -1,0 +1,126 @@
+# Percentage of trials reg-pub linkage, by year
+# Denominator n = 1902 (n trials)
+
+# By registry -------------------------------------------------------------
+
+links_by_year_reg <-
+  trials %>%
+  select(
+    id,
+    registry, # could also group by registry
+    completion_year,
+    has_reg_pub_link,
+    has_iv_trn_secondary_id,
+    has_iv_trn_abstract,
+    has_iv_trn_ft
+  ) %>%
+  rename(
+    "TRN in full-text" = has_iv_trn_ft,
+    "TRN in abstract" = has_iv_trn_abstract,
+    "TRN in PubMed metadata" = has_iv_trn_secondary_id,
+    "Publication in registration" = has_reg_pub_link
+  ) %>%
+  pivot_longer(c(-id, -registry, -completion_year),
+               names_to = "link_type",
+               values_to = "link",
+               values_drop_na = TRUE
+  ) %>%
+  mutate(
+    link_type = factor(
+      link_type,
+      # levels = c(
+      #   has_iv_trn_ft,
+      #   has_iv_trn_abstract,
+      #   has_iv_trn_secondary_id,
+      #   has_reg_pub_link
+      # ),
+      levels = c(
+        "TRN in full-text",
+        "TRN in abstract",
+        "TRN in PubMed metadata",
+        "Publication in registration"
+      )
+    )
+  ) %>%
+  group_by(link_type, registry, completion_year) %>%
+  summarise(
+    n_link_type= n(),
+    n_link = sum(link),
+    prop_link = n_link/n_link_type,
+    .groups = "drop"
+  )
+
+# Prepare plot ------------------------------------------------------------
+# library(showtext)
+# # font_add_google("Poppins", "Poppins")
+# font_add_google("Roboto", "Roboto")
+# font_add_google("Roboto Mono", "Roboto Mono")
+# showtext_auto()
+#
+# theme_set(theme_light(base_size = 18, base_family = "Roboto"))
+# theme_set(theme_light(base_size = 18, base_family = "Poppins"))
+# theme_set(ggthemes::theme_fivethirtyeight(base_size = 18, base_family = "Poppins"))
+
+plot_links_year_reg <-
+  links_by_year_reg %>%
+  rename(Registry = registry) %>%
+  ggplot(aes(x = completion_year, y = prop_link, color = Registry, linetype = Registry)) +
+  facet_wrap(~link_type, ncol = 1) +
+  geom_point() +
+  geom_line() +
[…]

### 2021-08-21 analysis/fig_plot-upset-link-type.R

@@ -1,0 +1,96 @@
+# Prepare trials with links
+# Transform links into list column of intersection sets
+trials_links <-
+  trials %>%
+  select(id,
+         registry,
+         has_reg_pub_link,
+         has_iv_trn_secondary_id,
+         has_iv_trn_abstract,
+         has_iv_trn_ft
+  ) %>%
+  rename(
+    "TRN in full-text" = has_iv_trn_ft,
+    "TRN in abstract" = has_iv_trn_abstract,
+    "TRN in PubMed metadata" = has_iv_trn_secondary_id,
+    "Publication in registration" = has_reg_pub_link
+  ) %>%
+  pivot_longer(
+    cols = c(-id, -registry),
+    names_to = "link"
+  ) %>%
+  filter(value == TRUE) %>%
+  group_by(id) %>%
+  mutate(links = list(link)) %>%
+  ungroup() %>%
+  select(-value, -link) %>%
+  distinct()
+
+# Prepare trials without links
+# Create dummy links list column
+trials_no_links <-
+  trials %>%
+  filter(
+    !has_reg_pub_link &
+      !has_iv_trn_secondary_id &
+      !has_iv_trn_abstract &
+      !has_iv_trn_ft
+  ) %>%
+  select(id,
+         registry
+  ) %>%
+  mutate(links = list(NULL))
+
+plot_upset_links_reg_pub_registry <-
+  bind_rows(trials_links, trials_no_links) %>%
+  rename(Registry = registry) %>%
+  ggplot(aes(x = links)) +
+
+  # TODO: switch to "dodge" and fix numbers
+  geom_bar(aes(fill = Registry), position = "stack") +
+  geom_text(stat='count', aes(label = after_stat(count)), vjust = -.5) +
+  scale_y_continuous(expand = expansion(mult = c(0, .08))) +
+  # scale_fill_grey() +
+  scale_fill_manual(values = registry_colors) +
+  ggupset::scale_x_upset() +
+  ylab("Number of trials") +
+  xlab(NULL) +
+  ggupset::theme_combmatrix(
+    # combmatrix.label.make_space = FALSE,
+    combmatrix.panel.line.size = 0,
+    combmatrix.label.text = element_text(family = "Roboto", size = 11)
+  ) +
+
+  theme(
+    legend.background = element_rect(color = "transparent", fill = "transparent"),
+    legend.position = c(.85, .9),
[…]

### 2021-08-21 analysis/tbl_stakeholders.R

@@ -1,0 +1,106 @@
+# Prepare actions and examples --------------------------------------------
+
+
+action_researchers <- "
+- Include TRN in abstract and full-text per CONSORT and ICMJE guidelines
+- Link publication in registry
+"
+
+example_researchers <- "—"
+
+action_registries <- "
+- Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registry
+- Link publications automatically using TRNs in bibliographic database metadata
+- Provide metadata on linked publication type (i.e., result, background)
+"
+
+example_registries <- "ClinicalTrials.gov uses PubMed TRN metadata to automatically link publications in the registration (National Library of Medicine, 2017). Our study shows this automated publication linking is responsible for the majority of linked publications in ClinicalTrials.gov and implies similar potential for DRKS."
+# [@nationallibraryofmedicine2017c]
+
+action_institutions <- "
+- Educate, support, and incentivize researchers on linking
+"
+
+example_institutions <- "
+Ethics review offices or UMC-core facilities for clinical research could give researchers a info sheet with trial registration and reporting guidelines when reviewing trial. Bonuses, i.e., *Leistungsorientierte Mittel (LOM)*, could be disseminated for trial registration, reporting, and linking.
+"
+
+action_journals <- "
+- Request TRN in specialized metadata field
+- Review abstract and full-text for TRN inclusion, using automated regexes and/or manual strategies
+- Provide TRN as metadata to bibliographic databases
+"
+
+example_journals <- "
+Taylor & Francis extracts TRNs from the abstract and full-text, submit this metadata to CrossRef and Pubmed, and displays the linked trial via Crossmark on the article page (CrossRef, 2020; Taylor & Francis Group, 2020)
+"
+# [@taylorfrancisgroup2020; @crossref2020c]
+#
+action_pubmed <- "
+- Integrate publisher-provided TRNs into metadata
+- Extract TRN as metadata from abstract and full-text, using automated regexes and/or manual strategies
+"
+
+example_pubmed <- "
[…]

### 2021-08-21 analysis/tbl_tbl-trials.R

@@ -1,0 +1,173 @@
+# Summarize characteristics of trials and publications included in sample
+
+# Summarize all trial info ------------------------------------------------
+
+# tbl_trials_full <-
+#   trials %>%
+#
+#   # Add indicators for none and all
+#   mutate(
+#     has_trn_all = case_when(
+#       has_trn_secondary_id & has_trn_abstract & (has_trn_ft_pmc | has_trn_ft_pdf) ~ TRUE,
+#       is.na(has_trn_secondary_id) | is.na(has_trn_abstract) | (is.na(has_trn_ft_pmc) & is.na(has_trn_ft_pdf)) ~ NA,
+#       TRUE ~ FALSE
+#     ),
+#     has_trn_none = if_else(
+#       !has_trn_secondary_id & !has_trn_abstract & (!has_trn_ft_pmc | !has_trn_ft_pdf),
+#       TRUE, FALSE, missing = FALSE
+#     )
+#   ) %>%
+#
+#   select(
+#     has_summary_results,
+#     is_multicentric,
+#     enrollment,
+#     main_sponsor,
+#     # recruitment_status, # could have changed
+#     phase,
+#     masking,
+#     allocation,
+#     intervention_type,
+#     # lead_cities, # separate table/analysis?
+#     starts_with("days_"),
+#     starts_with("has_trn") & !ends_with("intovalue"),
+#   ) %>%
+#   gtsummary::tbl_summary(
+#     label = list(
+#       has_trn_any ~ "Anywhere",
+#       has_trn_all ~ "Metadata + Abstract + Full-text (PMC or PDF)",
+#       has_trn_none ~ "Nowhere",
+#       has_trn_secondary_id ~ "PubMed Metadata" ,
+#       has_trn_abstract ~ "Abstract",
+#       has_trn_ft_pmc ~ "PMC Full-Text",
+#       has_trn_ft_pdf ~ "PDF Full-Text"
+#     ),
+#     statistic =
+#       # all_logical() ~
+#       c("has_trn_any",
+#         "has_trn_all",
+#         "has_trn_none",
+#         "has_trn_secondary_id",
+#         "has_trn_abstract",
+#         "has_trn_ft_pmc",
+#         "has_trn_ft_pdf") ~
+#       "{n}/{N_nonmiss} ({p}%)"#,
+#     # missing = "no"
+#   ) %>%
+#     # modify_caption("**Trial registration number (TRN) reporting across published PubMed-indexed results from German UMC-led registered trials**")
[…]

### 2021-09-06 docs/manuscript/manuscript.Rmd
Move manuscript and update gitignore

--- 2021-08-21 manuscript/manuscript.Rmd
@@
 shorttitle        : "Registration-publication links in German clinical trials"
-
 author: 
@@
       - Supervision
-
 affiliation:
@@
     institution   : "QUEST Center for Responsible Research, Berlin Institute of Health (BIH) at Charité Universitätsmedizin Berlin"
-
 authornote: >
-
 abstract: |
-  **Objective** To evaluate, using a novel automated approach, links between registration and publication across clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
-The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make clinical trial publications and registrations findable and increase the value of clinical research.
+  **Objective** To evaluate links between registration and publication across clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+Inadequate links make trial publications and registrations less findable and compromise evidence synthesis and health policy decision making.
+The World Health Organization (WHO) and others call for better adoption of this straightforward transparency practice.
+<!--#The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make clinical trial publications and registrations findable and increase the value of clinical research.-->
 
@@
 We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
-<!--# We verified our automated approach on a manually coded sample.
--->
 
-  **Main outcome measures** The proportion of trial registrations incorporating a structured reference (DOI or PMID) to a known results publication.
-The proportion of clinical trial results publications that reported the known TRN in the metadata, abstract, and full-text.
+  **Main outcome measures** The proportion of trial registrations that reference a known results publication.
+The proportion of results publications that report the known TRN in the metadata, abstract, and full-text.
 
   **Secondary analyses** We constructed exploratory logistic regression models to investigate the relationship between trial completion date, registry, and registration-publication linking.
-<!--# We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.
--->
 
-  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata, and only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft_abs")` trials met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
-In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links.
+  **Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
+In addition to improved linking by trialists, changes in the PubMed TRN capturing process (such as automated strategies like those developed in this study) and automated indexing of publications in DRKS would make trial registrations and results more findable.
+
+  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata, and only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft_abs")` of trials met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
+`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links.
 The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft")`).
@@
 Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
-
-  **Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
-In addition to improved linking by trialists, changes in the PubMed TRN capturing process (such as the incorporation of automated strategies like those used in this study) and automated indexing of publications in DRKS would make trial registrations and results more findable.
-
   **Open Data and Code** All code and the final analysis data for this study are available at https://github.com/maia-sh/reg-pub-link.
-Raw data (with the exception of the full-text of publications) is available at https://github.com/maia-sh/intovalue-data.<!--#TODO:move data processing into intovalue-data repo--><!--#**Registration** This study was not preregistered.
-The study protocol is available at [insert link].-->
-
+Raw data is available at https://github.com/maia-sh/intovalue-data.
 keywords          : "clinical trials, registration, reporting, meta-research, research transparency"
-wordcount         : "3551"
-
+wordcount         : "3604"
 bibliography      : ["references.bib", "r-references.bib"]
@@
 draft             : no
-
 lang              : "en-US"
@@
 classoption       : "man"
-output            : papaja::apa6_word
-
+output            : papaja::apa6_pdf
 header-includes   :
   - \usepackage{orcidlink}
-
 always_allow_html: true
-
 nocite: |
   @taylorfrancisgroup2020, @crossref2020c, @nationallibraryofmedicine2017c
+editor_options: 
+  chunk_output_type: console
 ---
@@
 abs_n_trials_link_none <- 
-  report_trials(n_trials_link_none)
+  report_trials(n_trials_link_none, pattern = "{p}% ({n})")
 abs_n_trials_link_ft <- 
-  inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0, pattern = "{n}, {p}% of trials")
+  inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0, pattern = "{p}%, {n}")
 abs_n_trials_link_reg_ctgov_auto <- 
@@
 
-Clinical trial evidence informs clinical practice and health policy decision making through reviews and meta-analyses.
-The validity of health evidence synthesis is jeopardized if relevant trial data are either left out due to an inability to find results through automated means based on inadequate trial metadata, or duplicated due to an inability to connect publications and their registrations.
-Incomplete or duplicated evidence threatens the validity of clinical decision-making and potentially misinforms decisions regarding the launch of subsequent clinical trials, which may affect the wellbeing of patients enrolled in them [@chan2014tl; @councilforinternationalorganizationsofmedicalsciences2016]<!--# undermines the justification for insurance reimbursements, drives the misallocation of healthcare resources-->.
-Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp; @chalmers2013b].
-Linking trial registrations and results publications makes each more findable, reducing the risk of information loss or duplication and increasing clinical research value [@kubben2019]<!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->.
-Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].
+Linking trial registrations and results publications makes each more findable and improves research transparency at minimal effort to researchers.
+Threaded evidence also empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t].
+If trial results cannot be found or linked to their registration, systematic reviewers may miss relevant data and draw invalid conclusions [@miron2020sd].
+Health policy decisions and clinical guidelines rely on such evidence synthesis, and incomplete evidence can misinform subsequent clinical trials, drive the misallocation of healthcare resources, and risk patient wellbeing [@chan2014tl; @councilforinternationalorganizationsofmedicalsciences2016].
 
-Trial results should be bidirectionally linked, meaning the inclusion of both a structured reference to publications in the registration, and trial registration numbers (TRN) in the full-text, abstract, and metadata of publications.
-The Consolidated Standards of Reporting Trials (CONSORT) as well as the ICMJE's *Recommendations for the Conduct, Reporting, Editing, and Publication of Scholarly Work in Medical Journals* call for reporting "trial registration number and name of the trial register" in both the full-text and abstract of trial results publication [@internationalcommitteeofmedicaljournaleditorsicmje2019; @hopewell2008pm].
-Reporting TRN solely in the full-text does not suffice since readers may not have access to the full-text or may base their judgment of the trial or topic on the abstract alone [@hopewell2008tl].
-Abstracts may also be published independently of a full-text publication, such as for conference abstracts [@hopewell2008pm].
+To inform clinical decision making with comprehensive trial evidence, legal regulations and ethical guidelines advocate results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp; @chalmers2013b].
+Trial results should be bidirectionally linked, meaning both a reference to publications in the registration, and trial registration numbers (TRNs) in the full-text, abstract, and metadata of publications.
+The Consolidated Standards of Reporting Trials (CONSORT) as well as the International Committee of Medical Journal Editors (ICMJE) ask trialists to report the "trial registration number and name of the trial register" in both the full-text and abstract of trial results publications [@internationalcommitteeofmedicaljournaleditorsicmje2019; @hopewell2008pm].
+Reporting a TRN solely in the full-text does not suffice, since readers may not have access to the full-text or may judge a trial on the abstract alone [@hopewell2008tl].
+Abstracts may also be published independently of a full-text publication, such as for conferences [@hopewell2008pm].
 
@@
 Such references to results publications within a registration allow readers to quickly identify and navigate to the trial findings.
-<!--# add policy background on why pub in reg --> Results references may also be entered manually in many registries, with varying degrees of structure.
+Results references may also be entered manually in selected registries, with varying degrees of structure [@miron2020sd; @moja2009t; @venugopal2021po].
 ClinicalTrials.gov, for example, provides fields for DOI and PubMed identifier (PMID), whereas the German Clinical Trials Registry (DRKS) offers an unstructured free-text field.
-<!--# add some background on ctgov vs.
-drks.
-and how they include trns and publications.
-and ctgov and pubmed by same nlm so may be better at seing those ids -->While readers of the registry may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval [@huser2013po].
+While registry readers may be able to find the referenced publication from the free-text, unique identifiers provide a more robust and structured link that allows for automated retrieval [@huser2013po].
 
@@
 The sampling strategies used in previous studies limit our ability to draw conclusions about registration-publication linking.
-Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included non-trial-result publications, and trials with unknown registration rates (as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo]); registration-based cohorts included included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b].
-Furthermore, these studies of publication links in registrations have primarily focused on ClinicalTrials.gov, and we are not aware of previous studies investigating publication links in DRKS.
-In this study, we limit our sample to registered clinical trials with published results which can be expected to link the registration and publication, and which allows us to draw conclusions about the prevalence of this responsible research practices.
+Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included publications beyond trial results, as well as trials with unknown registration rates, as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo].
+Registration-based cohorts included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b].
+Furthermore, these studies of publication links in registrations have primarily focused on ClinicalTrials.gov, and we are not aware of previous studies that investigate publication links in DRKS.
+In this study, we limit our sample to registered clinical trials with published results which can be expected to link the registration and publication, and which allows us to draw conclusions about the prevalence of this responsible research practice.
 
@@
 
-The exploratory study evaluates linking of registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or DRKS with completion dates between 2009 and 2017.
-We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions for the detection and classification of publication identifiers and TRNs, which may be applied to other trial cohorts.
+In this exploratory study, we evaluate the links between registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or DRKS with completion dates between 2009 and 2017.
+We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions to detect and classify publication identifiers and TRNs, which may be applied to other trial cohorts.
 <!--# We did not pre-specify the analyses.
@@
 We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021c and referred to as the "IntoValue" dataset [@riedel2021a].
-The dataset consists of clinical trials registered on ClinicalTrials.gov or DRKS and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.
+The dataset consists of clinical trials registered on ClinicalTrials.gov or DRKS, conducted at a German UMC, and completed between 2009 and 2017.
+Corresponding results publications were found via manual searches.
 
-We <!--# programmatically -->downloaded data from both registries on `r ifelse(aact_query_date == drks_query_date, aact_query_date, stop())`.
-<!--# We fetched additional fields needed for our analysis, such as references, as well as refreshed the fields previously included in our source trial datasets to ensure parity between the two trial cohorts and current registry data.
--->We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti].
+We downloaded data from both registries on `r ifelse(aact_query_date == drks_query_date, aact_query_date, stop())`.
+We queried ClinicalTrials.gov using the Clinical Trials Transformation Initiative (CTTI) Aggregate Content of ClinicalTrials.gov (AACT) via its PostgreSQL database API [@clinicaltrialstransformationinitiativectti].
 As DRKS does not provide an API, we built a webscraper to capture the necessary fields [@federalinstitutefurdrugsandmedicaldevicesbfarm].
-<!--# ClinicalTrials.gov requires more structured data input compared with DRKS; we attempted to tidy the DRKS output to extract additional information.
-For example, ClinicalTrials.gov has a separate summary results reporting section, whereas summary results are included with other references in DRKS.
-We inferred summary results based on common terminology (e.g., "study results", "Ergebnisbericht", "Abschlussbericht", "Studienergebnisse").-->
 
-We queried the PubMed Entrez Programming Utilities (E-utilities) API on `r pubmed_query_date` for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @R-rentrez].
+We queried the PubMed Entrez Programming Utilities (E-utilities) API on `r pubmed_query_date` for all trial results PMIDs<!--#, and excluded PMIDs that did not resolve--> [@usnationallibraryofmedicinetigtand; @R-rentrez].
 From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->
@@
 
-After updating registry data, we reapplied the IntoValue exclusion criteria (completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead; see @riedel2021c for further details on these criteria) and limited our sample to trials with results with a PMID and full-text publication.
+After updating registry data, we reapplied the IntoValue exclusion criteria: completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead.
+See @riedel2021c for further details on these criteria.
+We limited our sample to trials with results with a PMID and full-text publication.
 
@@
 
-As publication data is available in different formats depending on the registry, we developed parallel approaches to extract publication identifiers for each.
+As each registry formats references to publications differently, we developed parallel approaches to extract publication identifiers for each.
 For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT.
-We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field.
+We then used a regular expression (regex) to extract any DOIs from the citations.
 For DRKS, we scraped the reference type, citation, and URL, when available.
@@
 These additional TRNs could be cross-registrations of the same trial or provided as background or discussion.
-For this study, we were interested in whether known registrations were reported in the publication and so filtered for reported TRNs matching the known registrations.
-<!--# change this if considering cross-reg -->
+For this study, we were interested in whether the known registrations were reported in the publication.
 
@@
 The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_trials(n_trials_link_ft_only)` trials.
-The inclusion of the various link types ranged from `r inline_text(tbl_link_by_reg, has_iv_trn_abstract, stat_0, pattern = "{p}% ({n}/{N})")` in abstracts to `r inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0, pattern = "{p}% ({n}/{N})")` in full-text.
+The inclusion of the various link types ranged from `r inline_text(tbl_link_by_reg, has_iv_trn_abstract, stat_0)` in abstracts to `r inline_text(tbl_link_by_reg, has_iv_trn_ft, stat_0)` in full-text.
 Table \@ref(tab:tbl-link-by-reg) shows registration-publication links overall and by registry.
@@
 Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables.
-Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
+Completion year did not have a strong relationship with linking practices, although trials completed more recently were more likely to report the TRN in the abstract (`r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`).
 Figure \@ref(fig:plot-link-year) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.
@@
 
-Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with link type were generally more likely to have another type of link.
+Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with one link type were generally more likely to have another type of link.
 <!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")`).
-Futhermore, DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
+DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
 Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (`r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_1", pattern = "{p}%")` vs.
@@
 
-Trials registered in DRKS were also less likely than trials registered in ClinicalTrials.gov to reference the full-text primary outcome publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
-<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) ClinicalTrials.gov trials referencing the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
+Trials registered in DRKS were also less likely than trials registered in ClinicalTrials.gov to reference the primary outcome publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
+<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) ClinicalTrials.gov trials that reference the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
 Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")` vs.
@@
 We also found `r si_drks_misformatted` ids which we manually classified as true negative non-TRNs; however, using additional PubMed data, we determined these were severely misformatted DRKS ids (i.e., missing the preceding letters "DRKS" and just a string of numbers such as "00000711").
-Our regexes correctly classified all well formatted TRNs and non-TRNs, resulting in a sensitivity and specificity of `r ifelse(si_sens == si_spec, style_percent(si_sens), stop())`%.
+Our regexes correctly classified all well-formatted TRNs and non-TRNs, resulting in a sensitivity and specificity of `r ifelse(si_sens == si_spec, style_percent(si_sens), stop())`%.
 If we had instead categorized the `r si_drks_misformatted` severely misformatted DRKS ids as true positives, our regexes would have sensitivity of `r style_percent(si_sens_alt, digits = 1)`% and a specificity of `r style_percent(si_spec_alt)`%.
@@
 
-Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of research.
-Poor linking poses a barrier to identifying trial publications via automated approaches and instead requires researchers to perform intensive manual searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
+Linking of trial registrations and results publications plays an important role in research transparency and facilitates comprehensive evidence synthesis and informed health policy decision making.<!--# and can substantially increase the value of research.
+--> Poor linking poses a barrier to identifying trial publications via automated approaches and instead requires researchers to perform intensive manual searches to attempt to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
 This responsible research practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.
@@
 Our study shows that German UMCs can improve in both TRN reporting in publications and references to publications in the registration.
-In our sample (n = `r n_trials_ft`), `r report_trials(n_trials_link_none)` trials had no links between registration and publication and only `r report_trials(n_trials_link_full)` trials had the most comprehensive registration-publication links.
-Furthermore, only `r report_trials(n_trials_link_ft_abs)` trials in our sample fully met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
+In our sample (n = `r n_trials_ft`), `r report_trials(n_trials_link_none, pattern = "{p}% ({n})")` of trials had no links between registration and publication and only `r report_trials(n_trials_link_full, pattern = "{p}% ({n})")` of trials had the most comprehensive registration-publication links.
+Furthermore, only `r report_trials(n_trials_link_ft_abs, pattern = "{p}% ({n})")` of trials in our sample fully met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
 Linking practices showed at best minimal improvement over time.
@@
 Automated indexing of publications in ClinicalTrials.gov accounts for most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) publication references in the registry and drives the almost three-fold discrepancy with DRKS (`r inline_text(tbl_link_by_reg, has_reg_pub_link, "ClinicalTrials.gov", pattern = "{p}%, {n}/{N}")` vs.
-`r inline_text(tbl_link_by_reg, has_reg_pub_link, "DRKS", pattern = "{p}%, {n}/{N}")`).<!--# Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")`)--> Currently, DRKS allows  for only manual submission of references by trialists and does not index publications.<!--# add?--> Furthermore, registrations may refer to additional publications beyond results (e.g., up to 72 publications for one ClinicalTrials.gov registration in our sample), so registry metadata should encode publication type to support quick identification of results; while such publication type metadata is currently available in both registries, it is not systematically used, and many publications are categorized generically as a "paper" (DRKS) or "reference" (ClinicalTrials.gov).
+`r inline_text(tbl_link_by_reg, has_reg_pub_link, "DRKS", pattern = "{p}%, {n}/{N}")`).<!--# Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")`)--> Currently, DRKS allows for only manual submission of references by trialists and does not index publications.<!--# add?--> Furthermore, registrations may refer to additional publications beyond results (e.g., up to 72 publications for one ClinicalTrials.gov registration in our sample), so registry metadata should encode publication type to support quick identification of results.
+While such publication type metadata is currently available in both registries, it is not systematically used, and many publications are categorized generically as a "paper" (DRKS) or "reference" (ClinicalTrials.gov).
 
@@
 remove since not about trn formatting alone?
-Futhermore, ClinicalTrials.gov presumably relies not only on the reported TRN but also on the registry, and as such failed to index a publication (PMID: 26417987) despite the TRN (NCT01614301) appearing in the PubMed metadata, as the associated databank was incorrect (i.e., GENBANK).
+Furthermore, ClinicalTrials.gov presumably relies not only on the reported TRN but also on the registry, and as such failed to index a publication (PMID: 26417987) despite the TRN (NCT01614301) appearing in the PubMed metadata, as the associated databank was incorrect (i.e., GENBANK).
 -->
@@
 In contrast with previous studies relying on PubMed queries to identify potential (randomized) clinical trial results publications, we relied on a sample of bona fide results publications from registered trials, allowing us to evaluate the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.
-Futhermore, we used an automated approach including regular expressions with high sensitivity and specificity to identify and classify publication identifiers and TRNs which allowed for a larger sample size that manual data extraction would have permitted.
+Furthermore, we used an automated approach including regular expressions with high sensitivity and specificity to identify and classify publication identifiers and TRNs which allowed for a larger sample size that manual data extraction would have permitted.
 This automated strategy is scalable and can be applied to other trial sets.
@@
 Our input IntoValue dataset comprised German UMC-led trials in ClinicalTrials.gov and DRKS and may not reflect practices across other registries and/or countries.
-Future projects should look at multi-national samples of bona fide results publications.
+Future projects should look at multinational samples of bona fide results publications.
 Furthermore, we relied on IntoValue for trial deduplication and bona fide results publications.
@@
 In our sample alone, an additional `r report_trials(n_trials_link_ft_or_abs_no_si, n_trials_link_ft_or_abs, pattern = "{p}% ({n}/{N})")` of trials with TRN in abstract or full-text could have the TRN included in PubMed's metadata, and an additional `r report_trials(n_trials_drks_link_si_not_reg, n_trials_drks_link_si, pattern = "{p}% ({n}/{N})")` of DRKS trials with TRN in the PubMed metadata could be automatically indexed in the registration.
-With the uptake of the recommended actions across stakeholder groups, improved trial registration-publication links will allow for more comprehensive evidence synthesis and increase the value of clinical research.
+By adopting the recommended actions, stakeholders can improve trial registration-publication links and foster more comprehensive evidence synthesis and well-informed clinical guidelines and health policy decisions.
 

### 2021-09-06 analysis/tbl_stakeholders.R
Update stakeholder table

@@ -3,13 +3,13 @@
 
 action_researchers <- "
 - Include TRN in abstract and full-text per CONSORT and ICMJE guidelines
-- Link publication in registry
+- Link publication in registration
 "
 
 example_researchers <- "—"
 
 action_registries <- "
-- Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registry
+- Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registration
 - Link publications automatically using TRNs in bibliographic database metadata
 - Provide metadata on linked publication type (i.e., result, background)
 "
@@ -22,7 +22,7 @@ action_institutions <- "
 "
 
 example_institutions <- "
-Ethics review offices or UMC-core facilities for clinical research could give researchers a info sheet with trial registration and reporting guidelines when reviewing trial. Bonuses, i.e., *Leistungsorientierte Mittel (LOM)*, could be disseminated for trial registration, reporting, and linking.
+Ethics review offices or UMC-core facilities for clinical research could give researchers an info sheet with trial registration and reporting guidelines when reviewing trial. Bonuses, i.e., *Leistungsorientierte Mittel (LOM)*, could be disseminated for trial registration, reporting, and linking.
 "
 
 action_journals <- "
@@ -32,7 +32,7 @@ action_journals <- "
 "
 
 example_journals <- "
-Taylor & Francis extracts TRNs from the abstract and full-text, submit this metadata to CrossRef and Pubmed, and displays the linked trial via Crossmark on the article page (CrossRef, 2020; Taylor & Francis Group, 2020)
+Taylor & Francis extracts TRNs from the abstract and full-text, submits this metadata to CrossRef and Pubmed, and displays the linked trial via Crossmark on the article page (CrossRef, 2020; Taylor & Francis Group, 2020)
 "
 # [@taylorfrancisgroup2020; @crossref2020c]
 #

### 2021-09-06 docs/manuscript/manuscript.Rmd
Add Till Bruckner acknowledgement
Clarify German UMC lead/conducted by and industry sponsor

@@
 abstract: |
-  **Objective** To evaluate links between registration and publication across clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+  **Objective** To evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
 Inadequate links make trial publications and registrations less findable and compromise evidence synthesis and health policy decision making.
@@
 
-In this exploratory study, we evaluate the links between registrations and results publications in a cross-section of published clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or DRKS with completion dates between 2009 and 2017.
+In this exploratory study, we evaluate the links between registrations and results publications in a cross-section of published clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or DRKS with completion dates between 2009 and 2017.
 We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions to detect and classify publication identifiers and TRNs, which may be applied to other trial cohorts.
@@
 We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021c and referred to as the "IntoValue" dataset [@riedel2021a].
-The dataset consists of clinical trials registered on ClinicalTrials.gov or DRKS, conducted at a German UMC, and completed between 2009 and 2017.
+The dataset consists of clinical trials registered on ClinicalTrials.gov or DRKS, conducted by a German UMC, and completed between 2009 and 2017.
 Corresponding results publications were found via manual searches.
@@
 
-After updating registry data, we reapplied the IntoValue exclusion criteria: completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead.
+After updating registry data, we reapplied the IntoValue inclusion criteria: completion date between 2009 and 2017, interventional, complete based on study status, and conducted by a German UMC.
+Trials were considered to be conducted by a German UMC if a UMC was included as a trial sponsor, overall official, and/or responsible party in ClinicalTrials.gov or provided in study addresses in DRKS; trials with a German UMC as only a facility in ClinicalTrials.gov or recruitment location in DRKS were excluded.
 See @riedel2021c for further details on these criteria.
@@
 
-The IntoValue dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
+The IntoValue dataset includes all trials conducted by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`).
 After applying our exclusion criteria, our sample included `r n_trials_ft` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
@@
 
-<!-- (ref:tbl-trials-caption) **Characteristics of German UMC-led trials with published results** A trial was considered randomized if allocation included randomization.
+<!-- (ref:tbl-trials-caption) **Characteristics of German UMC-conducted trials with published results** A trial was considered randomized if allocation included randomization.
 A trial was considered prospectively registered if registered in the same or previous months to start date.
@@
 The approach also faces limitations.
-Our input IntoValue dataset comprised German UMC-led trials in ClinicalTrials.gov and DRKS and may not reflect practices across other registries and/or countries.
+Our input IntoValue dataset comprised trials conducted by German UMCs and registered in ClinicalTrials.gov and DRKS and may not reflect practices across other registries and/or countries.
 Future projects should look at multinational samples of bona fide results publications.
@@
 
-This study reveals shortcomings in the linking of German UMC-led trials registrations and results publications and highlights several promising avenues forward.
+This study reveals shortcomings of German UMCs in linking trial registrations and results publications and highlights several promising avenues forward.
 In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is straightforward and can be improved with action across stakeholders.
@@
 We would like to thank Kerstin Rubarth from the Charité Institut für Biometrie und Klinische Epidemiologie for her consultation on the statistical analysis, and Nicholas Devito for earlier versions of the trial registration number regular expressions.
+We also thank our colleague Till Bruckner for useful comments on an earlier version of this manuscript.
 

### 2021-09-06 analysis/tbl_tbl-trials.R

@@ -156,7 +156,7 @@ tbl_trials <-
   add_stat_label() %>%
 
   # modify_header(label = "**German UMC-led trials with published results**") %>%
-  modify_caption("**Characteristics of German UMC-led trials with published results** A trial was considered randomized if allocation included randomization. A trial was considered prospectively registered if registered in the same or previous months to start date. Summary results were taken from a structured data field in ClinicalTrials.gov, and determined based on manual inspection for terms such as *Ergebnisbericht* or *Abschlussbericht* in DRKS. *Top* journals refer to the journals with the greatest number of trial publications in our sample.") %>%
+  modify_caption("**Characteristics of German UMC-conducted trials with published results** A trial was considered randomized if allocation included randomization. A trial was considered prospectively registered if registered in the same or previous months to start date. Summary results were taken from a structured data field in ClinicalTrials.gov, and determined based on manual inspection for terms such as *Ergebnisbericht* or *Abschlussbericht* in DRKS. *Top* journals refer to the journals with the greatest number of trial publications in our sample.") %>%
   bold_labels() %>%
 
   # Remove rowname label

### 2021-09-06 docs/manuscript/manuscript.Rmd
Clarify study completion date

@@
 
-After updating registry data, we reapplied the IntoValue inclusion criteria: completion date between 2009 and 2017, interventional, complete based on study status, and conducted by a German UMC.
+After updating registry data, we reapplied the IntoValue inclusion criteria: study completion date between 2009 and 2017, interventional, complete based on study status, and conducted by a German UMC.
 Trials were considered to be conducted by a German UMC if a UMC was included as a trial sponsor, overall official, and/or responsible party in ClinicalTrials.gov or provided in study addresses in DRKS; trials with a German UMC as only a facility in ClinicalTrials.gov or recruitment location in DRKS were excluded.

### 2021-09-06 docs/manuscript/manuscript.Rmd
Add acknowledgement ct.gov manual pub link floor

@@
 Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")` vs.
-`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_2", pattern = "{p}%, {n}/{N}")`).
+`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_2", pattern = "{p}%, {n}/{N}")`); the number of manually linked publications in ClinicalTrials.gov may, however, be an underestimate of researcher efforts, since a researcher may attempt to manually link a publication in ClinicalTrials.gov, only to see it was already automatically indexed.
 Trials also had additional publications linked in the registrations, up to `r reg_pub_max_ctgov$n_reg_pub_any` in ClinicalTrials.gov (`r reg_pub_max_ctgov$id`) and `r reg_pub_max_drks$n_reg_pub_any` in DRKS (`r reg_pub_max_drks$id`).

### 2021-09-06 docs/manuscript/manuscript.Rmd
Mention protocols as non-results publication

@@
 The Consolidated Standards of Reporting Trials (CONSORT) as well as the International Committee of Medical Journal Editors (ICMJE) ask trialists to report the "trial registration number and name of the trial register" in both the full-text and abstract of trial results publications [@internationalcommitteeofmedicaljournaleditorsicmje2019; @hopewell2008pm].
-Reporting a TRN solely in the full-text does not suffice, since readers may not have access to the full-text or may judge a trial on the abstract alone [@hopewell2008tl].
+Reporting a TRN solely in the full-text does not suffice, since readers may not have access to the full-text or may screen a trial on the abstract alone [@hopewell2008tl].
 Abstracts may also be published independently of a full-text publication, such as for conferences [@hopewell2008pm].
@@
 The sampling strategies used in previous studies limit our ability to draw conclusions about registration-publication linking.
-Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included publications beyond trial results, as well as trials with unknown registration rates, as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo].
+Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included publications beyond trial results (e.g., protocols), as well as trials with unknown registration rates, as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo].
 Registration-based cohorts included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b].

### 2021-09-06 analysis/tbl_stakeholders.R

@@ -11,7 +11,7 @@ example_researchers <- "—"
 action_registries <- "
 - Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registration
 - Link publications automatically using TRNs in bibliographic database metadata
-- Provide metadata on linked publication type (i.e., result, background)
+- Provide metadata on linked publication type (i.e., result, background, protocol)
 "
 
 example_registries <- "ClinicalTrials.gov uses PubMed TRN metadata to automatically link publications in the registration (National Library of Medicine, 2017). Our study shows this automated publication linking is responsible for the majority of linked publications in ClinicalTrials.gov and implies similar potential for DRKS."

### 2021-09-06 docs/manuscript/manuscript.Rmd
Add ctgov results publication tutorial

@@
 nocite: |
-  @taylorfrancisgroup2020, @crossref2020c, @nationallibraryofmedicine2017c
+  @taylorfrancisgroup2020, @crossref2020c, @nationallibraryofmedicine2017c, @usnationallibraryofmedicine2020a
 editor_options: 
@@
 # Prepare directory
-dir_figures <- dir_create(here("manuscript", "figures"))
+dir_figures <- dir_create(here("docs", "manuscript", "figures"))
 

### 2021-09-06 analysis/tbl_stakeholders.R

@@ -6,7 +6,7 @@ action_researchers <- "
 - Link publication in registration
 "
 
-example_researchers <- "—"
+example_researchers <- "ClinicalTrials.gov provides a step-by-step tutorial for linking publications in the registration (U.S. National Library of Medicine, 2020)."
 
 action_registries <- "
 - Provide guidance on TRN formatting and reminder to include TRN in publications, and publications in registration

### 2021-09-06 docs/manuscript/manuscript.Rmd
Update appendix format to supplemental material

@@
 After applying our exclusion criteria, our sample included `r n_trials_ft` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial.
-Appendix \@ref(app-flow-trials) provides a flow diagram of the trial and publication screening.
+Supplemental Figure \@ref(fig:app-fig-flow-trials) provides a flow diagram of the trial and publication screening.
 Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.
@@
 `r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_2", pattern = "{p}%")`).
-Appendix \@ref(app-si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
+Supplemental Table \@ref(tab:app-tbl-si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.
 

### 2021-09-06 docs/manuscript/manuscript.Rmd
Add analysis of trials 2014 on for drks pubmed start date

@@
 <!--# point estimates in table should have CI, right?
--->Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the intersection of registration-publication links across the various types for ClinicalTrial.gov and DRKS trials.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.-->
+-->Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the intersection of registration-publication links across the various types for ClinicalTrial.gov and DRKS trials.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.--> As PubMed incorporated DRKS as a databank source in 2014, Supplemental Table \@ref(tab:app-tbl-link-by-reg-2014) shows the linking practices in trials published as of 2014 (n = `r inline_text(tbl_link_by_reg_2014, has_iv_trn_secondary_id, stat_0, pattern = "{N}")`), which reflect similar rates to Table \@ref(tab:tbl-link-by-reg) [@nationallibraryofmedicine2018].
 

### 2021-09-06 analysis/analysis_reg-pub-link.R

@@ -158,6 +158,34 @@ tbl_link_by_reg <-
   # add_overall(col_label = "**Overall**\nN = {N}") %>%
   modify_footnote(everything() ~ NA)
 
+# Prevalence of link types, overall and by registry
+# Trials published as of 2014 to account for MEDLINE incorporation of DRKS
+# https://www.nlm.nih.gov/bsd/medline_databank_source.html
+tbl_link_by_reg_2014 <-
+  trials %>%
+  filter(publication_date > as.Date("2013-12-31")) %>%
+  select(
+    registry,
+    has_iv_trn_ft,
+    has_iv_trn_abstract,
+    has_iv_trn_secondary_id,
+    has_reg_pub_link
+  ) %>%
+  gtsummary::tbl_summary(
+    by = registry,
+    label = list(
+      has_iv_trn_ft ~ "TRN in full-text",
+      has_iv_trn_abstract ~ "TRN in abstract",
+      has_iv_trn_secondary_id ~ "TRN in PubMed metadata",
+      has_reg_pub_link ~ "Publication in registration"
+    )
+  ) %>%
+  add_overall() %>%
+  modify_header(label = "") %>%
+  modify_caption("Registration-publication links overall and by registry for trials published as of 2014") %>%
+  bold_labels() %>%
+  modify_footnote(everything() ~ NA)
+
 # Manually-linked publication in registration -----------------------------
 
 tbl_reg_pub_link_manual <-

### 2021-09-07 docs/manuscript/manuscript.Rmd
Add clinical trials abstract

@@
 abstract: |
+  *ORIGINAL*
+
   **Objective** To evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
@@
 
-  **Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
-In addition to improved linking by trialists, changes in the PubMed TRN capturing process (such as automated strategies like those developed in this study) and automated indexing of publications in DRKS would make trial registrations and results more findable.
-
     **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata, and only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft_abs")` of trials met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
@@
 Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
+
+  **Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
+In addition to improved linking by trialists, changes in the PubMed TRN capturing process (such as automated strategies like those developed in this study) and automated indexing of publications in DRKS would make trial registrations and results more findable.
+
   **Open Data and Code** All code and the final analysis data for this study are available at https://github.com/maia-sh/reg-pub-link.
 Raw data is available at https://github.com/maia-sh/intovalue-data.
+
+  *CLINICAL TRIALS*
+
+  **Background/Aims** Informed clinical guidance and health policy relies on the ability of clinicians, policymakers, and guideline developers to find comprehensive clinical evidence and interlink components of the same clinical trial.
+To support the finding and linking of trial evidence, the World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) ask researchers to provide the trial registration number (TRN) in their publication and a reference to the publication in the registration.
+This straightforward transparency practice costs researchers minimal effort and makes evidence synthesis more thorough and efficient.
+The aim of this cross-sectional study is to evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+Secondary aims are to develop an automated pipeline that can be applied to other cohorts of trial registrations and publications, and to provide stakeholders, from trialists to registries, with concrete guidance to improve trial links.
+
+  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and results publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017.
+We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
+  We evaluated the proportion of trial registrations that reference a known results publication, and the proportion of results publications that report the known TRN in the metadata, abstract, and full-text.
+We also constructed exploratory logistic regression models to investigate the relationship between trial completion date, registry, and registration-publication linking.
+
+  **Results, positive or negative?
+which values to include?** `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_not_both_ft_abs")` of trials fail to reference TRNs in both the full-text and abstract, in breach of long-standing CONSORT and ICMJE guidelines.
+`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg")` of trial registrations do not contain links to their results publications.
+`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links, so registration and publication could be associated only by manual searching and screening.
+Overall, ClinicalTrials.gov trials were better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists.
+Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata.
+Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
+
+  **Conclusions** German UMCs have not comprehensively linked trial registrations and publications, despite established recommendations.
+This shortcoming threatens the quality of evidence synthesis and medical practice, and  burdens researchers to manually search and link trial data.
+Researchers could easily improve this by copy-and-pasting references between their trial registrations and publications.
+Other stakeholders could make trial evidence even more findable and interlinked, e.g., PubMed could capture additional TRNs using automated strategies (like those developed in this study), and DRKS could automatically index publications from PubMed.
+
 keywords          : "clinical trials, registration, reporting, meta-research, research transparency"
@@
   report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")
+abs_n_trials_link_reg <- 
+  inline_text(tbl_link_by_reg, has_reg_pub_link, stat_0, pattern = "{p}% ({n})")
+abs_n_trials_link_not_both_ft_abs <- 
+  report_trials(nrow(trials) - n_trials_link_ft_abs, pattern = "{p}% ({n})")
 ```

### 2021-09-07 docs/manuscript/manuscript.Rmd
Change upset plot, remove registry and change to percent

@@
 <!--# point estimates in table should have CI, right?
--->Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the intersection of registration-publication links across the various types for ClinicalTrial.gov and DRKS trials.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.--> As PubMed incorporated DRKS as a databank source in 2014, Supplemental Table \@ref(tab:app-tbl-link-by-reg-2014) shows the linking practices in trials published as of 2014 (n = `r inline_text(tbl_link_by_reg_2014, has_iv_trn_secondary_id, stat_0, pattern = "{N}")`), which reflect similar rates to Table \@ref(tab:tbl-link-by-reg) [@nationallibraryofmedicine2018].
+-->Figure \@ref(fig:plot-upset-links-reg-pub) shows the percentage of trials with each combination of links between registration and publication.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.--> As PubMed incorporated DRKS as a databank source in 2014, Supplemental Table \@ref(tab:app-tbl-link-by-reg-2014) shows the linking practices in trials published as of 2014 (n = `r inline_text(tbl_link_by_reg_2014, has_iv_trn_secondary_id, stat_0, pattern = "{N}")`), which reflect similar rates to Table \@ref(tab:tbl-link-by-reg) [@nationallibraryofmedicine2018].
 
@@
 
-```{r plot-upset-links-reg-pub-registries, fig.cap = "Trials with all combinations of links between registration and publication in ClinicalTrials.gov and DRKS", out.width="80%"}
+```{r plot-upset-links-reg-pub, fig.cap = "Percentage trials with each combination of links between registration and publication", out.width="80%"}
 knitr::include_graphics(path(dir_figures, "plot-upset-link-type.pdf"))

### 2021-09-07 analysis/fig_plot-upset-link-type.R

@@ -3,7 +3,6 @@
 trials_links <-
   trials %>%
   select(id,
-         registry,
          has_reg_pub_link,
          has_iv_trn_secondary_id,
          has_iv_trn_abstract,
@@ -15,10 +14,7 @@ trials_links <-
     "TRN in PubMed metadata" = has_iv_trn_secondary_id,
     "Publication in registration" = has_reg_pub_link
   ) %>%
-  pivot_longer(
-    cols = c(-id, -registry),
-    names_to = "link"
-  ) %>%
+  pivot_longer(cols = -id, names_to = "link") %>%
   filter(value == TRUE) %>%
   group_by(id) %>%
   mutate(links = list(link)) %>%
@@ -36,24 +32,25 @@ trials_no_links <-
       !has_iv_trn_abstract &
       !has_iv_trn_ft
   ) %>%
-  select(id,
-         registry
-  ) %>%
+  select(id) %>%
   mutate(links = list(NULL))
 
-plot_upset_links_reg_pub_registry <-
+plot_upset_links_reg_pub <-
   bind_rows(trials_links, trials_no_links) %>%
-  rename(Registry = registry) %>%
   ggplot(aes(x = links)) +
-
-  # TODO: switch to "dodge" and fix numbers
-  geom_bar(aes(fill = Registry), position = "stack") +
-  geom_text(stat='count', aes(label = after_stat(count)), vjust = -.5) +
-  scale_y_continuous(expand = expansion(mult = c(0, .08))) +
-  # scale_fill_grey() +
-  scale_fill_manual(values = registry_colors) +
+  geom_bar() +
+  geom_text(
+    stat = 'count',
+    aes(label = scales::percent(after_stat(count)/nrow(trials), accuracy = 0.1)),
+    vjust = -.5,
+    size = 3.5) +
+  scale_y_continuous(
+    label = scales::label_percent(scale = 100/nrow(trials), accuracy = 1),
+    breaks = scales::breaks_width(nrow(trials)/20),
+    expand = expansion(mult = c(0, .05))
+  ) +
   ggupset::scale_x_upset() +
-  ylab("Number of trials") +
+  ylab("Percentage of trials") +
   xlab(NULL) +
   ggupset::theme_combmatrix(
     # combmatrix.label.make_space = FALSE,
@@ -67,7 +64,6 @@ plot_upset_links_reg_pub_registry <-
[…]

### 2021-09-07 docs/manuscript/manuscript.Rmd
Change citation style to sage vancouver for clinical trials journal

@@
 always_allow_html: true
+csl: sage-vancouver.csl
 nocite: |

### 2021-09-07 docs/manuscript/manuscript.Rmd
Ignore deprecated scripts from git and analyses
Fix typo

--- 2021-09-07 docs/manuscript/manuscript.Rmd
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
 header-includes   :
@@
 ```{r analysis, cache = TRUE}
-dir_ls(here("analysis")) %>%
+dir_ls(here("analysis"), type = "file") %>%
   walk(source)
@@
 
-```{r plot-upset-links-reg-pub, fig.cap = "Percentage trials with each combination of links between registration and publication", out.width="80%"}
+```{r plot-upset-links-reg-pub, fig.cap = "Percentage of trials with each combination of links between registration and publication", out.width="80%"}
 knitr::include_graphics(path(dir_figures, "plot-upset-link-type.pdf"))

### 2021-09-08 README.Rmd
Add readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

r_citations <- papaja::cite_r(here::here("docs", "manuscript", "r-references.bib"))
```

This repository contains code for the project "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers." The manuscript is computationally reproducible using this repository. Note that a few manual formatting tweaks to tables and figures were done for the final [manuscript](https://doi.org/10.1101/2021.08.23.21262478).

To reproduce the manuscript:

1.  Clone this repository
2.  Run `renv::restore()` to ensure you have the necessary packages and versions
3.  Knit [manuscript.Rmd](https://github.com/maia-sh/reg-pub-link/blob/master/docs/manuscript/manuscript.Rmd)

This project uses `r r_citations`.

### 2021-09-08 docs/manuscript/manuscript.Rmd
Remove old abstract, edit data/code section

@@
 abstract: |
-  *ORIGINAL*
-
-  **Objective** To evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
-Inadequate links make trial publications and registrations less findable and compromise evidence synthesis and health policy decision making.
-The World Health Organization (WHO) and others call for better adoption of this straightforward transparency practice.
-<!--#The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make clinical trial publications and registrations findable and increase the value of clinical research.-->
-
-  **Design** Cross-sectional bibliographic study
-
-  **Setting** German UMC clinical trials
-
-  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017.
-We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
-
-  **Main outcome measures** The proportion of trial registrations that reference a known results publication.
-The proportion of results publications that report the known TRN in the metadata, abstract, and full-text.
-
-  **Secondary analyses** We constructed exploratory logistic regression models to investigate the relationship between trial completion date, registry, and registration-publication linking.
-
-    **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata, and only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft_abs")` of trials met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
-`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links.
-The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft")`).
-ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists.
-Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata.
-Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
-
-  **Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications.
-In addition to improved linking by trialists, changes in the PubMed TRN capturing process (such as automated strategies like those developed in this study) and automated indexing of publications in DRKS would make trial registrations and results more findable.
-
-  **Open Data and Code** All code and the final analysis data for this study are available at https://github.com/maia-sh/reg-pub-link.
-Raw data is available at https://github.com/maia-sh/intovalue-data.
-
-  *CLINICAL TRIALS*
-
-  **Background/Aims** Informed clinical guidance and health policy relies on the ability of clinicians, policymakers, and guideline developers to find comprehensive clinical evidence and interlink components of the same clinical trial.
+  **Background/Aims** Informed clinical guidance and health policy relies on the ability of clinicians, policymakers, and guideline developers to find comprehensive clinical evidence and link registrations and publications of the same clinical trial together.
 To support the finding and linking of trial evidence, the World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) ask researchers to provide the trial registration number (TRN) in their publication and a reference to the publication in the registration.
 This straightforward transparency practice costs researchers minimal effort and makes evidence synthesis more thorough and efficient.
-The aim of this cross-sectional study is to evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
-Secondary aims are to develop an automated pipeline that can be applied to other cohorts of trial registrations and publications, and to provide stakeholders, from trialists to registries, with concrete guidance to improve trial links.
+This cross-sectional study aims to evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+Secondary aims are to develop an automated pipeline that can be applied to other cohorts of trial registrations and publications, and to provide stakeholders, from trialists to registries, with guidance to improve trial links.
 
-  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and results publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017.
-We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
-  We evaluated the proportion of trial registrations that reference a known results publication, and the proportion of results publications that report the known TRN in the metadata, abstract, and full-text.
-We also constructed exploratory logistic regression models to investigate the relationship between trial completion date, registry, and registration-publication linking.
+  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and results publications for a cohort of registered, published trials conducted across German UMCs and completed between 2009 and 2017.
+We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and TRNs in publication metadata, abstract, and full-text.
 
-  **Results, positive or negative?
-which values to include?** `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_not_both_ft_abs")` of trials fail to reference TRNs in both the full-text and abstract, in breach of long-standing CONSORT and ICMJE guidelines.
-`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg")` of trial registrations do not contain links to their results publications.
-`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links, so registration and publication could be associated only by manual searching and screening.
-Overall, ClinicalTrials.gov trials were better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists.
+  **Results** In breach of long-standing CONSORT and ICMJE guidelines, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_not_both_ft_abs")` of trials failed to reference TRNs in both the abstract and full-text of their results publications.
+Furthermore, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg")` of trial registrations did not contain links to their results publications.
+`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links, so associating registration and publication required manual searching and screening.
+Overall, ClinicalTrials.gov trials were better linked than DRKS trials, and PubMed and registry infrastructures appear to drive this difference.
 Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata.
@@
 Researchers could easily improve this by copy-and-pasting references between their trial registrations and publications.
-Other stakeholders could make trial evidence even more findable and interlinked, e.g., PubMed could capture additional TRNs using automated strategies (like those developed in this study), and DRKS could automatically index publications from PubMed.
+Other stakeholders could make trial evidence even more findable and linked, e.g., PubMed could capture additional TRNs using automated strategies (like those developed in this study), and DRKS could automatically index publications from PubMed.
 
@@
 
-Data collection, preparation, and analysis was performed in `r r_citation`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> All code, the final analysis data, and a list of all packages used for this study are available at <https://github.com/maia-sh/reg-pub-link>.
+Data collection, preparation, and analysis was performed in `r r_citation`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> Code to recreate the analysis dataset, rerun the analysis, and generate this manuscript is available at <https://github.com/maia-sh/reg-pub-link>.
 Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>.
+A STROBE checklist for reporting cross-sectional studies in provided in the Supplement.
 

### 2021-09-09 docs/manuscript/manuscript.Rmd
Remove commented text from manuscript

@@
 ```{r packages}
+
+# Data munging
 library(dplyr)
 library(tidyr)
+library(purrr)
 library(stringr)
 library(glue)
-library(purrr)
+
+# Files
 library(here)
+library(fs)
 library(readr)
-library(fs)
 
+# Plots and diagrams
 library(ggplot2)
 library(showtext)
+library(DiagrammeR)
 
-library(DiagrammeR)
+# Tables
 library(gtsummary)
 library(gt)
+library(kableExtra)
+
+# Manuscript
 library(papaja)
 library(knitr)
-library(kableExtra)
 ```
@@
 ```{r figures-setup}
-# figures (plots and tables)
 
-# Prepare directory
+# Prepare directory for figures (plots and tables)
 dir_figures <- dir_create(here("docs", "manuscript", "figures"))
 
-# theme_gtsummary_compact()
-
-# Set themes
-lightgray <- "gray75" #"coral3" # #"gray60"
-darkgray <- "gray35" #"gray10"
-registry_colors <- c("DRKS" = darkgray, "ClinicalTrials.gov" = lightgray)
-# registry_linetype <- c("DRKS" = "longdash", "ClinicalTrials.gov" = "solid")
-
-# font_add_google("Poppins", "Poppins")
+# Prepare fonts
 font_add_google("Roboto", "Roboto")
@@
 
-theme_set(theme_light(
-  # base_size = 18,
-  base_family = "Roboto"
-))
+# Prepare colors
+lightgray <- "gray75"
+darkgray <- "gray35"
+registry_colors <- c("DRKS" = darkgray, "ClinicalTrials.gov" = lightgray)
+
+# Set theme
+theme_set(theme_light(base_family = "Roboto"))
 
@@
   text = element_text(family = "Roboto"),
-  # axis.title = element_text(size = 16),
-  axis.text = element_text(family = "Roboto Mono",
-                           size = 8),
+  axis.text = element_text(family = "Roboto Mono",size = 8)
 )
+
 ```
@@
 In this exploratory study, we evaluate the links between registrations and results publications in a cross-section of published clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or DRKS with completion dates between 2009 and 2017.
-We also looked at the relationship between different link types, as well as registry and change in practices over time.<!--# By relying on a cohort of known trial registrations and publications, we are able to draw conclusions about registration-publication links, both with regards to the rate of structured links to known results publications in its registration, and the rate of reporting a trial's known TRN in its publication full-text, abstract, and metadata.--> We developed an automated and scalable approach for data collection and extraction using regular expressions to detect and classify publication identifiers and TRNs, which may be applied to other trial cohorts.
-<!--# We did not pre-specify the analyses.
-We also explored the reporting of additional TRNs, beyond the known associated TRN, in publications, including reported cross-registrations, meaning more than one TRN in more than one registry associated with a single publication.
-We also looked at reporting PMC-version full-text.
--->
+We also looked at the relationship between different link types, as well as registry and change in practices over time.
+We developed an automated and scalable approach for data collection and extraction using regular expressions to detect and classify publication identifiers and TRNs, which may be applied to other trial cohorts.
 
@@
 
-We queried the PubMed Entrez Programming Utilities (E-utilities) API on `r pubmed_query_date` for all trial results PMIDs<!--#, and excluded PMIDs that did not resolve--> [@usnationallibraryofmedicinetigtand; @R-rentrez].
-From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->
+We queried the PubMed Entrez Programming Utilities (E-utilities) API on `r pubmed_query_date` for all trial results PMIDs [@usnationallibraryofmedicinetigtand; @R-rentrez].
+From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.
 
-We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies?
--->[@ourresearch; @R-roadoi].
+We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step [@ourresearch; @R-roadoi].
 We then used the GROBID machine learning library for technical and scientific publications (v0.6.1) via Python to parse the PDFs into machine-readable XMLs [@2008; @lopez2021].
@@
 
-We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (available at <https://github.com/maia-sh/ctregistries/blob/master/inst/extdata/registries.csv>) and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
-To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata<!--# for true positives and true negatives-->.
+We developed regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (available at <https://github.com/maia-sh/ctregistries/blob/master/inst/extdata/registries.csv>) and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text.
+To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata.
 For each source, we first extracted all unique TRN patterns within a source.
@@
 We generated descriptive statistics on trial and publication characteristics, overall and by registry.
-We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables.
+We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.
+To explore change over time, the relationships between types of links, and differences between registries, we ran logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables.
 Regressions were performed as exploratory analyses, and all variables were included in each regression model.
@@
 To explore registry differences for only manually linked publications, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration.
-<!--#add automatic links to regression for pub in registry?
-in which case would need to recode NA as FALSE.
-To explore whether automated links explains registry differences in publication links in the registry...-->
 
@@
 
-Data collection, preparation, and analysis was performed in `r r_citation`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> Code to recreate the analysis dataset, rerun the analysis, and generate this manuscript is available at <https://github.com/maia-sh/reg-pub-link>.
+Data collection, preparation, and analysis was performed in `r r_citation`.
+Code to recreate the analysis dataset, rerun the analysis, and generate this manuscript is available at <https://github.com/maia-sh/reg-pub-link>.
 Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>.
@@
 Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.
-
-<!-- (ref:tbl-trials-caption) **Characteristics of German UMC-conducted trials with published results** A trial was considered randomized if allocation included randomization.
-A trial was considered prospectively registered if registered in the same or previous months to start date.
-Summary results were taken from a structured data field in ClinicalTrials.gov, and determined based on manual inspection for terms such as *Ergebnisbericht* or *Abschlussbericht* in DRKS.
-*Top* journals refer to the journals with the greatest number of trial publications in our sample.
--->
 
@@
 Table \@ref(tab:tbl-link-by-reg) shows registration-publication links overall and by registry.
-<!--# point estimates in table should have CI, right?
--->Figure \@ref(fig:plot-upset-links-reg-pub) shows the percentage of trials with each combination of links between registration and publication.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.--> As PubMed incorporated DRKS as a databank source in 2014, Supplemental Table \@ref(tab:app-tbl-link-by-reg-2014) shows the linking practices in trials published as of 2014 (n = `r inline_text(tbl_link_by_reg_2014, has_iv_trn_secondary_id, stat_0, pattern = "{N}")`), which reflect similar rates to Table \@ref(tab:tbl-link-by-reg) [@nationallibraryofmedicine2018].
+Figure \@ref(fig:plot-upset-links-reg-pub) shows the percentage of trials with each combination of links between registration and publication.
+As PubMed incorporated DRKS as a databank source in 2014, Supplemental Table \@ref(tab:app-tbl-link-by-reg-2014) shows the linking practices in trials published as of 2014 (n = `r inline_text(tbl_link_by_reg_2014, has_iv_trn_secondary_id, stat_0, pattern = "{N}")`), which reflect similar rates to Table \@ref(tab:tbl-link-by-reg) [@nationallibraryofmedicine2018].
 
@@
 ```{r tbl-trials-regression}
-# options(knitr.kable.NA = "**")
-as_kable(tbl_trials_regression, format = "pipe") #%>% 
-  # kable_styling(latex_options = "scale_down")
-  # landscape()
+as_kable(tbl_trials_regression, format = "pipe")
 ```
@@
 
-Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with one link type were generally more likely to have another type of link.
-<!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")`).
+Across the adjusted multivariate models, trials with one link type were generally more likely to have another type of link.
+In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")`).
 DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
@@
 Trials registered in DRKS were also less likely than trials registered in ClinicalTrials.gov to reference the primary outcome publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
-<!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) ClinicalTrials.gov trials that reference the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
+Most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) ClinicalTrials.gov trials that reference the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS.
 Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")` vs.
@@
 
-Linking of trial registrations and results publications plays an important role in research transparency and facilitates comprehensive evidence synthesis and informed health policy decision making.<!--# and can substantially increase the value of research.
---> Poor linking poses a barrier to identifying trial publications via automated approaches and instead requires researchers to perform intensive manual searches to attempt to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
+Linking of trial registrations and results publications plays an important role in research transparency and facilitates comprehensive evidence synthesis and informed health policy decision making.
+Poor linking poses a barrier to identifying trial publications via automated approaches and instead requires researchers to perform intensive manual searches to attempt to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr].
 This responsible research practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.
@@
 Linking practices showed at best minimal improvement over time.
-The upward trend in reporting in full-text and abstracts suggests this practice is gaining traction, however more slowly than advisable per CONSORT and ICMJE guidelines.<!--# shorten--> Trials registered in ClinicalTrials.gov were overall better linked than trials registered in DRKS.<!--# We found substantial differences between registries with ClinicalTrials.gov better linked than DRKS.
---> These differences are in part beyond trialists' control and reliant on bibliometric databases (i.e., PubMed) and registries, namely (1) generating PubMed metadata from TRNs in the abstract or full-text, and (2) automated indexing of publications in the registry.
+The upward trend in reporting in full-text and abstracts suggests this practice is gaining traction, however more slowly than advisable per CONSORT and ICMJE guidelines.
+Trials registered in ClinicalTrials.gov were overall better linked than trials registered in DRKS.
+These differences are in part beyond trialists' control and reliant on bibliometric databases (i.e., PubMed) and registries, namely (1) generating PubMed metadata from TRNs in the abstract or full-text, and (2) automated indexing of publications in the registry.
 
@@
 However, we found that while trials with a TRN in the abstract were indeed more likely to include the TRN in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), TRNs in the full-text were not more likely to appear in the metadata (`r inline_text(aor_si, variable = "has_iv_trn_ft", pattern = "aOR {estimate} [{ci}]")`).
-Furthermore, our data suggest that NLM staff are better at extracting ClinicalTrials.gov than DRKS TRNs (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).<!--# As such, it appears that PubMed metadata misses TRNs from the full-text as well as DRKS trials.-->
+Furthermore, our data suggest that NLM staff are better at extracting ClinicalTrials.gov than DRKS TRNs (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).
 
 Automated indexing of publications in ClinicalTrials.gov accounts for most (`r report_trials(n_trials_link_reg_ctgov_auto, n_trials_link_reg_ctgov, pattern = "{p}%, {n}/{N}")`) publication references in the registry and drives the almost three-fold discrepancy with DRKS (`r inline_text(tbl_link_by_reg, has_reg_pub_link, "ClinicalTrials.gov", pattern = "{p}%, {n}/{N}")` vs.
-`r inline_text(tbl_link_by_reg, has_reg_pub_link, "DRKS", pattern = "{p}%, {n}/{N}")`).<!--# Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (`r inline_text(tbl_reg_pub_link_manual, has_reg_pub_link, column = "stat_1", pattern = "{p}%, {n}/{N}")`)--> Currently, DRKS allows for only manual submission of references by trialists and does not index publications.<!--# add?--> Furthermore, registrations may refer to additional publications beyond results (e.g., up to 72 publications for one ClinicalTrials.gov registration in our sample), so registry metadata should encode publication type to support quick identification of results.
+`r inline_text(tbl_link_by_reg, has_reg_pub_link, "DRKS", pattern = "{p}%, {n}/{N}")`).
+Currently, DRKS allows for only manual submission of references by trialists and does not index publications.
+Furthermore, registrations may refer to additional publications beyond results (e.g., up to 72 publications for one ClinicalTrials.gov registration in our sample), so registry metadata should encode publication type to support quick identification of results.
 While such publication type metadata is currently available in both registries, it is not systematically used, and many publications are categorized generically as a "paper" (DRKS) or "reference" (ClinicalTrials.gov).
@@
 For example, in our visual inspection of the PubMed metadata TRNs, we found `r si_drks_misformatted` severely misformatted DRKS TRNS (i.e., numbers only with no preceding letters), which we could only identify as DRKS TRNs using additional metadata and which prevented the regex from classifying them as TRNs based on the pattern alone.
-<!--#Classifying these cases as true negative, our regexes correctly classified all TRNs and non-TRNs in our sample's the PubMed metadata secondary identifier field (sensitivity and specificity of 100%).
-remove since not about trn formatting alone?
-Furthermore, ClinicalTrials.gov presumably relies not only on the reported TRN but also on the registry, and as such failed to index a publication (PMID: 26417987) despite the TRN (NCT01614301) appearing in the PubMed metadata, as the associated databank was incorrect (i.e., GENBANK).
--->
 
@@
 Furthermore, we relied on IntoValue for trial deduplication and bona fide results publications.
-The dataset may have had a small number of unaccounted for cross-registrations (e.g., DRKS00004156 and NCT00215683) and publications that are not trial results (e.g., systematic reviews, conference abstract books, etc.<!--# DRKS00009565 associated publication is 10.1097/aap.0000000000000573, a systematic review, with many reported TRNs-->).
+The dataset may have had a small number of unaccounted for cross-registrations (e.g., DRKS00004156 and NCT00215683) and publications that are not trial results (e.g., systematic reviews, conference abstract books, etc.).
 Finally, this automated approach faces limitations of the software on which it is built.
 While PubMed updated their website in 2020, their API reflects the previous backend and has subtle differences to the web version (personal communication via NLM Helpdesk Ticket CAS-575119-Y6D6Y4).
-While GROBID algorithms are trained on academic papers and have been used in large-scale bibliometric projects [e.g., @wang2020ac], parsing PDFs to XMLs may introduce some errors.<!--#Finally, the approach relies on quality input text data, and since journal articles are still most commonly distributed as PDFs, parsing errors may reduce text quality.
-We minimized this risk by relying on GROBID to parse full-text PDFs into XML with the body separated out from the abstract, header, and back.
-While GROBID algorithms are trained on academic papers and have been used in large-scale bibliometric projects [e.g., @wang2020ac], it does not perfectly classify the full-text main body.
-As such, TRNs appearing in the main body main be mistakenly excluded and TRNs appearing in the header or back material may be mistakenly included.-->
+While GROBID algorithms are trained on academic papers and have been used in large-scale bibliometric projects [e.g., @wang2020ac], parsing PDFs to XMLs may introduce some errors.
 
@@
 In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is straightforward and can be improved with action across stakeholders.
-<!--#Drawing from the study findings and additional context, -->Table \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.
+Table \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.
 
 ```{r tbl-stakeholders}
-# https://www.anthonyschmidt.co/post/2020-06-03-making-apa-tables-with-gt/
 knitr::include_graphics(path(dir_figures, "tbl-stakeholders.pdf"))
@@
 DS is a member of the Sanofi Advisory Bioethics Committee and receives an honorarium for his contribution to meetings.
-<!--# BGC?
--->
 

### 2021-09-09 analysis/tbl_stakeholders.R

@@ -1,5 +1,6 @@
-# Prepare actions and examples --------------------------------------------
+# https://www.anthonyschmidt.co/post/2020-06-03-making-apa-tables-with-gt/
 
+# Prepare actions and examples --------------------------------------------
 
 action_researchers <- "
 - Include TRN in abstract and full-text per CONSORT and ICMJE guidelines

### 2021-09-14 docs/manuscript/manuscript.Rmd
Update abstract and add zenodo for raw data

@@
 abstract: |
-  **Background/Aims** Informed clinical guidance and health policy relies on the ability of clinicians, policymakers, and guideline developers to find comprehensive clinical evidence and link registrations and publications of the same clinical trial together.
+  **Background/Aims** Informed clinical guidance and health policy relies on clinicians, policymakers, and guideline developers finding comprehensive clinical evidence and linking registrations and publications of the same clinical trial.
 To support the finding and linking of trial evidence, the World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) ask researchers to provide the trial registration number (TRN) in their publication and a reference to the publication in the registration.
-This straightforward transparency practice costs researchers minimal effort and makes evidence synthesis more thorough and efficient.
-This cross-sectional study aims to evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
+This practice costs researchers minimal effort and makes evidence synthesis more thorough and efficient.
+Nevertheless, trial evidence appears inadequately linked, and the extent of trial links in Germany remains unquantified.
+This cross-sectional study aims to evaluate links between registrations and publications across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS).
 Secondary aims are to develop an automated pipeline that can be applied to other cohorts of trial registrations and publications, and to provide stakeholders, from trialists to registries, with guidance to improve trial links.
@@
   **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and results publications for a cohort of registered, published trials conducted across German UMCs and completed between 2009 and 2017.
-We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and TRNs in publication metadata, abstract, and full-text.
+We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and TRNs in publication metadata, abstracts, and full-texts.
 
@@
 `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links, so associating registration and publication required manual searching and screening.
-Overall, ClinicalTrials.gov trials were better linked than DRKS trials, and PubMed and registry infrastructures appear to drive this difference.
-Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata.
+Overall, ClinicalTrials.gov trials were better linked than DRKS trials; PubMed and registry infrastructures appear to drive this difference.TRNs were more likely to be transferred to PubMed metadata from abstracts for ClinicalTrials.gov trials than for DRKS trials.
 Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.
@@
   **Conclusions** German UMCs have not comprehensively linked trial registrations and publications, despite established recommendations.
-This shortcoming threatens the quality of evidence synthesis and medical practice, and burdens researchers to manually search and link trial data.
+This shortcoming threatens the quality of evidence synthesis and medical practice, and burdens researchers with manually searching and linking trial data.
 Researchers could easily improve this by copy-and-pasting references between their trial registrations and publications.
-Other stakeholders could make trial evidence even more findable and linked, e.g., PubMed could capture additional TRNs using automated strategies (like those developed in this study), and DRKS could automatically index publications from PubMed.
+Other stakeholders could build on this practice, e.g., PubMed could capture additional TRNs using automated strategies (like those developed in this study), and DRKS could automatically index publications from PubMed.
 
 keywords          : "clinical trials, registration, reporting, meta-research, research transparency"
-wordcount         : "3604"
+wordcount         : "3564"
 bibliography      : ["references.bib", "r-references.bib"]
@@
 classoption       : "man"
-output            : papaja::apa6_word
+output            : papaja::apa6_pdf
 header-includes   :
@@
 Code to recreate the analysis dataset, rerun the analysis, and generate this manuscript is available at <https://github.com/maia-sh/reg-pub-link>.
-Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>.
+Raw data (with the exception of the full-text of publications) is available at <https://doi.org/10.5281/zenodo.5506434>, and code for generating the raw data is available at <https://github.com/maia-sh/intovalue-data>.
 A STROBE checklist for reporting cross-sectional studies in provided in the Supplement.

### 2021-09-14 docs/manuscript/manuscript.Rmd
Fix typo

@@
 The sampling strategies used in previous studies limit our ability to draw conclusions about registration-publication linking.
-Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included publications beyond trial results (e.g., protocols), as well as trials with unknown registration rates, as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo].
+Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included publications beyond trial results (e.g., protocols), as well as trials with unknown registration rates, as not all trials are registered [@trinquart2018bm; @gopal2018t; @denneny2019bo].
 Registration-based cohorts included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b].

### 2022-10-13 docs/manuscript/manuscript.Rmd
Fix decimal point

@@
 
-Per our inclusion criteria all trials in our sample (n = `r n_trials_ft`) were registered and had a publication, however only `r report_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning the publication linked in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata.
-Disregarding metadata, which is largely beyond trialists' control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_trials(n_trials_link_ft_abs_reg)`.
-An additional `r report_trials(n_trials_link_ft_abs - n_trials_link_ft_abs_reg)` met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
-In contrast, we found `r report_trials(n_trials_link_none)` trials with no links in either the registration or the publication.
+Per our inclusion criteria all trials in our sample (n = `r n_trials_ft`) were registered and had a publication, however only `r report_trials(n_trials_link_full, digits = 1)` trials had the most comprehensive registration-publication linking, meaning the publication linked in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata.
+Disregarding metadata, which is largely beyond trialists' control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_trials(n_trials_link_ft_abs_reg, digits = 1)`.
+An additional `r report_trials(n_trials_link_ft_abs - n_trials_link_ft_abs_reg, digits = 0)` met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract.
+In contrast, we found `r report_trials(n_trials_link_none, digits = 1)` trials with no links in either the registration or the publication.
 The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_trials(n_trials_link_ft_only)` trials.
